---
title: "Analysis code associated with Hoffman et al. Decreases in lake and reservoir water level..." # MEL consider linking title to manuscript?
author: Kathryn K. Hoffman
format: html
editor: visual
---

This script reads in the extracted data from the systematic literature review.

## Libraries

Install and load dependencies.

```{r packages}
#| message: false
#| warning: false
# MEL remember to add install.packages statements for each of these!

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")

pacman::p_load(janitor, tidyverse, ggplot2, dplyr, tidyr, naniar, plotly, patchwork, here, #plotting and data manipulation packages
               rnaturalearth, rnaturalearthdata, mapview, sf, #mapping packages
               rstatix, rfriend) #chi-square analysis packages

#MASS needs to be installed, but is installed and loaded later because of conflicts with dplyr


```

#Read in data

```{r read in data}
data <- "Systematic/Combined_08Oct.csv"
init_data <- read.csv(here(data))
init_data <- clean_names(init_data)
head(init_data)
```

# QAQC

## Excluded papers

When more than one paper included the same monitoring period and location (duplicate data) in a given lake or reservoir, we retained only the paper that had the longer data collection period.

Papers were excluded for the following reasons:

Abirhire 2019: Has 2 years of data, Abirhire 2023 has 8 years of data BUT it's not presented in a way that is particularly usable \> doesn't fit as well. Excluded 2019 because that was our criteria \#### MEL: sorry, not following this - are you saying the 2023 paper data isn't presented in a way that's particularly usable? and which criterion was 2019 excluded on?

Beaver 2013: contacted the author and there were no water level data available

Braga 2015: duplicate data with Medeiros paper, and Medeiros paper has an extra year of data, so retaining Medeiros and dropping Braga

Braga 2020: duplicated data with deMendonca paper, deMendonca paper has a few more months of data, so retaining deMendonca paper and dropping Braga

Cao 2016 & Jia 2022: Both Poyang papers, which were excluded from our analysis because Poyang has a potomophase

Christensen 2015: Compared two rule curves, but didn't compare a high and low condition - the later rule curve generally meant higher water level, but not for all of the reservoirs. There's not a way to answer our questions with the data provided.

Geraldes 2007 - this paper presented two reservoirs. For the Serra Serrada Reservoir, the data was duplicate Geraldes 2005. Azivo Reservoir is still included from Geraldes 2007.

Hart 2004: Data is either duplicated or more data is available in Hart 2006, so excluding Hart 2004

Kisand 2004 & Wang 2022: these papers were not clear enough to extract data from.

Lacerda 2018: same data are available in Santos, and Lacerda has a stronger focus on remote sensing, which was not our target data type

Naselli Flores 2003: duplicate data with other Naselli Flores 2003 paper, retained the one with more available data

Nõges 2010 & Tuvikene 2011 & Noges 1999: duplicate data with other Nõges papers

Sakharova 2018: 3 years of data, Mineeva has longer time series

Sharip 2019: paper got duplicated, removed the duplicate version. Also, Bera Lake was not more than 4.5m deep, so does not meet our depth criterion and was excluded.

```{r exclude papers}
#| warning: false
data_excl <- init_data |> 
  filter(!name_yr %in% c("Beaver 2013","Jia 2022","Cao 2016","Wang 2022","Kisand 2004", "Hart 2004", "Nõges 2010", "Tuvikene 2011", "Christensen 2015", "Braga 2015", "Braga 2020", "Lacerda 2018", "Abirhire 2019", "Sakharova 2018", "Noges 1999")) |>
  filter(!res_name == "Serra Serrada Reservoir") |> #Geraldes 2007
  filter(!res_name == "Bera Lake") |> #Sharip 2019
  filter(!title %in% c("Steady-state assemblages in a Mediterranean hypertrophic reservoir. The role of Microcystis ecomorphological variability in maintaining an apparent equilibrium")) |>
  # filter(res_name != "Dourado" | name_yr != "deMednonca 2018") |> #Naselli-Flores 2003
#Data is either duplicated or more data is available in Hart 2006, so excluding Hart 2004
  filter(!notes %in% c("Ignore this version, everything has been added to multiple reservoirs sheet! KKH")) #Sharip duplicate
#Beaver no WL data
#Jia 
#duplicate Sharip

excluded_papers <- anti_join(init_data, data_excl)

duplicate_waterbodies <- data_excl |> 
  select(name_yr, res_name, country, study_years, consecutive, phys_response, notes) #used this to order data alphabetically and manually check for dups that were not named exactly the same

```

## Fixing typos and assigning trophic levels

Fix typos and assign trophic levels based on chla and P concentrations if available.

```{r fix some data}
data <- data_excl |> 
  mutate(tp = mean_tp) |> 
  separate_wider_delim(mean_tp, " ", names = c("tp_ugL", "unit_tp"), too_few = "align_start") |> 
  mutate(mean_chla = ifelse(mean_chla == "0.07 to 34.62 ug/L", "8", mean_chla)) |> #change Rai 1978 to 8ug/L, actual range between 0 and 14 ug/L, original range from pheophytin
  mutate(mean_chla = ifelse(mean_chla == "15.7ug/L", "15.7 ug/L", mean_chla)) |> #fix missing space in chla value
  mutate(chla = mean_chla) |> 
  separate_wider_delim(mean_chla, " ", names = c("chla_ugL", "unit_chla"), too_few = "align_start") |>
  mutate(res_name = ifelse(res_name == "Rybinsk", "Rybinsk Reservoir", #standardize reservoir names
                    ifelse(res_name == "Cruzeta man-made lake", "Cruzeta", 
                    ifelse(res_name == "Castanhao", "Castanhão reservoir",
                    ifelse(name_yr == "Bouvy 2003", "Tapacura Reservoir", res_name))))) |> 
  filter((!is.na(cov_num)|!is.na(num_water)|!is.na(lat_dd))) |> #remove empty or duplicate columns from the df
  mutate(fullpond_sa = ifelse(fullpond_sa == "~600", 600, fullpond_sa), #fix character in numeric columns
         fullpond_sa = as.numeric(fullpond_sa)) |> 
  mutate(obs_maxdepth = ifelse(obs_maxdepth == "SEK had 12, 12 max at sampling station", 12, obs_maxdepth), #fix text in numeric columns
         mindepth = ifelse(mindepth == "SEK had 9, 9 was min at sampling station", 9, mindepth)) |> 
  mutate(elev_decrease_max = ifelse(grepl("2001", elev_decrease_max), "-6", elev_decrease_max), # to use the 2003 value given in the elev_decrease_max column. 
         elev_increase_max = ifelse(elev_increase_max == 0, NA, elev_increase_max)) |> #to remove increases of 0
  mutate(long_dd = ifelse(res_name == "Lake Buchanan", -98.4, long_dd), #fix incorrect latlongs
         lat_dd = ifelse(lat_dd == 103.466667 & res_name == "Zipingpu Reservoir", 30.983333, lat_dd), 
         long_dd = ifelse(long_dd == 30.983333 & res_name == "Zipingpu Reservoir", 103.466667, long_dd),
         long_dd = ifelse(long_dd == 6.77888889 & res_name == "Serra Serrada", -6.77888889, long_dd)) |> 
  mutate(tp_ugL = as.numeric(tp_ugL)) |> #add in trophic state based on phosphorus and chlorophyll values. 
  mutate(mean_tp = case_when( #correct units so that mean_tp is in ug/L
    unit_tp == "mg/L" ~ tp_ugL * 1000, 
    unit_tp == "mg/m3" ~ tp_ugL / 1000, 
    TRUE ~ tp_ugL)) |> 
  mutate(chla_ugL = as.numeric(chla_ugL)) |>  #correct units so that chla_ugL is in ug/L
  mutate(mean_chla = case_when(
    unit_chla == "mg/L" ~ chla_ugL * 1000, 
    unit_chla == "mg/m3" ~ chla_ugL / 1000, 
     TRUE ~ chla_ugL)) |> 
  # fix text and notes in trophic column
  mutate(trophic = ifelse(trophic == "Other: according to authors, varies according to water level; never more than mesotrophic unless water level is low", "mesotrophic", trophic)) %>% 
  mutate(trophic = ifelse(trophic == "", "not reported", trophic)) |> 
  mutate(chla_ugL = as.numeric(chla_ugL)) |>  
  mutate(tp_ugL = as.numeric(tp_ugL)) |> 
  #assign trophic status (if not reported) based on chlorophyll and total phosphorus concentrations
  mutate(trophic_status_chla = ifelse(trophic == "not reported" & chla_ugL <= 2.6, "oligotrophic", 
                               ifelse(trophic == "not reported" & chla_ugL >2.6 & chla_ugL <=7.3, "mesotrophic", 
                               ifelse(trophic == "not reported" & chla_ugL >7.3 & chla_ugL <= 56, "eutrophic", 
                               ifelse(trophic == "not reported" & chla_ugL >56, "hypereutrophic",
                               ifelse(trophic == "not reported" & is.na(chla_ugL), trophic, trophic)))))) %>% 
  mutate(trophic_status_tp = ifelse(tp_ugL <= 12, "oligotrophic", 
                             ifelse(tp_ugL >=12 & tp_ugL <=24, "mesotrophic", 
                             ifelse(tp_ugL >24 & tp_ugL <= 70, "eutrophic", 
                             ifelse(tp_ugL >70, "hypereutrophic",
                             ifelse(is.na(tp_ugL), trophic, NA)))))) %>%  
  mutate(trophic_status_combined = ifelse(trophic == "not reported" & !is.na(trophic_status_tp), trophic_status_tp,
                                   ifelse(trophic == "not reported" & !is.na(trophic_status_chla), trophic_status_chla, trophic))) |> 
  #combine multi trophic state options
  mutate(trophic_status_reduced = ifelse(trophic_status_combined %in% c("oligotrophic"), "oligotrophic", 
                                  ifelse(trophic_status_combined %in% c("oligo-mesotrophic", "mesotrophic"), "mesotrophic", 
                                  ifelse(trophic_status_combined %in% c("meso-eutrophic", "eutrophic"), "eutrophic", 
                                  ifelse(trophic_status_combined %in% c("hypereutrophic", "eu-hypereutrophic"), "hypereutrophic", "not reported"))))) |> 
  filter(res_name != "") #remove any rows that don't have a reservoir name because they are duplicate with reservoirs in the multiple reservoir sheet
write_csv(data, here("Systematic/qaqced.csv"))
```

# Dataset summary statistics

```{r quick study info}
#number of waterbodies
reservoirs <- data |> select(c(res_name)) |> unique()
#there are a total of 4 TGRs, so there are 65-3 reservoirs, = 62 total lakes!
countries <- data |> select(c(country)) |> unique() #25 countries - Argentina/Uruguay, Estonia/Russia, one blank = 22 total countries
studies <- data |> select(c(name_yr)) |> unique() #61 studies, some duplicate authors but no duplicate studies.
```

# Plots

## Phytoplankton drivers

Physical variables that we considered as potential drivers of phytoplankton community change: water temperature, water level, Secchi disk depth/transparency, light (PAR, photic zone), turbidity, total suspended solids, total dissolved solids, and other

Physical variables that we considered as potential drivers of phytoplankton community change: pH, DO, DOC, DIC, DOM, POM, TP, SRP, ORP (orthophosphate), TN, TKN, NO3, NO2, NH4, DIN, metals, conductivity, salinity, and other

```{r}
drivers <- data |> 
  mutate(tp = ifelse(grepl("TP", chem_drive), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_drive), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_drive), 1, 0), 
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         do = ifelse(grepl("DO", chem_drive), 1, 0),
         srp = ifelse(grepl("SRP", chem_drive), 1, 0),
         no3 = ifelse(grepl("NO3", chem_drive), 1, 0),
         cond = ifelse(grepl("conductivity", chem_drive), 1, 0),
         metals = ifelse(grepl("metals", chem_drive), 1, 0),
         orp = ifelse(grepl("ORP", chem_drive), 1, 0),
         no2 = ifelse(grepl("NO2", chem_drive), 1, 0),
         dic = ifelse(grepl("DIC", chem_drive), 1, 0),
       dom = ifelse(grepl("DOM", chem_drive), 1, 0),
       pom = ifelse(grepl("POM", chem_drive), 1, 0),
       tkn = ifelse(grepl("TKN", chem_drive), 1, 0),
         doc = ifelse(grepl("DOC", chem_drive), 1, 0),
         din = ifelse(grepl("DIN", chem_drive), 1, 0),
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         salinity = ifelse(grepl("salinity", chem_drive), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_drive), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_drive), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_drive), 1, 0),
         transparency = ifelse(grepl("transparency", phys_drive), 1, 0),
         turb = ifelse(grepl("turbidity", phys_drive), 1, 0),
         light = ifelse(grepl("PAR", phys_drive), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid", phys_drive), 1, 0), 
         strat = ifelse(grepl("mix|stab|strat", phys_drive), 1, 0),
         restime = ifelse(grepl("residence", phys_drive), 1, 0), 
         flush_discharge = ifelse(grepl("flush|discharge", phys_drive), 1, 0), 
         inflow = ifelse(grepl("inflow", phys_drive), 1, 0),
         precip = ifelse(grepl("precip|rain", phys_drive), 1, 0),
         air_temp = ifelse(grepl("air", phys_drive), 1, 0),
         other_phys = ifelse(grepl("Other", phys_drive), 1, 0))

other_phys <- drivers |> 
  filter(other_phys == 1) |> 
  select(name_yr, title, num_water, phys_drive)

  unique(data$phys_response)
  
  #WLchangevars is whether or not vars changed under WL change
WLchangevars <- data |> 
mutate(not_assessed = ifelse(grepl("not assessed", phys_response), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_response), 1, 0),
         transparency = ifelse(grepl("transparency", phys_response), 1, 0),
         turb = ifelse(grepl("turbidity", phys_response), 1, 0),
         light = ifelse(grepl("PAR", phys_response), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_response), 1, 0), 
       #this group of variables is commented out and not assessed because it was not part of our extraction list  
       #strat = ifelse(grepl("mix|stab|strat", phys_response), 1, 0),
         #restime = ifelse(grepl("residence", phys_response), 1, 0), 
         #flush_discharge = ifelse(grepl("flush|discharge", phys_response), 1, 0), 
         #inflow = ifelse(grepl("inflow", phys_response), 1, 0),
         #precip = ifelse(grepl("precip|rain", phys_response), 1, 0),
         #air_temp = ifelse(grepl("air", phys_response), 1, 0),
         other_phys = ifelse(grepl("Other", phys_response), 1, 0), 
       watercolor = ifelse(grepl("water color", phys_response), 1, 0)) |> 
mutate(tp = ifelse(grepl("TP", chem_response), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_response), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_response), 1, 0), 
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         do = ifelse(grepl("DO", chem_response), 1, 0),
         srp = ifelse(grepl("SRP", chem_response), 1, 0),
         no3 = ifelse(grepl("NO3", chem_response), 1, 0),
         cond = ifelse(grepl("conductivity", chem_response), 1, 0),
         metals = ifelse(grepl("metals", chem_response), 1, 0),
         orp = ifelse(grepl("ORP", chem_response), 1, 0),
         no2 = ifelse(grepl("NO2", chem_response), 1, 0),
         doc = ifelse(grepl("DOC", chem_response), 1, 0),
       dic = ifelse(grepl("DIC", chem_response), 1, 0),
       dom = ifelse(grepl("DOM", chem_response), 1, 0),
       pom = ifelse(grepl("POM", chem_response), 1, 0),
       tkn = ifelse(grepl("TKN", chem_response), 1, 0),
         din = ifelse(grepl("DIN", chem_response), 1, 0),
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         salinity = ifelse(grepl("salinity", chem_response), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_response), 1, 0)) 

#spreadvars is whether or not the var was measured
spreadvars <- data |> 
  mutate(filt_chla = ifelse(grepl("filtered chlorophyll", bio_vars), 1, 0),                       #biotic variables
         phyto_micros = ifelse(grepl("phytoplankton microscopy", bio_vars), 1, 0), 
         single_cyano_taxon = ifelse(grepl("single cyanobacterial taxon", bio_vars), 1, 0), 
         sensor_chla = ifelse(grepl("sensor", bio_vars), 1, 0),
         other_bio = ifelse(grepl("Other", bio_vars), 1, 0)) |> 
  mutate(tp = ifelse(grepl("TP", chem_vars), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_vars), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_vars), 1, 0), 
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         do = ifelse(grepl("DO", chem_vars), 1, 0),
         srp = ifelse(grepl("SRP", chem_vars), 1, 0),
         no3 = ifelse(grepl("NO3", chem_vars), 1, 0),
         cond = ifelse(grepl("conductivity", chem_vars), 1, 0),
         metals = ifelse(grepl("metals", chem_vars), 1, 0),
         orp = ifelse(grepl("ORP", chem_vars), 1, 0),
         no2 = ifelse(grepl("NO2", chem_vars), 1, 0),
         doc = ifelse(grepl("DOC", chem_vars), 1, 0),
          dic = ifelse(grepl("DIC", chem_vars), 1, 0),
       dom = ifelse(grepl("DOM", chem_vars), 1, 0),
       pom = ifelse(grepl("POM", chem_vars), 1, 0),
       tkn = ifelse(grepl("TKN", chem_vars), 1, 0),
         din = ifelse(grepl("DIN", chem_vars), 1, 0),
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         salinity = ifelse(grepl("salinity", chem_vars), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_vars), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_vars), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_vars), 1, 0),
         transparency = ifelse(grepl("transparency", phys_vars), 1, 0),
         turb = ifelse(grepl("turbidity", phys_vars), 1, 0),
         light = ifelse(grepl("PAR", phys_vars), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing #these are commented out below
        tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_vars), 1, 0), 
         #strat = ifelse(grepl("mix|stab|strat", phys_vars), 1, 0),
        # restime = ifelse(grepl("residence", phys_vars), 1, 0), 
        # flush_discharge = ifelse(grepl("flush|discharge", phys_vars), 1, 0), 
        # inflow = ifelse(grepl("inflow", phys_vars), 1, 0),
        # precip = ifelse(grepl("precip|rain", phys_vars), 1, 0),
        # air_temp = ifelse(grepl("air", phys_vars), 1, 0),
         other_phys = ifelse(grepl("Other", phys_vars), 1, 0)) 
        # watercolor = ifelse(grepl("water color", phys_vars), 1, 0))

```

Look at phytoplankton and cyanobacterial drivers separately. Phytoplankton first.

```{r phyto drivers}
#separate out phytoplankton and cyanobacteria drivers
phyto_only_drivers <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:125)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_drivers <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:125)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

#WL change vars
phyto_only_WLchangevars <- WLchangevars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_WLchangevars <- WLchangevars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

#not WL change vars! This is anything measured
phyto_only_vars <- spreadvars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:124)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_vars <- spreadvars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:124)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

chemvars_drive_phytos <- phyto_only_drivers |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))

#only including variables that were measured at least 10 times in the extracted data
chemvars_phytos <- phyto_only_vars |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))

physvars_drive_phytos <- phyto_only_drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))

# physvarsWL_phytos <- phyto_only_WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp))

physvars_phytos <- phyto_only_vars |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))
# summary(vars) 

freqchem_drive_phytos <- as.data.frame(colSums(chemvars_drive_phytos))
chemvars_drive_phytos <- setNames(cbind(rownames(freqchem_drive_phytos), freqchem_drive_phytos, row.names = NULL), 
         c("names", "sum"))
# freqchemWL_phyto <- as.data.frame(colSums(chemvarsWL_phytos))
freqchem_phyto <- as.data.frame(colSums(chemvars_phytos))
# chemvarsWL_phytos <- setNames(cbind(rownames(freqchemWL), freqchemWL_phyto, row.names = NULL), 
#          c("names", "sum"))
chemvars_phytos <- setNames(cbind(rownames(freqchem_phyto), 
                                  freqchem_phyto, row.names = NULL), c("names", "sum"))

# chemvarsWL |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +  ggtitle("Chem variables that changed in response to WL")
# want a comparison with what was measured!
chemvars1_phytos <- chemvars_phytos |> 
  rename(measured = sum, 
         name = names)

# drivers identified out of all vars measured
chemicalvars_phytos <- cbind(chemvars1_phytos, chemvars_drive_phytos)

chem <- chemicalvars_phytos |> 
  filter(!name == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("Chemical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#99bbab", "#ddefcd"),
  labels = c("Important", "Not Important"), 
  name = "Phytoplankton: \nDid authors identify \nthis driver as important \nor not important?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

#physical variables

freqphys_drive <- as.data.frame(colSums(physvars_drive_phytos))
# freqphysWL <- as.data.frame(colSums(physvarsWL_phytos))
freqphys_phytos <- as.data.frame(colSums(physvars_phytos))

physvars2_drive_phytos <- setNames(cbind(rownames(freqphys_drive), freqphys_drive, row.names = NULL), 
         c("names", "sum"))

physvars1_phytos <- setNames(cbind(rownames(freqphys_phytos), freqphys_phytos, row.names = NULL), 
         c("name", "measured"))

physvars_phytos <- cbind(physvars1_phytos, physvars2_drive_phytos)

phys <- physvars_phytos |> 
    select(c(names, measured, sum)) |> 
  filter(!names == "other_phys") |> 
  rename(id = sum) |> 
   mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("Physical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#99bbab", "#ddefcd"),
  labels = c("Important", "Not Important"), 
  name = "Phytoplankton: \nDid authors identify \nthis driver as important \nor not important?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

phytos <- phys/chem
phytos

ggsave(paste0(here(),"/Systematic/Plots/phytovars.jpg"), plot = phytos, height = 5, width = 6.5, units = "in")

```

Physical and Chemical drivers for cyanobacteria

```{r cyanos vars only}
# chemvars_drive_cyanos <- cyano_only_drivers |> select(c(other_chem, tp, tn, nh4, ph, do, srp, no3, cond, metals, orp, no2, doc, din, ph, salinity, dom, pom, dic, tkn))

# only vars that have more than 10
chemvars_drive_cyanos <- cyano_only_drivers |> select(c(cond, do, nh4, no2, no3, ph, srp, tn, tp))

chemvars_cyanos <- cyano_only_vars |> select(c(cond, do, nh4, no2, no3, ph, srp, tn, tp))

physvars_drive_cyanos <- cyano_only_drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))

# physvarsWL_phytos <- phyto_only_WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp))

physvars_cyanos <- cyano_only_vars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, water_level))

physvars_cyanos <- cyano_only_vars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, water_level))
# summary(vars) 

freqchem_drive_cyanos <- as.data.frame(colSums(chemvars_drive_cyanos))
chemvars_drive_cyanos <- setNames(cbind(rownames(freqchem_drive_cyanos), freqchem_drive_cyanos, row.names = NULL), 
         c("names", "sum"))
# freqchemWL_cyano <- as.data.frame(colSums(chemvarsWL_cyanos))
freqchem_cyano <- as.data.frame(colSums(chemvars_cyanos))
# chemvarsWL_cyanos <- setNames(cbind(rownames(freqchemWL), freqchemWL_cyano, row.names = NULL), 
#          c("names", "sum"))
chemvars_cyanos <- setNames(cbind(rownames(freqchem_cyano), 
                                  freqchem_cyano, row.names = NULL), c("names", "sum"))

# chemvarsWL |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +  ggtitle("Chem variables that changed in response to WL")
# want a comparison with what was measured!
chemvars1_cyanos <- chemvars_cyanos |> 
  rename(measured = sum, 
         name = names)
chemicalvars_cyanos <- cbind(chemvars1_cyanos, chemvars_drive_cyanos)


chemvars_cyanos_drivers <- chemicalvars_cyanos |> 
    select(c(names, measured, sum)) |> 
  filter(!names == "other_phys") |> 
  rename(id = sum) |> 
   mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
 # ggtitle("Chemical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#4dd0e1","#b2ebf2"),
  labels = c("Important", "Not Important"), 
  name = "Cyanobacteria: \nDid authors identify \nthis driver as important \nor not important?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

#physical variables

freqphys_drive <- as.data.frame(colSums(physvars_drive_cyanos))
# freqphysWL <- as.data.frame(colSums(physvarsWL_cyanos))
freqphys_cyanos <- as.data.frame(colSums(physvars_cyanos))

physvars2_drive_cyanos <- setNames(cbind(rownames(freqphys_drive), freqphys_drive, row.names = NULL), 
         c("names", "sum"))

physvars1_cyanos <- setNames(cbind(rownames(freqphys_cyanos), freqphys_cyanos, row.names = NULL), 
         c("name", "measured"))

physvars_cyanos <- cbind(physvars1_cyanos, physvars2_drive_cyanos)

physvars_cyanos_drivers <- physvars_cyanos |> 
    select(c(names, measured, sum)) |> 
  filter(!names == "other_phys") |> 
  rename(id = sum) |> 
   mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
 # ggtitle("Physical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#4dd0e1","#b2ebf2"),
  labels = c("Important", "Not Important"), 
  name = "Cyanobacteria: \nDid authors identify \nthis driver as important \nor not important?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

cyanos <- physvars_cyanos_drivers/chemvars_cyanos_drivers
cyanos

ggsave(paste0(here(),"/Systematic/Plots/cyanodrivervars.jpg"), plot = cyanos, height = 5, width = 6.5, units = "in")

combined <- (physvars_cyanos_drivers/chemvars_cyanos_drivers) | (phys / chem)
combined <- ((phys/physvars_cyanos_drivers) | (chem/chemvars_cyanos_drivers)) + plot_layout(guides = "collect") &
  theme(
    legend.position = "right",
    legend.box = "vertical"
  )
combined

ggsave(paste0(here(),"/Systematic/Plots/drivers_cyanophyto.jpg"), plot = combined, height = 4, width = 8, units = "in")

```

## Map

Make map figure.

```{r}

na_strings <- c("NA", "N A", "N / A", "N/A", "N/ A", "Not Available", "NOt available", "", "not reported")

 latlong <- data %>%
  mutate(
    study_years = ifelse(
      grepl("36 \\(satellite data\\); 12 \\(field data\\)", study_years),"12", 
      ifelse(grepl("36 \\(satellite data\\); 2 \\(field data\\)", study_years),"2",study_years)), 
    study_years = as.numeric(study_years)) |> 
  mutate(depth_m = ifelse(!is.na(fullpond_maxdepth), fullpond_maxdepth, obs_maxdepth), 
         depth_m = as.numeric(depth_m), 
         years_studied = ifelse(study_years == 1, "1", 
                                ifelse(study_years >1 & study_years <=5, "2-5", 
                                       ifelse(study_years >5 & study_years <= 10, "6-10", 
                                              ifelse(study_years > 10, "more than 10", NA)))), 
         years_studied = factor(years_studied, levels = c("1", "2-5", "6-10", "more than 10"))) |> 
  select("lat_dd", "long_dd", "res_name", "decrease_phyto_response", "increase_phyto_response", "decrease_cyano_response", "increase_cyano_response", "depth_m", "trophic_status_reduced", "years_studied") |> 
  mutate(lat_dd = ifelse(lat_dd == 103.466667 & res_name == "Zipingpu Reservoir", 30.983333, lat_dd)) |> 
  mutate(long_dd = ifelse(long_dd == 30.983333 & res_name == "Zipingpu Reservoir", 103.466667, long_dd)) |> 
  mutate(long_dd = ifelse(long_dd == 6.77888889 & res_name == "Serra Serrada", -6.77888889, long_dd)) |> 
    replace_with_na_all(condition = ~.x %in% na_strings) |> 
    subset(!is.na(long_dd)) |>  
  subset(!is.na(lat_dd)) |> 
  mutate(increase_phyto_response = ifelse(increase_phyto_response == "Not measured"|increase_phyto_response == ""|(grepl("Other", increase_phyto_response)), NA, increase_phyto_response)) |>
    mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured"|decrease_phyto_response == ""|(grepl("Other", decrease_phyto_response)), NA, decrease_phyto_response)) |>
   mutate(increase_cyano_response = ifelse(increase_cyano_response == "Not measured"|increase_cyano_response == ""|(grepl("Other", increase_cyano_response)), NA, increase_cyano_response)) |>
   mutate(decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured"|decrease_cyano_response == ""|(grepl("Other", decrease_cyano_response)), NA, decrease_cyano_response)) |>
  filter(!(is.na(decrease_phyto_response) & is.na(increase_phyto_response) & is.na(decrease_cyano_response) & is.na(increase_cyano_response))) %>%
  mutate(increase_decrease_mosaic = ifelse(((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)) & (!is.na(increase_phyto_response) | !is.na(increase_cyano_response))),"Both",
                                           ifelse((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)),"Decrease",
                                                  ifelse((!is.na(increase_phyto_response) | !is.na(increase_cyano_response)),"Increase", NA))))

world <- ne_countries(scale = "medium", returnclass = "sf")
#world <- map_data("world")

litdata_map <- latlong %>% 
  mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured", NA, decrease_phyto_response), 
         increase_phyto_response = ifelse(increase_phyto_response == "Not measured", NA, increase_phyto_response), 
         decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured", NA, decrease_cyano_response), 
         increase_cyano_response = ifelse(increase_cyano_response == "Not measured", NA, increase_cyano_response)) |> 
   mutate(mapvals = ifelse(((!is.na(decrease_phyto_response)|!is.na(increase_phyto_response)) & (!is.na(decrease_cyano_response)|!is.na(increase_cyano_response))), "Both",
                           ifelse(!is.na(decrease_phyto_response)|!is.na(increase_phyto_response), "Phytoplankton", 
                                 ifelse(!is.na(decrease_cyano_response)|!is.na(increase_cyano_response), "Cyanobacteria", NA)))) 

litdata_map_sf <- st_as_sf(litdata_map,
                           coords = c("long_dd", "lat_dd"),
                           crs = 4326,
                           remove = FALSE)

sitemap <- ggplot(world) +
  geom_sf(fill = "white") +
 # coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white", color = "black") +
  #geom_sf_label(stat = "sf_coordinates")+
  geom_sf(data = litdata_map_sf,
          aes(shape = as.factor(mapvals),
              color = as.factor(mapvals)),
          size = 2) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("#608687", "#4dd0e1", "darkgreen"),
                     name = "Variable") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  coord_sf() +
  theme_minimal()
sitemap

sitemap_trophic <- ggplot(world) +
  geom_sf(data = world, fill = "white") +
  #coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white") +
  #geom_sf_label(stat = "sf_coordinates")+
  # geom_sf(data = litdata_map_sf,
  #         aes(shape = as.factor(mapvals),
  #             color = as.factor(increase_decrease_mosaic))) +
    geom_sf(data = litdata_map_sf, size = 3, alpha = 0.6,
          aes(color = as.factor(increase_decrease_mosaic))) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("lightgreen", "darkgreen", "blue", "blue"),
                     name = "Was it a \nwater level \nincrease or \ndecrease? \n") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  theme_minimal() + 
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(legend.text  = element_text(size = 16),
        legend.title = element_text(size = 20))
sitemap_trophic

sitemap_years <- ggplot(world) +
  geom_sf(data = world, fill = "white") +
  #coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white") +
  #geom_sf_label(stat = "sf_coordinates")+
  # geom_sf(data = litdata_map_sf,
  #         aes(shape = as.factor(mapvals),
  #             color = as.factor(increase_decrease_mosaic))) +
    geom_sf(data = litdata_map_sf, size = 3, alpha = 0.6,
          aes(color = as.factor(years_studied))) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("lightgreen", "darkgreen", "blue", "purple"),
                     name = "How many \nyears studied?") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  theme_minimal() + 
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(legend.text  = element_text(size = 16),
        legend.title = element_text(size = 20))
sitemap_years

ggsave(paste0(here(),"/Systematic/Plots/sitemap.jpg"), plot = sitemap, height = 4, width = 6.5, units = "in")

ggsave(paste0(here(),"/Systematic/Plots/sitemapposter.jpg"), plot = sitemap_trophic, height = 4, width = 10, units = "in")
```

## Calculate proportion water level change

```{r}
#annotated code

# Create a processed dataset calculating proportional water level changes
prop_dl <- data |>
  
  # Replace empty strings in obs_maxdepth with NA
  mutate(obs_maxdepth = ifelse(obs_maxdepth == "", NA, obs_maxdepth)) |> 
  
  # Create "usedepth": use observed max depth if available,
  # otherwise rely on full pond max depth
  mutate(usedepth = ifelse(is.na(obs_maxdepth), fullpond_maxdepth, obs_maxdepth)) |> 
  
  # Convert elevation decrease and depthvalues to numeric
  mutate(elev_decrease_max = as.numeric(elev_decrease_max), 
         usedepth = as.numeric(usedepth)) |> 
  
  # Calculate proportional decrease in water level
  # Only calculate if elev_increase_max is NA (i.e., decrease-only cases)
  mutate(prop_1delta_wl_decrease =
           ifelse(is.na(elev_increase_max), elev_decrease_max/usedepth, NA)) |> 
  
  # Calculate proportional decrease relative to depth
  mutate(prop_alldelta_wl_decrease = elev_decrease_max/usedepth) |> 
  
  # Ensure decreases are negative values
  mutate(prop_alldelta_wl_decrease =
           ifelse(prop_alldelta_wl_decrease > 0,
                  prop_alldelta_wl_decrease * -1,
                  prop_alldelta_wl_decrease)) |> 
  
  # Ensure decrease values in the single-change version are also negative
  mutate(prop_1delta_wl_decrease =
           ifelse(prop_1delta_wl_decrease > 0,
                  prop_1delta_wl_decrease * -1,
                  prop_1delta_wl_decrease)) |>
  
  # Calculate proportional increase in water level
  mutate(prop_alldelta_wl_increase = elev_increase_max/usedepth) |> 
  
  # Duplicate proportional increase calculation for "1 delta" variable
  mutate(prop_1delta_wl_increase = elev_increase_max/usedepth) |> 
  
  # Create a combined proportional delta water level variable
  # Use increase if decrease is NA
  #prop_delta_wl prioritizes wl decreases!
  mutate(prop_delta_wl =
           ifelse(is.na(prop_alldelta_wl_decrease),
                  prop_alldelta_wl_increase,
                  prop_alldelta_wl_decrease)) |> 
  
  # Alternative combined variable prioritizing increases instead
  mutate(prop_delta_wl_priorityincrease =
           ifelse(is.na(prop_alldelta_wl_increase),
                  prop_alldelta_wl_decrease,
                  prop_alldelta_wl_increase))


# Prepare dataset used for boxplots and response analysis
boxplotdata2 <- prop_dl %>% 
  
  # Convert phytoplankton response to binary for water level increase
  # 1 = increase in phytoplankton
  # 0 = decrease or no significant response
  mutate(phyto_binary_increase =
           ifelse(increase_phyto_response == "No significant response", 0, 
           ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
           ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  
  # Binary phytoplankton response for water level decrease
  mutate(phyto_binary_decrease =
           ifelse(decrease_phyto_response == "No significant response", 0, 
           ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
           ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  
  # Combine phyto responses depending on direction of water level change
  mutate(combined_phyto_response_binary = case_when(
    is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
    is.na(phyto_binary_decrease) ~ phyto_binary_increase,
    TRUE ~ phyto_binary_decrease
  )) |> 
  
  # Alternative combined phyto response prioritizing increases
  mutate(combined_phyto_response_binary_increasepriority = case_when(
    prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
    is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
    TRUE ~ phyto_binary_increase
  )) |> 
  
  # Binary cyanobacteria response for water level increase
  mutate(cyano_binary_increase =
           ifelse(increase_cyano_response == "No significant response", 0, 
           ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
           ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  
  # Binary cyanobacteria response for water level decrease
  mutate(cyano_binary_decrease =
           ifelse(decrease_cyano_response == "No significant response", 0, 
           ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
           ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  
  # cyanobacteria response for water level increase
  mutate(cyano_response_wl_up =
           ifelse(increase_cyano_response == "No significant response", "Not significant", 
           ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
           ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
           ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
  
  # cyanobacteria response for water level decrease
  mutate(cyano_response_wl_down =
           ifelse(decrease_cyano_response == "No significant response", "Not significant", 
           ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
           ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
           ifelse(decrease_cyano_response == "shift",
                  "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  
  # Combined cyanobacteria binary response depending on WL direction
  mutate(combined_cyano_response_binary = case_when(
    is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
    is.na(cyano_binary_decrease) ~ cyano_binary_increase,
    TRUE ~ cyano_binary_decrease
  )) |>
  
  # Combined cyanobacteria response prioritizing increases
  mutate(combined_cyano_response_binary_increasepriority = case_when(
    prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
    prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
    TRUE ~ cyano_binary_increase
  )) |> 
  
  # Combined cyanobacteria response text
  mutate(combined_cyano_response =
           ifelse(is.na(decrease_cyano_response),
                  cyano_response_wl_up,
                  decrease_cyano_response)) %>% 
  
  # Combined cyanobacteria response prioritizing increase response text
  mutate(combined_cyano_response_increasepriority =
           ifelse(is.na(cyano_response_wl_up),
                  decrease_cyano_response,
                  cyano_response_wl_up)) %>% 
  
  # phytoplankton response for water level increase
  mutate(phyto_response_wl_up =
           ifelse(increase_phyto_response == "No significant response", "Not significant", 
           ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
           ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
           ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
  
  # phytoplankton response for water level decrease
  mutate(phyto_response_wl_down =
           ifelse(decrease_phyto_response == "No significant response", "Not significant", 
           ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
           ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
           ifelse(decrease_phyto_response == "shift",
                  "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  
  # Combined phytoplankton response text
  mutate(combined_phyto_response =
           ifelse(is.na(phyto_response_wl_down),
                  phyto_response_wl_up,
                  phyto_response_wl_down)) |> 
  
  # Combined phytoplankton response prioritizing increase direction
  mutate(combined_phyto_response_increasepriority =
           ifelse(is.na(phyto_response_wl_up),
                  phyto_response_wl_down,
                  phyto_response_wl_up))
```

```{r}
boxplotdata2 <- prop_dl %>% 
  
  # ---------------------------------------------------
  # Convert phytoplankton responses to binary variables
  # 1 = increase in phytoplankton
  # 0 = decrease OR no significant response
  # NA = response not reported
  # ---------------------------------------------------

  # Binary response when water level increases
  mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 0, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 

  # Binary response when water level decreases
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 0, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%

  # ---------------------------------------------------
  # Combine phytoplankton responses depending on
  # the direction of water level change
  # ---------------------------------------------------

  mutate(combined_phyto_response_binary = case_when(
    
    # If decrease response is missing AND the WL change was negative,
    # return NA (no usable response)
    is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
    
    # If decrease response missing but increase response exists,
    # use the increase response
    is.na(phyto_binary_decrease) ~ phyto_binary_increase,
    
    # Otherwise use the decrease response
    TRUE ~ phyto_binary_decrease
  )) |> 

  # Alternative version that prioritizes increase responses
  mutate(combined_phyto_response_binary_increasepriority = case_when(
    
    # If WL change indicates a decrease
    prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
    
    # If increase response missing but WL increased
    is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
    
    # Default: use increase response
    TRUE ~ phyto_binary_increase
  )) |> 


  # ---------------------------------------------------
  # Convert cyanobacteria responses to binary variables
  # ---------------------------------------------------

  # Binary cyanobacteria response when water level increases
  mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 

  # Binary cyanobacteria response when water level decreases
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%


  # ---------------------------------------------------
  # Create human-readable cyanobacteria responses
  # for water level increases
  # ---------------------------------------------------

  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                                ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 


  # Human-readable cyanobacteria responses
  # for water level decreases
  mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                                ifelse(decrease_cyano_response == "shift", 
                                       "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 


  # ---------------------------------------------------
  # Combine cyanobacteria responses based on WL direction
  # ---------------------------------------------------

  mutate(combined_cyano_response_binary = case_when(
    
    # If decrease response missing and WL change negative
    is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
    
    # If decrease response missing use increase response
    is.na(cyano_binary_decrease) ~ cyano_binary_increase,
    
    # Otherwise use decrease response
    TRUE ~ cyano_binary_decrease
  )) |>


  # Alternative combined response prioritizing increases
  mutate(combined_cyano_response_binary_increasepriority = case_when(
    
    prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
    
    prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
    
    TRUE ~ cyano_binary_increase
  )) |> 


  # ---------------------------------------------------
  # Combine cyanobacteria response text variables
  # ---------------------------------------------------

  mutate(combined_cyano_response =
           ifelse(is.na(decrease_cyano_response),
                  cyano_response_wl_up,
                  decrease_cyano_response)) %>% 

  mutate(combined_cyano_response_increasepriority =
           ifelse(is.na(cyano_response_wl_up),
                  decrease_cyano_response,
                  cyano_response_wl_up)) %>% 


  # ---------------------------------------------------
  # Create readable phytoplankton response text variables
  # ---------------------------------------------------

  # Phytoplankton response to WL increase
  mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                                ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 


  # Phytoplankton response to WL decrease
  mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                                ifelse(decrease_phyto_response == "shift", 
                                       "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 


  # ---------------------------------------------------
  # Combine phytoplankton response text variables
  # ---------------------------------------------------

  mutate(combined_phyto_response =
           ifelse(is.na(phyto_response_wl_down),
                  phyto_response_wl_up,
                  phyto_response_wl_down)) |> 

  mutate(combined_phyto_response_increasepriority =
           ifelse(is.na(phyto_response_wl_up),
                  phyto_response_wl_down,
                  phyto_response_wl_up))
```

```{r max depth and proportion water level}
#annotated version above
# prop_dl <- data |>
#   mutate(obs_maxdepth = ifelse(obs_maxdepth == "", NA, obs_maxdepth)) |> 
#   mutate(usedepth = ifelse(is.na(obs_maxdepth), fullpond_maxdepth, obs_maxdepth)) |> 
#   #select(c("obs_maxdepth", "fullpond_maxdepth", "usedepth")) |> 
#   mutate(elev_decrease_max = as.numeric(elev_decrease_max)) |> 
#   mutate(usedepth = as.numeric(usedepth)) |> 
#   mutate(prop_1delta_wl_decrease = ifelse(is.na(elev_increase_max), elev_decrease_max/usedepth, NA)) |> 
#   mutate(prop_alldelta_wl_decrease = elev_decrease_max/usedepth) |> 
#   mutate(prop_alldelta_wl_decrease = ifelse(prop_alldelta_wl_decrease > 0, prop_alldelta_wl_decrease* -1, prop_alldelta_wl_decrease)) |> 
#   mutate(prop_1delta_wl_decrease = ifelse(prop_1delta_wl_decrease > 0, prop_1delta_wl_decrease* -1, prop_1delta_wl_decrease)) |>
#   mutate(prop_alldelta_wl_increase = elev_increase_max/usedepth) |> 
#   mutate(prop_1delta_wl_increase = elev_increase_max/usedepth) |> 
#   #select(c("obs_maxdepth", "fullpond_maxdepth", "usedepth", "prop_1delta_wl_decrease", "prop_alldelta_wl_decrease", "elev_decrease_max", "elev_increase_max", "prop_alldelta_wl_increase", "prop_1delta_wl_increase"))
#   mutate(prop_delta_wl = ifelse(is.na(prop_alldelta_wl_decrease), prop_alldelta_wl_increase, prop_alldelta_wl_decrease)) |> 
#   mutate(prop_delta_wl_priorityincrease = ifelse(is.na(prop_alldelta_wl_increase), prop_alldelta_wl_decrease, prop_alldelta_wl_increase))
# 
# boxplotdata2 <- prop_dl %>% 
#   #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
#     mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 0, 
#                                 ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
#                                 ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
#   mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 0, 
#                                 ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
#                                 ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
#   # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
#   #          is.na() %>% 
#   #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
#     mutate(combined_phyto_response_binary = case_when(
#   is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
#   is.na(phyto_binary_decrease) ~ phyto_binary_increase,
#   TRUE ~ phyto_binary_decrease
# )) |> 
#     mutate(combined_phyto_response_binary_increasepriority = case_when(
#  prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
#   is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
#   TRUE ~ phyto_binary_increase
# )) |> 
#     #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
#     mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
#                                 ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
#                                 ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
#   mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
#                                 ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
#                                 ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
#   # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
#   #          is.na() %>% 
#   #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
#     #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
#   mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
#                                 ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
#                                 ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
#                                ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
#     mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
#                                 ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
#                                 ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
#                                ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
#   mutate(combined_cyano_response_binary = case_when(
#   is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
#   is.na(cyano_binary_decrease) ~ cyano_binary_increase,
#   TRUE ~ cyano_binary_decrease
# )) |>
#        mutate(combined_cyano_response_binary_increasepriority = case_when(
#   prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
#   prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
#   TRUE ~ cyano_binary_increase
# )) |> 
#    mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
#   mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
#     mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
#                                 ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
#                                 ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
#                                ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
#     mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
#                                 ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
#                                 ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
#                                ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
#   mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
#   mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))

```

```{r}
# boxplotdata2 <- prop_dl %>% 
#   #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
#   #phytoplankton response to water level increases
#     mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 0, 
#                                 ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
#                                 ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
#   mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 0, 
#                                 ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
#                                 ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
#   # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
#   #          is.na() %>% 
#   #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
#     mutate(combined_phyto_response_binary = case_when(
#   is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
#   is.na(phyto_binary_decrease) ~ phyto_binary_increase,
#   TRUE ~ phyto_binary_decrease
# )) |> 
#     mutate(combined_phyto_response_binary_increasepriority = case_when(
#  prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
#   is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
#   TRUE ~ phyto_binary_increase
# )) |> 
#     #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
#     mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
#                                 ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
#                                 ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
#   mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
#                                 ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
#                                 ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
#   # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
#   #          is.na() %>% 
#   #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
#     #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
#   mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
#                                 ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
#                                 ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
#                                ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
#     mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
#                                 ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
#                                 ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
#                                ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
#   mutate(combined_cyano_response_binary = case_when(
#   is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
#   is.na(cyano_binary_decrease) ~ cyano_binary_increase,
#   TRUE ~ cyano_binary_decrease
# )) |>
#        mutate(combined_cyano_response_binary_increasepriority = case_when(
#   prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
#   prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
#   TRUE ~ cyano_binary_increase
# )) |> 
#    mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
#   mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
#     mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
#                                 ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
#                                 ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
#                                ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
#     mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
#                                 ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
#                                 ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
#                                ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
#   mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
#   mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))
```

## Set up logistic regression for effect of water level change on phytoplankton and cyanobacteria communities

We ran the logistic regression three ways. In the manuscript, the logistic regression includes both water level increases and decreases. We also prioritized only water level increases (any lake that had both a water level increase and decrease was kept as only an increase) and subsequently prioritized only water level decreases (any lake that had both a water level increase and decrease was kept as only a decrease)

```{r}
phytoincrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced) |> 
  #mutate(phyto_binary_decrease = NA) |> 
  filter(!is.na(phyto_response_wl_up)) |> 
  filter(!is.na(prop_1delta_wl_increase))

phytodecrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced) |> 
  filter(!is.na(phyto_response_wl_down)) |> 
  filter(!is.na(prop_1delta_wl_decrease))

phyto_logreg <- rbind(phytodecrease_logreg, phytoincrease_logreg)

phytos <- phyto_logreg |> 
  mutate(response = as.numeric(ifelse(!is.na(prop_1delta_wl_decrease), phyto_binary_decrease, 
                                 ifelse(!is.na(prop_1delta_wl_increase), phyto_binary_increase, NA))), 
         deltawl = as.numeric(ifelse(!is.na(prop_1delta_wl_decrease), prop_1delta_wl_decrease, 
                                 ifelse(!is.na(prop_1delta_wl_increase), prop_1delta_wl_increase, NA)))) |> 
  dplyr::select(response, deltawl, name_yr, trophic_status_reduced) |> 
  mutate(type = "phyto")


ggplot(phytos, aes(x = deltawl, y = response)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.4) +
  stat_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = TRUE) + labs(
    x = "Proportional Water Level Change",
    y = "Probability of Phytoplankton Increase") +
  theme_classic()

cyanoincrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced) |> 
  mutate(cyano_binary_decrease = NA) |> 
  filter(!is.na(cyano_response_wl_up)) |> 
  filter(!is.na(prop_1delta_wl_increase))

cyanodecrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced) |> 
  filter(!is.na(cyano_response_wl_down)) |> 
  filter(!is.na(prop_1delta_wl_decrease))

cyano_logreg <- rbind(cyanodecrease_logreg, cyanoincrease_logreg)

cyanos <- cyano_logreg |> 
  mutate(
    response = as.numeric(
      ifelse(!is.na(prop_1delta_wl_decrease), cyano_binary_decrease, 
      ifelse(!is.na(prop_1delta_wl_increase), cyano_binary_increase, NA))
    ), 
    deltawl = as.numeric(
      ifelse(!is.na(prop_1delta_wl_decrease), prop_1delta_wl_decrease, 
      ifelse(!is.na(prop_1delta_wl_increase), prop_1delta_wl_increase, NA))
    )
  ) |>
  mutate(type = "cyano") |> 
  dplyr::select(response, deltawl, name_yr, trophic_status_reduced, type)


ggplot(cyanos, aes(x = deltawl, y = response)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.4) +
  stat_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = TRUE) +
  labs(
    x = "Proportional Water Level Change",
    y = "Probability of Cyanobacteria Increase"
  ) +
  theme_classic()

all_logreg <- rbind(phytos, cyanos)


ggplot(all_logreg, aes(x = deltawl, y = response)) +
  geom_point(aes(color = trophic_status_reduced), alpha = 0.6, size = 4) +
  geom_smooth(
    method = "glm",
    method.args = list(family = "binomial"),
    se = FALSE,
    color = "black"
  ) +
  facet_wrap(~type) +
  theme_bw() +
  labs(
    x = "Proportional Water Level Change",
    y = "Probability of Increase"
  )

logregform_phyto <-  glm(formula=response ~ deltawl, data=phytos, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response ~ deltawl, data=cyanos, family=binomial)
summary(logregform_cyano)

```

## Plot logistic regression with combined water level increase/decrease

This chunk will generate logistic regression plots for both phytoplankton and cyanobacteria.

```{r broken}
#| eval: false
#| include: false
logregdata_increase1 <- logregdata_increase |>
  rename(prop_delta_wl = prop_delta_wl_priorityincrease)

new_all <- bind_rows(logregdata_decrease, logregdata_increase1) |>
  select(-c(wl_priority)) |>
  unique()

all <- new_all %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) +
  #geom_jitter()+
 ylab("Response to changes in water level") +
  xlab("Proportion water level change") +
  ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10),
        axis.title.y=element_text(angle=90, vjust = .5, size = 12),
        legend.title=element_blank())
all
#117733
#factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic"))
logreg_phyto <- new_all %>% 
  filter(type == "phyto") |> 
  # ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  ggplot(aes(x = prop_delta_wl, y = response_binary, shape = as.factor(type))) +
  scale_color_manual(values = c("#3355DB", "#FFC51E", "#B1C3A2", "#549C79", "gray"), name = "")+
  scale_shape_manual(values = c(20, 10, 11, 12, 13)) +
  # geom_point(size = 10, alpha = 0.5, aes(color = as.factor(trophic_status_reduced))) +
    geom_point(
    aes(
      color = as.factor(factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic", "not reported"))),
      shape = as.factor(type),
      text = paste(
        "Name / Year:", name_yr,
        "<br>Trophic status:", trophic_status_reduced,
        "<br>Δ WL:", prop_delta_wl,
        "<br>Response:", response_binary
      )
    ),
    size = 14,
    alpha = 0.75
  ) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 6, color = "black") +
  #geom_jitter()+
 # ylab("Phytoplankton response to changes in water level") +
 #  xlab("Proportion water level change") +
   ylab("") +
  xlab("") +
  #ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 0.5, size = 20),
        axis.text.y = element_text(hjust = 1, size = 20),
        axis.title.y=element_text(angle=90, hjust = 1, size = 28),
        axis.title.x=element_text(vjust = 1, size = 28),
        legend.text  = element_text(size = 28),
        legend.title=element_blank())

plotly_p <- ggplotly(logreg_phyto)

plotly_p # I have this as a plotly to be able to see exactly what's going into this for now. Useful, but can probably be removed for publishing
logreg_phyto

logregdata_phyto <- new_all %>% subset(type == "phyto")
logregdata_cyano <- new_all %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)

ggsave(paste0(here(),"/Systematic/Plots/logregposter.jpg"), plot = logreg_phyto, height = 9.5, width = 20, units = "in")
ggsave(paste0(here(),"/Systematic/Plots/logreg.jpg"), plot = sitemap, height = 4, width = 8, units = "in")

```

## Ordinal Logistic Regression

```{r}
boxplotdat_ordinal <- prop_dl %>% 
  #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
    mutate(phyto_ord_increase = ifelse(increase_phyto_response == "No significant response", 0, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", -1, 
                               ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_ord_decrease = ifelse(decrease_phyto_response == "No significant response", 0, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", -1, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    mutate(combined_phyto_response_ord = case_when(
  is.na(phyto_ord_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(phyto_ord_decrease) ~ phyto_ord_increase,
  TRUE ~ phyto_ord_decrease
)) |> 
    mutate(combined_phyto_response_ord_increasepriority = case_when(
 prop_delta_wl_priorityincrease < 0 ~ phyto_ord_decrease,
  is.na(phyto_ord_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_ord_increase,
  TRUE ~ phyto_ord_increase
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_ord_increase = ifelse(increase_cyano_response == "No significant response", 0, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_ord_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response_ord = case_when(
  is.na(cyano_ord_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(cyano_ord_decrease) ~ cyano_ord_increase,
  TRUE ~ cyano_ord_decrease
)) |>
       mutate(combined_cyano_response_ord_increasepriority = case_when(
  prop_delta_wl_priorityincrease < 0 ~ cyano_ord_decrease,
  prop_delta_wl_priorityincrease > 0 ~ cyano_ord_increase,
  TRUE ~ cyano_ord_increase
)) |> 
   mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
  mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
  mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))

require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)

phytoincreasedata <- boxplotdat_ordinal |>  
  dplyr::select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_ord, name_yr, res_name) |> 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_ord)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_ord = combined_phyto_response_ord, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto", 
         wlchange = "increase")

phytodecreasedata <- boxplotdat_ordinal %>% 
    dplyr::select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_ord, name_yr, res_name) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_ord)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_ord = combined_phyto_response_ord, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto", 
         wlchange = "decrease")


  
cyanodata <- boxplotdat_ordinal %>% 
  dplyr::select(prop_delta_wl, combined_cyano_response, combined_cyano_response_ord, name_yr, trophic_status_reduced, res_name) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_ord)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_ord = combined_cyano_response_ord, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdat_ordinal %>% 
  dplyr::select(prop_delta_wl, combined_phyto_response, combined_phyto_response_ord, name_yr, trophic_status_reduced, res_name) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_ord)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_ord, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
#logregdata_decrease <- rbind(cyanodata, phytodata)
#logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"
logreg_phyto_ord <- rbind(phytoincreasedata, phytodecreasedata)

logreg_phyto_ord <- logreg_phyto_ord |> 
  mutate(prop_delta_wl = ifelse(wlchange == "increase", prop_alldelta_wl_increase, 
                                ifelse(wlchange == "decrease", prop_alldelta_wl_decrease, NA))) |> dplyr::select(-c(prop_alldelta_wl_decrease, prop_alldelta_wl_increase))

logreg_phyto_ord$response_ord <- factor(
logreg_phyto_ord$response_ord,
  levels = c(-1, 0, 1),
  labels = c("Decrease", "Not significant", "Increase"),
  ordered = TRUE
)

library(MASS)

model <- polr(
response_ord ~ prop_delta_wl,
  data = logreg_phyto_ord,
  Hess = TRUE
)

summary(model)

ctable <- coef(summary(model))

p_values <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

ctable <- cbind(ctable, "p value" = p_values)

ctable

ggplot(logreg_phyto_ord, 
       aes(x = prop_delta_wl, 
           y = response_ord)) +
  geom_jitter(height = 0.1) +
  theme_minimal()

logreg_phyto_ord$prop_delta_wl2 <- logreg_phyto_ord$prop_delta_wl^2
library(MASS)

model_quad <- polr(response_ord ~ prop_delta_wl + prop_delta_wl2,
                   data = logreg_phyto_ord,
                   Hess = TRUE)

summary(model_quad)

ctable <- coef(summary(model_quad))

p_values <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

anova(model, model_quad)

ctable <- cbind(ctable, "p value" = p_values)

ctable

#install.packages("ggeffects")
library(ggeffects)

pred_quad <- ggpredict(model_quad,
                       terms = "prop_delta_wl [all]")

plot(pred_quad)


help <- boxplotdat_ordinal |> dplyr::select(phyto_ord_decrease, phyto_ord_increase, phyto_response_wl_up, phyto_response_wl_down, prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl)


```

## 

## Set up and run Chi-squared tests

These Chi-square tests assess the effect of increases and decreases in water level on phytoplankton and cyanobacteria (do they experience an increase, a decrease, or no change?).

```{r chisq tests of independence}
# Create contingency table
tab_cyano_all <- boxplotdata2 %>%
  pivot_longer(
    cols = c(cyano_response_wl_up, cyano_response_wl_down),
    names_to = "water_direction",
    values_to = "response"
  ) %>%
  mutate(response = gsub(" ", "", response)) %>%
  filter(!is.na(response)) %>%
  with(table(water_direction, response))

# Run chi-square test
# Are there differences in cyanobacterial responses to increases vs. decreases in water level?
chisq_test(tab_cyano_all)
f_chisq_test(tab_cyano_all) #Bonferroni - adj. p

tab_cyano_down <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(cyano_response_wl_down) ~ "wl_down",
      is.na(cyano_response_wl_down) & !is.na(cyano_response_wl_up) ~ "wl_up",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(cyano_response_wl_down) ~ cyano_response_wl_down,
      is.na(cyano_response_wl_down) & !is.na(cyano_response_wl_up) ~ cyano_response_wl_up,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)

tab_cyano_down

tab_cyano_down <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_cyano_down)

chisq_test(tab_cyano_down)
f_chisq_test(tab_cyano_down)
stats::fisher.test(tab_cyano_down)


tab_cyano_up <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(cyano_response_wl_up) ~ "wl_up",                     # priority
      is.na(cyano_response_wl_up) & !is.na(cyano_response_wl_down) ~ "wl_down",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(cyano_response_wl_up) ~ cyano_response_wl_up,        # priority
      is.na(cyano_response_wl_up) & !is.na(cyano_response_wl_down) ~ cyano_response_wl_down,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)


tab_cyano_up <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_cyano_up)
chi_cyano_up <- stats::chisq.test(tab_cyano_up)
f_chisq_test(tab_cyano_up)


# Create contingency table
tab_phyto_all <- boxplotdata2 %>%
  pivot_longer(
    cols = c(phyto_response_wl_up, phyto_response_wl_down),
    names_to = "water_direction",
    values_to = "response"
  ) %>%
  mutate(response = gsub(" ", "", response)) %>%
  filter(!is.na(response)) %>%
  with(table(water_direction, response))

# Run chi-square test
# Are there differences in cyanobacterial responses to increases vs. decreases in water level?
chisq_test(tab_phyto_all)
f_chisq_test(tab_phyto_all) #Bonferroni - adj. p

#wl decrease priority
tab_phyto <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(phyto_response_wl_down) ~ "wl_down",
      is.na(phyto_response_wl_down) & !is.na(phyto_response_wl_up) ~ "wl_up",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(phyto_response_wl_down) ~ phyto_response_wl_down,
      is.na(phyto_response_wl_down) & !is.na(phyto_response_wl_up) ~ phyto_response_wl_up,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)
tab_phyto
tab_phyto_decrease <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_phyto)
tab_phyto_decrease

stats::chisq.test(tab_phyto_decrease)
f_chisq_test(tab_phyto_decrease)

tab_phyto_increase <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(phyto_response_wl_up) ~ "wl_up",                        # priority to wl_up
      is.na(phyto_response_wl_up) & !is.na(phyto_response_wl_down) ~ "wl_down",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(phyto_response_wl_up) ~ phyto_response_wl_up,           # priority to wl_up
      is.na(phyto_response_wl_up) & !is.na(phyto_response_wl_down) ~ phyto_response_wl_down,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)
tab_phyto_increase
phyto_increase_chi <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_phyto_increase)
f_chisq_test(phyto_increase_chi)

stats::fisher.test(tab_cyano_down)
stats::fisher.test(tab_cyano_up)
stats::fisher.test(tab_phyto_decrease)
stats::fisher.test(phyto_increase_chi)
```

## Chi-squared test plots

```{r barplots for chisq}
library(dplyr)
cyano_chi <- ggplot(as.data.frame(tab_cyano_all), aes(x = water_direction, y = Freq, fill = response)) +
  geom_bar(stat = "identity", position = "fill") +
     geom_text(aes(label = Freq),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 6) +
  labs(x = "",
       y = "Proportion",
       fill = "Response") + #,
      # title = "Cyanobacteria Response by Water Level Direction") +
  scale_x_discrete(labels = c(
  "cyano_response_wl_up" = "Water level increase",
  "cyano_response_wl_down" = "Water level decrease"))+
 scale_fill_manual(
  values = c("#b2ebf2", "#4dd0e1", "#e0e0e0"),   # 3 colors
  name = "Did \ncyanobacteria\nincrease or \ndecrease?",
  labels = c("Decrease", "Increase", "No Change")) +
  theme_bw() + 
    theme(axis.text.x = element_text(hjust = 0.5, size = 12),
        #axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=90, vjust = 0.5, size = 12),
        axis.title.x=element_text(vjust = 1, size = 12),
        #legend.text  = element_blank(),
        legend.title=element_text(size = 12), 
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm')) 

df_plot_phyto <- as.data.frame(tab_phyto_all)

phyto_chi <- ggplot(df_plot_phyto, aes(x = water_direction, y = Freq, fill = response)) +
  geom_bar(stat = "identity", position = "fill") +
     geom_text(aes(label = Freq),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 6) +
  labs(x = "",
       y = "Proportion",
       fill = "Response") +#,
      # title = "Phytoplankton Response by Water Level Direction") +
    scale_x_discrete(labels = c(
  "phyto_response_wl_up" = "Water level increase",
  "phyto_response_wl_down" = "Water level decrease"))+
  scale_fill_manual(values = c("#ddefcd", "#99bbab", "#e0e0e0"), 
                    name = "Did \nphytoplankton\nincrease or \ndecrease?",
  labels = c("Decrease", "Increase", "No Change")) +
  theme_bw() + 
    theme(axis.text.x = element_text(hjust = 0.5, size = 12),
        #axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=90, vjust = 0.5, size = 12),
        axis.title.x=element_text(vjust = 1, size = 12),
        #legend.text  = element_blank(),
        legend.title=element_text(size = 12), 
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm')) 

chi <- phyto_chi / cyano_chi + #+ plot_annotation(title = "Chi-Square") +
  plot_layout(axes = "collect")


chi

ggsave(paste0(here(), "/Systematic/Plots/chi.jpg"), plot = chi, height = 5, width = 6.5, units = "in")
```

# Supplemental figures

## Trophic Levels

```{r trophic levels}
library(dplyr)
#install.packages("conflicted")
library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
#consider subsetting these to phyto and cyano specific studies?
non_dups <- data |> select(res_name, trophic_status_reduced) |> 
  unique()

#plot only reduced trophic state levels and only reservoirs that are not duplicated
#from 81 -> 72
trophic <- non_dups |> 
  filter(trophic_status_reduced != "not reported") |> 
  ggplot(aes(x= factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic")))) + 
  geom_bar(fill = "#608687") +
  xlab("Trophic Status") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

trophic

ggsave(paste0(here(),"/Systematic/Plots/trophic.jpg"), plot = trophic, height = 3, width = 6.5, units = "in")

```

## Logistic regression prioritizing increases in water level

Logistic regression figures that prioritize either increases or decreases in water level, rather than both as plotted above.

```{r}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary_increasepriority, 
         response = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "increase")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary_increasepriority, 
         response = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")|> 
  mutate(wl_priority = "increase")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_increase <- rbind(cyanodata, phytodata)
#"#7A524B","#3C7F29","#8DBAE9"

logreg_wlincrease <- logregdata_increase %>% ggplot(aes(x = prop_delta_wl_priorityincrease, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes \nin water level") + 
  xlab("Proportion water level change") + 
  ggtitle("Priority WL increases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
logreg_wlincrease

logregdata_phyto <- logregdata_increase %>% subset(type == "phyto")
logregdata_cyano <- logregdata_increase %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
ggsave(paste0(here(),"/Systematic/Plots/logreg_wlincrease.jpg"), plot = logreg_wlincrease, height = 3, width = 6.5, units = "in")
```

## Logistic regression prioritizing decreases in water level

```{r wl decreases}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_decrease <- rbind(cyanodata, phytodata)
logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"

logreg_wldecrease <- logregdata %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes \nin water level") + 
  xlab("Proportion water level change") + 
  ggtitle("Priority WL decrease")+
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
logreg_wldecrease

logregdata_phyto <- logregdata %>% subset(type == "phyto")
logregdata_cyano <- logregdata %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
both_logreg <- logreg_wldecrease/logreg_wlincrease
ggsave(paste0(here(),"/Systematic/Plots/logreg_wldecrease.jpg"), plot = logreg_wlincrease, height = 3, width = 6.5, units = "in")
ggsave(paste0(here(),"/Systematic/Plots/bothlogreg.jpg"), plot = both_logreg, height = 6, width = 6.5, units = "in")
```

## Barplot: number of reservoir by continent

```{r possible supplement figures}

#countries by continent
countries_sf <- data |> 
  select(lat_dd, long_dd, res_name, cov_num, name_yr) |> 
  filter(!is.na(lat_dd) & !is.na(long_dd)) |> 
  st_as_sf(coords = c("long_dd", "lat_dd"), crs = 4326) 
world <- ne_countries(scale = "medium", returnclass = "sf")
df_with_country <- st_join(countries_sf, world["name_long"])
df_final <- df_with_country %>% 
  st_drop_geometry() %>%
  rename(country = name_long)
df_final <- df_final %>% 
  mutate(continent = world$continent[match(country, world$name_long)])

countries <- left_join(data, df_final, by = c("res_name", "cov_num", "name_yr")) |> select(c("res_name", "name_yr", "country.x", "country.y", "continent")) |> 
  mutate(country.x = ifelse(country.x == "", NA, country.x)) |> 
  mutate(country = ifelse(!is.na(country.x), country.x, country.y)) |> 
  mutate(continent.a = case_when(
    country %in% c("Brazil", "Argentina", "Uruguay", "Argentina and Uruguay") ~ "South America",
      country %in% c("United States", "USA/Canada", "Canada", "Mexico") ~ "North America",
      country %in% c("Portugal", "England", "Italy", "France", "Serbia", 
                     "Estonia", "Estonia/Russia") ~ "Europe",
      country %in% c("Russia") ~ "Europe/Asia",
      country %in% c("China", "Taiwan", "Malaysia", "Mongolia", "Lebanon",
                     "Cambodia", "Vietnam") ~ "Asia",
       country %in% c("Australia") ~ "Oceania",
    country %in% c("South Africa") ~ "Africa",
      TRUE ~ NA_character_)) |> 
  mutate(continent = ifelse(!is.na(continent.a), continent.a, continent)) |> 
  select(res_name, country, continent) |> 
  unique()

countries |> 
  filter(!is.na(continent)) |> 
  ggplot(aes(x=continent))+
  geom_bar() + 
  xlab("Continent") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")
```

## Histogram of maximum reservoir depths

```{r}
#histogram of maxdepths

maxdepth <- data |> 
  subset(!is.na(fullpond_maxdepth)) %>% 
  ggplot(aes(x=fullpond_maxdepth)) + 
  geom_histogram() +
  xlab("Max depth at Full Pond") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave(paste0(here(),"/Systematic/Plots/maxdepth.jpg"), plot = maxdepth, height = 3, width = 6.5, units = "in")
```

## Drivers for phyto + cyano combined

drivers and look at variables that changed with WL change

```{r}
drivers <- data |> 
  mutate(tp = ifelse(grepl("TP", chem_drive), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_drive), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_drive), 1, 0), 
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         do = ifelse(grepl("DO", chem_drive), 1, 0),
         srp = ifelse(grepl("SRP", chem_drive), 1, 0),
         no3 = ifelse(grepl("NO3", chem_drive), 1, 0),
         cond = ifelse(grepl("conductivity", chem_drive), 1, 0),
         metals = ifelse(grepl("metals", chem_drive), 1, 0),
         orp = ifelse(grepl("ORP", chem_drive), 1, 0),
         no2 = ifelse(grepl("NO2", chem_drive), 1, 0),
         dic = ifelse(grepl("DIC", chem_drive), 1, 0),
       dom = ifelse(grepl("DOM", chem_drive), 1, 0),
       pom = ifelse(grepl("POM", chem_drive), 1, 0),
       tkn = ifelse(grepl("TKN", chem_drive), 1, 0),
         doc = ifelse(grepl("DOC", chem_drive), 1, 0),
         din = ifelse(grepl("DIN", chem_drive), 1, 0),
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         salinity = ifelse(grepl("salinity", chem_drive), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_drive), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_drive), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_drive), 1, 0),
         transparency = ifelse(grepl("transparency", phys_drive), 1, 0),
         turb = ifelse(grepl("turbidity", phys_drive), 1, 0),
         light = ifelse(grepl("PAR", phys_drive), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid", phys_drive), 1, 0), 
         strat = ifelse(grepl("mix|stab|strat", phys_drive), 1, 0),
         restime = ifelse(grepl("residence", phys_drive), 1, 0), 
         flush_discharge = ifelse(grepl("flush|discharge", phys_drive), 1, 0), 
         inflow = ifelse(grepl("inflow", phys_drive), 1, 0),
         precip = ifelse(grepl("precip|rain", phys_drive), 1, 0),
         air_temp = ifelse(grepl("air", phys_drive), 1, 0),
         other_phys = ifelse(grepl("Other", phys_drive), 1, 0))

other_phys <- drivers |> 
  filter(other_phys == 1) |> 
  select(name_yr, title, num_water, phys_drive)

  unique(data$phys_response)
  
  #WLchangevars is whether or not vars changed under WL change
WLchangevars <- data |> 
mutate(not_assessed = ifelse(grepl("not assessed", phys_response), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_response), 1, 0),
         transparency = ifelse(grepl("transparency", phys_response), 1, 0),
         turb = ifelse(grepl("turbidity", phys_response), 1, 0),
         light = ifelse(grepl("PAR", phys_response), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_response), 1, 0), 
         #strat = ifelse(grepl("mix|stab|strat", phys_response), 1, 0),
         #restime = ifelse(grepl("residence", phys_response), 1, 0), 
         #flush_discharge = ifelse(grepl("flush|discharge", phys_response), 1, 0), 
         #inflow = ifelse(grepl("inflow", phys_response), 1, 0),
         #precip = ifelse(grepl("precip|rain", phys_response), 1, 0),
         #air_temp = ifelse(grepl("air", phys_response), 1, 0),
         other_phys = ifelse(grepl("Other", phys_response), 1, 0), 
       watercolor = ifelse(grepl("water color", phys_response), 1, 0)) |> 
mutate(tp = ifelse(grepl("TP", chem_response), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_response), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_response), 1, 0), 
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         do = ifelse(grepl("DO", chem_response), 1, 0),
         srp = ifelse(grepl("SRP", chem_response), 1, 0),
         no3 = ifelse(grepl("NO3", chem_response), 1, 0),
         cond = ifelse(grepl("conductivity", chem_response), 1, 0),
         metals = ifelse(grepl("metals", chem_response), 1, 0),
         orp = ifelse(grepl("ORP", chem_response), 1, 0),
         no2 = ifelse(grepl("NO2", chem_response), 1, 0),
         doc = ifelse(grepl("DOC", chem_response), 1, 0),
       dic = ifelse(grepl("DIC", chem_response), 1, 0),
       dom = ifelse(grepl("DOM", chem_response), 1, 0),
       pom = ifelse(grepl("POM", chem_response), 1, 0),
       tkn = ifelse(grepl("TKN", chem_response), 1, 0),
         din = ifelse(grepl("DIN", chem_response), 1, 0),
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         salinity = ifelse(grepl("salinity", chem_response), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_response), 1, 0)) 

#spreadvars is whether or not the var was measured
spreadvars <- data |> 
  mutate(filt_chla = ifelse(grepl("filtered chlorophyll", bio_vars), 1, 0),                       #biotic variables
         phyto_micros = ifelse(grepl("phytoplankton microscopy", bio_vars), 1, 0), 
         single_cyano_taxon = ifelse(grepl("single cyanobacterial taxon", bio_vars), 1, 0), 
         sensor_chla = ifelse(grepl("sensor", bio_vars), 1, 0),
         other_bio = ifelse(grepl("Other", bio_vars), 1, 0)) |> 
  mutate(tp = ifelse(grepl("TP", chem_vars), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_vars), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_vars), 1, 0), 
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         do = ifelse(grepl("DO", chem_vars), 1, 0),
         srp = ifelse(grepl("SRP", chem_vars), 1, 0),
         no3 = ifelse(grepl("NO3", chem_vars), 1, 0),
         cond = ifelse(grepl("conductivity", chem_vars), 1, 0),
         metals = ifelse(grepl("metals", chem_vars), 1, 0),
         orp = ifelse(grepl("ORP", chem_vars), 1, 0),
         no2 = ifelse(grepl("NO2", chem_vars), 1, 0),
         doc = ifelse(grepl("DOC", chem_vars), 1, 0),
          dic = ifelse(grepl("DIC", chem_vars), 1, 0),
       dom = ifelse(grepl("DOM", chem_vars), 1, 0),
       pom = ifelse(grepl("POM", chem_vars), 1, 0),
       tkn = ifelse(grepl("TKN", chem_vars), 1, 0),
         din = ifelse(grepl("DIN", chem_vars), 1, 0),
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         salinity = ifelse(grepl("salinity", chem_vars), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_vars), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_vars), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_vars), 1, 0),
         transparency = ifelse(grepl("transparency", phys_vars), 1, 0),
         turb = ifelse(grepl("turbidity", phys_vars), 1, 0),
         light = ifelse(grepl("PAR", phys_vars), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing #these are commented out below
        tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_vars), 1, 0), 
         #strat = ifelse(grepl("mix|stab|strat", phys_vars), 1, 0),
        # restime = ifelse(grepl("residence", phys_vars), 1, 0), 
        # flush_discharge = ifelse(grepl("flush|discharge", phys_vars), 1, 0), 
        # inflow = ifelse(grepl("inflow", phys_vars), 1, 0),
        # precip = ifelse(grepl("precip|rain", phys_vars), 1, 0),
        # air_temp = ifelse(grepl("air", phys_vars), 1, 0),
         other_phys = ifelse(grepl("Other", phys_vars), 1, 0)) 
        # watercolor = ifelse(grepl("water color", phys_vars), 1, 0))

# unique(data$chem_drive)
# unique(data$phys_drive)

#drivers
chemvars_drive <- drivers |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
freqchem_drive <- as.data.frame(colSums(chemvars_drive))
chemvars_drive <- setNames(cbind(rownames(freqchem_drive), freqchem_drive, row.names = NULL), 
         c("names", "sum"))

physvars_drive <- drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))
freqphys_drive <- as.data.frame(colSums(physvars_drive))
physvars_drive <- setNames(cbind(rownames(freqphys_drive), freqphys_drive, row.names = NULL), 
         c("names", "sum"))

#all measured
chemvars <- spreadvars |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
freqchem <- as.data.frame(colSums(chemvars))
chemvars <- setNames(cbind(rownames(freqchem), freqchem, row.names = NULL), 
         c("names", "measured"))

physvars <- spreadvars |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))
freqphys <- as.data.frame(colSums(physvars))
physvars <- setNames(cbind(rownames(freqphys), freqphys, row.names = NULL), 
         c("names", "measured"))

physvars_noWL <- spreadvars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys))
freqphys_noWL<- as.data.frame(colSums(physvars_noWL))
physvars_noWL <- setNames(cbind(rownames(freqphys_noWL), freqphys_noWL, row.names = NULL), 
         c("names", "measured"))

#changes with WL change
chemvarsWL <- WLchangevars |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
freqchemWL <- as.data.frame(colSums(chemvarsWL))
chemvarsWL <- setNames(cbind(rownames(freqchemWL), freqchemWL, row.names = NULL), 
         c("names", "sum"))

physvarsWL <- WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys))
freqphysWL <- as.data.frame(colSums(physvarsWL))
physvarsWL <- setNames(cbind(rownames(freqphysWL), freqphysWL, row.names = NULL), 
         c("names", "sum"))

#physvarsWL_all <- WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp)) # this is all possible variables, our analysis was limited to only the ones we specifically extracted

chemvars_drive_comb <-  right_join(chemvars, chemvars_drive)
physvars_drive_comb <- right_join(physvars, physvars_drive)

chemvarsWL_comb <- right_join(chemvarsWL, chemvars, by = "names")
physvarsWL_comb <- right_join(physvars_noWL, physvarsWL, by = "names")

#Make supplemental Figs
#water level change
chemWL <- chemvarsWL_comb |> 
  filter(!names == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  #ggtitle("Chemical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#8680cc", "#cfcdef"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as changing \nwith water level?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

physWL <- physvarsWL_comb |> 
  filter(!names == "other_phys") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#8680cc", "#cfcdef"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as changing \nwith water level?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))


chemWL <- chemvarsWL_comb |> 
  filter(!names == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#8680cc", "#cfcdef"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as changing \nwith water level?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

#overall response
physdrive <- physvars_drive_comb |> 
 filter(!names == "other_phys") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(measured = if_else(measured == 42, 61, measured),
         notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("Physical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#CC6060", "#F2DADA"), 
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as driving \nresponse?") +
  theme_bw() + 
  labs(y = "") +
    theme(#legend.position = "none",
  axis.text.x  = element_text(hjust = .5, size = 6),
  axis.title.x = element_blank(),
  axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

chemdrive <- chemvars_drive_comb |> 
  filter(!names == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("Chemical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#CC6060", "#F2DADA"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as driving \nresponse?") +
  theme_bw() + 
  labs(y = "") +
  theme(#legend.position = "none",
  axis.text.x  = element_text(hjust = .5, size = 6),
  axis.title.x = element_blank(),
  axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

combined <- ((chemdrive/physdrive) | (chemWL / physWL)) + plot_layout(guides = "collect") &
  theme(
    legend.position = "right",
    legend.box = "vertical"
  )
combined

ggsave(paste0(here(), "/Systematic/Plots/drivers_all_WL.jpg"), plot = combined, height = 4, width = 8, units = "in")
```

## Reverse LogReg

```{r}

boxplotdata2_opposite <- prop_dl %>% 
  #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
    mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 1, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 1, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    mutate(combined_phyto_response_binary = case_when(
  is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(phyto_binary_decrease) ~ phyto_binary_increase,
  TRUE ~ phyto_binary_decrease
)) |> 
    mutate(combined_phyto_response_binary_increasepriority = case_when(
 prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
  is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
  TRUE ~ phyto_binary_increase
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response_binary = case_when(
  is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(cyano_binary_decrease) ~ cyano_binary_increase,
  TRUE ~ cyano_binary_decrease
)) |>
       mutate(combined_cyano_response_binary_increasepriority = case_when(
  prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
  prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
  TRUE ~ cyano_binary_increase
)) |> 
   mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
  mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
  mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))
```

## Set up logistic regression for effect of water level change on phytoplankton and cyanobacteria communities

We ran the logistic regression three ways. In the manuscript, the logistic regression includes both water level increases and decreases. We also prioritized only water level increases (any lake that had both a water level increase and decrease was kept as only an increase) and subsequently prioritized only water level decreases (any lake that had both a water level increase and decrease was kept as only a decrease)

```{r}
phytoincrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced) |> 
  #mutate(phyto_binary_decrease = NA) |> 
  filter(!is.na(phyto_response_wl_up)) |> 
  filter(!is.na(prop_1delta_wl_increase))

phytodecrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced) |> 
  filter(!is.na(phyto_response_wl_down)) |> 
  filter(!is.na(prop_1delta_wl_decrease))

phyto_logreg <- rbind(phytodecrease_logreg, phytoincrease_logreg)

phytos <- phyto_logreg |> 
  mutate(response = as.numeric(ifelse(!is.na(prop_1delta_wl_decrease), phyto_binary_decrease, 
                                 ifelse(!is.na(prop_1delta_wl_increase), phyto_binary_increase, NA))), 
         deltawl = as.numeric(ifelse(!is.na(prop_1delta_wl_decrease), prop_1delta_wl_decrease, 
                                 ifelse(!is.na(prop_1delta_wl_increase), prop_1delta_wl_increase, NA)))) |> 
  dplyr::select(response, deltawl, name_yr, trophic_status_reduced) |> 
  mutate(type = "phyto")


ggplot(phytos, aes(x = deltawl, y = response)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.4) +
  stat_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = TRUE) + labs(
    x = "Proportional Water Level Change",
    y = "Probability of Phytoplankton Increase") +
  theme_classic()

cyanoincrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced) |> 
  mutate(cyano_binary_decrease = NA) |> 
  filter(!is.na(cyano_response_wl_up)) |> 
  filter(!is.na(prop_1delta_wl_increase))

cyanodecrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced) |> 
  filter(!is.na(cyano_response_wl_down)) |> 
  filter(!is.na(prop_1delta_wl_decrease))

cyano_logreg <- rbind(cyanodecrease_logreg, cyanoincrease_logreg)

cyanos <- cyano_logreg |> 
  mutate(
    response = as.numeric(
      ifelse(!is.na(prop_1delta_wl_decrease), cyano_binary_decrease, 
      ifelse(!is.na(prop_1delta_wl_increase), cyano_binary_increase, NA))
    ), 
    deltawl = as.numeric(
      ifelse(!is.na(prop_1delta_wl_decrease), prop_1delta_wl_decrease, 
      ifelse(!is.na(prop_1delta_wl_increase), prop_1delta_wl_increase, NA))
    )
  ) |>
  mutate(type = "cyano") |> 
  dplyr::select(response, deltawl, name_yr, trophic_status_reduced, type)


ggplot(cyanos, aes(x = deltawl, y = response)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.4) +
  stat_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = TRUE) +
  labs(
    x = "Proportional Water Level Change",
    y = "Probability of Cyanobacteria Increase"
  ) +
  theme_classic()

all_logreg <- rbind(phytos, cyanos)


ggplot(all_logreg, aes(x = deltawl, y = response)) +
  geom_point(aes(color = trophic_status_reduced), alpha = 0.6, size = 4) +
  geom_smooth(
    method = "glm",
    method.args = list(family = "binomial"),
    se = FALSE,
    color = "black"
  ) +
  facet_wrap(~type) +
  theme_bw() +
  labs(
    x = "Proportional Water Level Change",
    y = "Probability of Increase"
  )

logregform_phyto <-  glm(formula=response ~ deltawl, data=phytos, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response ~ deltawl, data=cyanos, family=binomial)
summary(logregform_cyano)


# cyanodecreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
#   filter(!is.na(prop_delta_wl_priorityincrease)) %>%
#   filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
#   filter(prop_alldelta_wl_decrease<0) %>% 
#   rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
#          response_increase = combined_cyano_response_increasepriority) %>% 
#   mutate(type = "cyano")

# cyanoincreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
#   filter(!is.na(prop_delta_wl_priorityincrease)) %>%
#   filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
#   filter(prop_alldelta_wl_increase>0) %>% 
#   rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
#          response_increase = combined_cyano_response_increasepriority) %>% 
#   mutate(type = "cyano")
# 
# phytoincreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
#   filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
#   filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
#   filter(prop_alldelta_wl_increase>0) %>% 
#   rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
#          response_increase = combined_phyto_response_increasepriority) %>% 
#   mutate(type = "phyto")
# 
# phytodecreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
#   filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
#   filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
#   filter(prop_alldelta_wl_decrease<0) %>% 
#   rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
#          response_increase = combined_phyto_response_increasepriority) %>% 
#   mutate(type = "phyto")
#   
# cyanodata <- boxplotdata2_opposite %>% 
#   select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
#   filter(!is.na(prop_delta_wl_priorityincrease)) %>%
#   filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
#   # filter(prop_delta_wl<0) %>% 
#   rename(response_binary = combined_cyano_response_binary_increasepriority, 
#          response = combined_cyano_response_increasepriority) %>% 
#   mutate(type = "cyano") |> 
#   mutate(wl_priority = "increase")
# 
# phytodata <- boxplotdata2_opposite %>% 
#   select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
#   filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
#   filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
#   # filter(prop_delta_wl<0) %>% 
#   rename(response_binary = combined_phyto_response_binary_increasepriority, 
#          response = combined_phyto_response_increasepriority) %>% 
#   mutate(type = "phyto")|> 
#   mutate(wl_priority = "increase")
# # logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
# logregdata_increase <- rbind(cyanodata, phytodata)
# #"#7A524B","#3C7F29","#8DBAE9"
# 
# cyanodecreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
#   filter(!is.na(prop_delta_wl)) %>%
#   filter(!is.na(combined_cyano_response_binary)) %>% 
#   filter(prop_alldelta_wl_decrease<0) %>% 
#   rename(response_binary = combined_cyano_response_binary, 
#          response = combined_cyano_response) %>% 
#   mutate(type = "cyano")
# 
# cyanoincreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
#   filter(!is.na(prop_delta_wl)) %>%
#   filter(!is.na(combined_cyano_response_binary)) %>% 
#   filter(prop_alldelta_wl_increase>0) %>% 
#   rename(response_binary = combined_cyano_response_binary, 
#          response = combined_cyano_response) %>% 
#   mutate(type = "cyano")
# 
# phytoincreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
#   filter(!is.na(prop_delta_wl)) %>% 
#   filter(!is.na(combined_phyto_response_binary)) %>% 
#   filter(prop_alldelta_wl_increase>0) %>% 
#   rename(response_binary = combined_phyto_response_binary, 
#          response = combined_phyto_response) %>% 
#   mutate(type = "phyto")
# 
# phytodecreasedata <- boxplotdata2_opposite %>% 
#   select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
#   filter(!is.na(prop_delta_wl)) %>% 
#   filter(!is.na(combined_phyto_response_binary)) %>% 
#   filter(prop_alldelta_wl_decrease<0) %>% 
#   rename(response_binary = combined_phyto_response_binary, 
#          response = combined_phyto_response) %>% 
#   mutate(type = "phyto")
#   
# cyanodata <- boxplotdata2_opposite %>% 
#   select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr, trophic_status_reduced) %>% 
#   filter(!is.na(prop_delta_wl)) %>%
#   filter(!is.na(combined_cyano_response_binary)) %>% 
#   # filter(prop_delta_wl<0) %>% 
#   rename(response_binary = combined_cyano_response_binary, 
#          response = combined_cyano_response) %>% 
#   mutate(type = "cyano") |> 
#   mutate(wl_priority = "decrease")
# 
# phytodata <- boxplotdata2_opposite %>% 
#   select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
#   filter(!is.na(prop_delta_wl)) %>% 
#   filter(!is.na(combined_phyto_response_binary)) %>% 
#   # filter(prop_delta_wl<0) %>% 
#   rename(response_binary = combined_phyto_response_binary, 
#          response = combined_phyto_response) %>% 
#   mutate(type = "phyto") |> 
#   mutate(wl_priority = "decrease")
# 
# # logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
# logregdata_decrease <- rbind(cyanodata, phytodata)
# logregdata <- logregdata_decrease
# #"#7A524B","#3C7F29","#8DBAE9"

```

## Plot logistic regression with combined water level increase/decrease

This chunk will generate logistic regression plots for both phytoplankton and cyanobacteria.

```{r}
logregdata_increase1 <- logregdata_increase |>
  rename(prop_delta_wl = prop_delta_wl_priorityincrease)

new_all <- bind_rows(logregdata_decrease, logregdata_increase1) |>
  select(-c(wl_priority)) |>
  unique()

all <- new_all %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) +
  #geom_jitter()+
 ylab("Response to changes in water level") +
  xlab("Proportion water level change") +
  ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10),
        axis.title.y=element_text(angle=90, vjust = .5, size = 12),
        legend.title=element_blank())
all
#117733
#factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic"))
logreg_phyto <- new_all %>% 
  filter(type == "phyto") |> 
  # ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  ggplot(aes(x = prop_delta_wl, y = response_binary, shape = as.factor(type))) +
  scale_color_manual(values = c("#3355DB", "#FFC51E", "#B1C3A2", "#549C79", "gray"), name = "")+
  scale_shape_manual(values = c(20, 10, 11, 12, 13)) +
  # geom_point(size = 10, alpha = 0.5, aes(color = as.factor(trophic_status_reduced))) +
    geom_point(
    aes(
      color = as.factor(factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic", "not reported"))),
      shape = as.factor(type),
      text = paste(
        "Name / Year:", name_yr,
        "<br>Trophic status:", trophic_status_reduced,
        "<br>Δ WL:", prop_delta_wl,
        "<br>Response:", response_binary
      )
    ),
    size = 14,
    alpha = 0.75
  ) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 6, color = "black") +
  #geom_jitter()+
 # ylab("Phytoplankton response to changes in water level") +
 #  xlab("Proportion water level change") +
   ylab("") +
  xlab("") +
  #ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 0.5, size = 20),
        axis.text.y = element_text(hjust = 1, size = 20),
        axis.title.y=element_text(angle=90, hjust = 1, size = 28),
        axis.title.x=element_text(vjust = 1, size = 28),
        legend.text  = element_text(size = 28),
        legend.title=element_blank())

plotly_p <- ggplotly(logreg_phyto)

plotly_p # I have this as a plotly to be able to see exactly what's going into this for now. Useful, but can probably be removed for publishing
logreg_phyto

logregdata_phyto <- new_all %>% subset(type == "phyto")
logregdata_cyano <- new_all %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)

ggsave(paste0(here(),"/Systematic/Plots/logregposter_opposite.jpg"), plot = logreg_phyto, height = 9.5, width = 20, units = "in")
ggsave(paste0(here(),"/Systematic/Plots/logreg_opposite.jpg"), plot = all, height = 4, width = 8, units = "in")

```

## LogReg no NS response

```{r}

boxplotdata2_opposite <- prop_dl %>% 
  #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
    mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", NA, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", NA, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    mutate(combined_phyto_response_binary = case_when(
  is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(phyto_binary_decrease) ~ phyto_binary_increase,
  TRUE ~ phyto_binary_decrease
)) |> 
    mutate(combined_phyto_response_binary_increasepriority = case_when(
 prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
  is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
  TRUE ~ phyto_binary_increase
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response_binary = case_when(
  is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(cyano_binary_decrease) ~ cyano_binary_increase,
  TRUE ~ cyano_binary_decrease
)) |>
       mutate(combined_cyano_response_binary_increasepriority = case_when(
  prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
  prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
  TRUE ~ cyano_binary_increase
)) |> 
   mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
  mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
  mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))
```

## Set up logistic regression for effect of water level change on phytoplankton and cyanobacteria communities

We ran the logistic regression three ways. In the manuscript, the logistic regression includes both water level increases and decreases. We also prioritized only water level increases (any lake that had both a water level increase and decrease was kept as only an increase) and subsequently prioritized only water level decreases (any lake that had both a water level increase and decrease was kept as only a decrease)

```{r}
cyanodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary_increasepriority, 
         response = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "increase")

phytodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary_increasepriority, 
         response = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")|> 
  mutate(wl_priority = "increase")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_increase <- rbind(cyanodata, phytodata)
#"#7A524B","#3C7F29","#8DBAE9"

cyanodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_decrease <- rbind(cyanodata, phytodata)
logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"

```

## Plot logistic regression with combined water level increase/decrease

This chunk will generate logistic regression plots for both phytoplankton and cyanobacteria.

```{r}
logregdata_increase1 <- logregdata_increase |>
  rename(prop_delta_wl = prop_delta_wl_priorityincrease)

new_all <- bind_rows(logregdata_decrease, logregdata_increase1) |>
  select(-c(wl_priority)) |>
  unique()

all <- new_all %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) +
  #geom_jitter()+
 ylab("Response to changes in water level") +
  xlab("Proportion water level change") +
  ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10),
        axis.title.y=element_text(angle=90, vjust = .5, size = 12),
        legend.title=element_blank())
all
#117733
#factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic"))
logreg_phyto <- new_all %>% 
  filter(type == "phyto") |> 
  # ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  ggplot(aes(x = prop_delta_wl, y = response_binary, shape = as.factor(type))) +
  scale_color_manual(values = c("#3355DB", "#FFC51E", "#B1C3A2", "#549C79", "gray"), name = "")+
  scale_shape_manual(values = c(20, 10, 11, 12, 13)) +
  # geom_point(size = 10, alpha = 0.5, aes(color = as.factor(trophic_status_reduced))) +
    geom_point(
    aes(
      color = as.factor(factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic", "not reported"))),
      shape = as.factor(type),
      text = paste(
        "Name / Year:", name_yr,
        "<br>Trophic status:", trophic_status_reduced,
        "<br>Δ WL:", prop_delta_wl,
        "<br>Response:", response_binary
      )
    ),
    size = 14,
    alpha = 0.75
  ) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 6, color = "black") +
  #geom_jitter()+
 # ylab("Phytoplankton response to changes in water level") +
 #  xlab("Proportion water level change") +
   ylab("") +
  xlab("") +
  #ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 0.5, size = 20),
        axis.text.y = element_text(hjust = 1, size = 20),
        axis.title.y=element_text(angle=90, hjust = 1, size = 28),
        axis.title.x=element_text(vjust = 1, size = 28),
        legend.text  = element_text(size = 28),
        legend.title=element_blank())

plotly_p <- ggplotly(logreg_phyto)

plotly_p # I have this as a plotly to be able to see exactly what's going into this for now. Useful, but can probably be removed for publishing
logreg_phyto

logregdata_phyto <- new_all %>% subset(type == "phyto")
logregdata_cyano <- new_all %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)

ggsave(paste0(here(),"/Systematic/Plots/logregposter_ns.jpg"), plot = logreg_phyto, height = 9.5, width = 20, units = "in")
ggsave(paste0(here(),"/Systematic/Plots/logreg_ns.jpg"), plot = all, height = 4, width = 8, units = "in")

```
