---
title: "SystematicReview_Final"
format: html
editor: visual
---

This script reads in the extracted data from the systematic literature review.

## Libraries

libraries for the whole script!

```{r packages}
#| message: false
#| warning: false
library(janitor)
library(ggplot2)
library(tidyverse)
library(dplyr)

library(rnaturalearth)
library(rnaturalearthdata)
library(mapview)

library(naniar)

library(sf)
library(dplyr)
library(rnaturalearth)
library(rnaturalearthdata)

library(here)
library(plotly)
library(patchwork)
```

#Read in data I preprocessed colnames for the data.

```{r read in data}
data <- "Systematic/Combined_08Oct.csv"
init_data <- read.csv(here(data))
init_data <- clean_names(init_data)
head(init_data)
```

# QAQC

## Excluded papers

Papers were excluded for the following reasons:

"Abirhire 2019": Has 2 years of data, Abirhire 2023 has 8 years of data BUT it's not presented in a way that is particularly usable \> doesn't fit as well. Excluded 2019 because that was our criteria

Beaver 2013: we contacted the author and there was no WL data available

Braga 2015: duplicate data with Medeiros paper, and Medeiros paper has an extra year of data

Braga 2020: duplicated data with deMendonca paper, deMendonca paper has a few more months of data

Cao 2016 & Jia 2022: Poyang

Christensen 2015: Compared two rule curves, but didn't compare a high and low condition - the later rule curve generally meant higher water level, but not for all of the reservoirs. There's not a way to answer our questions with the data provided.

Geraldes 2007 - Serra Serrada Reservoir: duplicate data with Geraldes 2005. Azivo Reservoir is still included from Geraldes 2007. Azivo Reservoir has no water level change and the compared time periods do not give a change in water level. 

Hart 2004: Data is either duplicated or more data is available in Hart 2006, so excluding Hart 2004

Kisand 2004 & Wang 2022: these papers were not clear enough to extract meaningful data from.

Lacerda 2018: same data are available in Santos, and Lacerda has a stronger focus on remote sensing

Naselli Flores 2003: duplicate data with other Naselli Flores 2003 paper, retained the one with more data

Nõges 2010 & Tuvikene 2011 & Noges 1999: duplicate data with other Nõges papers

Sakharova 2018: 3 years of data, Mineeva has longer time series

Sharip: paper got duplicated, removed the duplicate version. Also, Bera lake not more than 4.5m deep

```{r exclude papers}
#| warning: false
data_excl <- init_data |> 
  filter(!name_yr %in% c("Beaver 2013","Jia 2022","Cao 2016","Wang 2022","Kisand 2004", "Hart 2004", "Nõges 2010", "Tuvikene 2011", "Christensen 2015", "Braga 2015", "Braga 2020", "Lacerda 2018", "Abirhire 2019", "Sakharova 2018", "Noges 1999", "Geraldes 2007")) |>
  filter(!res_name == "Serra Serrada Reservoir") |> #Geraldes 2007
  filter(!res_name == "Bera Lake") |> #Sharip 2019
  filter(!title %in% c("Steady-state assemblages in a Mediterranean hypertrophic reservoir. The role of Microcystis ecomorphological variability in maintaining an apparent equilibrium")) |>
  # filter(res_name != "Dourado" | name_yr != "deMednonca 2018") |> #Naselli-Flores 2003
#Data is either duplicated or more data is available in Hart 2006, so excluding Hart 2004
  filter(!notes %in% c("Ignore this version, everything has been added to multiple reservoirs sheet! KKH"))
#Beaver no WL data
#Jia 
#duplicate Sharip

excluded_papers <- anti_join(init_data, data_excl)

duplicate_waterbodies <- data_excl |> 
  select(name_yr, res_name, country, study_years, consecutive, phys_response, notes) #used this to order data alphabetically and manually check for dups that were not named exactly the same

```

## Typos, trophic levels

Fix typos, add trophic levels based on chla and P concentrations if available

Need to fix Vrutci - Water level increase is 626.7-621.3 and water level decrease is 621.3-615

```{r fix some data}
data <- data_excl |> 
  mutate(tp = mean_tp) |> 
  separate_wider_delim(mean_tp, " ", names = c("tp_ugL", "unit_tp"), too_few = "align_start") |> 
  mutate(mean_chla = ifelse(mean_chla == "0.07 to 34.62 ug/L", "8", mean_chla)) |> #change Rai 1978 to 8ug/L, actual range between 0 and 14 ug/L, original range from pheophyton
  mutate(mean_chla = ifelse(mean_chla == "15.7ug/L", "15.7 ug/L", mean_chla)) |> #fix missing space in chla value
  mutate(chla = mean_chla) |> 
  separate_wider_delim(mean_chla, " ", names = c("chla_ugL", "unit_chla"), too_few = "align_start") |>
  mutate(res_name = ifelse(res_name == "Rybinsk", "Rybinsk Reservoir", #standardize reservoir names
                    ifelse(res_name == "Cruzeta man-made lake", "Cruzeta", 
                    ifelse(res_name == "Castanhao", "Castanhão reservoir",
                    ifelse(name_yr == "Bouvy 2003", "Tapacura Reservoir", res_name))))) |> 
  filter((!is.na(cov_num)|!is.na(num_water)|!is.na(lat_dd))) |> #remove empty or duplicate columns from the df
  mutate(fullpond_sa = ifelse(fullpond_sa == "~600", 600, fullpond_sa), #fix character in numeric columns
         fullpond_sa = as.numeric(fullpond_sa)) |> 
   mutate(increase_phyto_response = ifelse(name_yr == "Li 2024", "Not measured", increase_phyto_response), 
         increase_cyano_response = ifelse(name_yr == "Li 2024", "Not measured", increase_cyano_response)) |> # there was no water level increase recorded for this paper and it focused on the effects of water level decreases/reservoir operation that resulted
  mutate(obs_maxdepth = ifelse(obs_maxdepth == "SEK had 12, 12 max at sampling station", 12, obs_maxdepth), #fix text in numeric columns
         mindepth = ifelse(mindepth == "SEK had 9, 9 was min at sampling station", 9, mindepth)) |> 
  mutate(elev_decrease_max = ifelse(grepl("2001", elev_decrease_max), "-6", elev_decrease_max), # to use the 2003 value
         elev_increase_max = ifelse(elev_increase_max == 0, NA, elev_increase_max)) |> #to remove increases of 0
  mutate(long_dd = ifelse(res_name == "Lake Buchanan", -98.4, long_dd), #fix incorrect latlongs
         lat_dd = ifelse(lat_dd == 103.466667 & res_name == "Zipingpu Reservoir", 30.983333, lat_dd), 
         long_dd = ifelse(long_dd == 30.983333 & res_name == "Zipingpu Reservoir", 103.466667, long_dd),
         long_dd = ifelse(long_dd == 6.77888889 & res_name == "Serra Serrada", -6.77888889, long_dd)) |> 
  mutate(tp_ugL = as.numeric(tp_ugL)) |> #add in trophic state based on phosphorus and chlorophyll values. 
  mutate(mean_tp = case_when(
    unit_tp == "mg/L" ~ tp_ugL * 1000, 
    unit_tp == "mg/m3" ~ tp_ugL / 1000, 
    TRUE ~ tp_ugL)) |> 
  mutate(chla_ugL = as.numeric(chla_ugL)) |> 
  mutate(mean_chla = case_when(
    unit_chla == "mg/L" ~ chla_ugL * 1000, 
    unit_chla == "mg/m3" ~ chla_ugL / 1000, 
     TRUE ~ chla_ugL)) |> 
  mutate(trophic = ifelse(trophic == "Other: according to authors, varies according to water level; never more than mesotrophic unless water level is low", "mesotrophic", trophic)) %>% 
  mutate(trophic = ifelse(trophic == "", "not reported", trophic)) %>% 
  mutate(chla_ugL = as.numeric(chla_ugL)) |>  
  mutate(tp_ugL = as.numeric(tp_ugL)) |> 
  mutate(trophic_status_chla = ifelse(trophic == "not reported" & chla_ugL <= 2.6, "oligotrophic", 
                               ifelse(trophic == "not reported" & chla_ugL >2.6 & chla_ugL <=7.3, "mesotrophic", 
                               ifelse(trophic == "not reported" & chla_ugL >7.3 & chla_ugL <= 56, "eutrophic", 
                               ifelse(trophic == "not reported" & chla_ugL >56, "hypereutrophic",
                               ifelse(trophic == "not reported" & is.na(chla_ugL), trophic, trophic)))))) %>% 
  mutate(trophic_status_tp = ifelse(tp_ugL <= 12, "oligotrophic", 
                             ifelse(tp_ugL >=12 & tp_ugL <=24, "mesotrophic", 
                             ifelse(tp_ugL >24 & tp_ugL <= 70, "eutrophic", 
                             ifelse(tp_ugL >70, "hypereutrophic",
                             ifelse(is.na(tp_ugL), trophic, NA)))))) %>%  
  mutate(trophic_status_combined = ifelse(trophic == "not reported" & !is.na(trophic_status_tp), trophic_status_tp,
                                   ifelse(trophic == "not reported" & !is.na(trophic_status_chla), trophic_status_chla, trophic))) |> 
  mutate(trophic_status_reduced = ifelse(trophic_status_combined %in% c("oligotrophic"), "oligotrophic", #combine multi trophic state options
                                  ifelse(trophic_status_combined %in% c("oligo-mesotrophic", "mesotrophic"), "mesotrophic", 
                                  ifelse(trophic_status_combined %in% c("meso-eutrophic", "eutrophic"), "eutrophic", 
                                  ifelse(trophic_status_combined %in% c("hypereutrophic", "eu-hypereutrophic"), "hypereutrophic", "not reported"))))) |> 
  filter(res_name != "") #remove any rows that don't have a reservoir name because they are duplicate with reservoirs in the multiple reservoir sheet
write_csv(data, here("Systematic/qaqced.csv"))
```

#Quick info

```{r quick study info}
#number of waterbodies
reservoirs <- data |> select(c(res_name)) |> unique()
#there are a total of 4 TGRs, so there are 64-3 reservoirs, = 61 total lakes!
countries <- data |> select(c(country)) |> unique() #25 countries - Argentina/Uruguay, Estonia/Russia, one blank = 22 total countries
studies <- data |> select(c(name_yr)) |> unique() #60 studies, some duplicate authors but no duplicate studies.
```

# Plots

## Trophic Levels

```{r trophic levels}
#consider subsetting these to phyto and cyano specific studies?

non_dups <- data |> select(res_name, trophic_status_reduced) |> 
  unique()

#plot only reduced trophic state levels and only reservoirs that are not duplicated
#from 81 > 72
trophic <- non_dups |> 
  filter(trophic_status_reduced != "not reported") |> 
  ggplot(aes(x= factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic")))) + 
  geom_bar(fill = "#608687") +
  xlab("Trophic Status") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

trophic
ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/trophic.jpg", plot = trophic, height = 3, width = 6.5, units = "in")

```


Phytoplankton drivers

List of physical variables that we extracted that were measured and considered as drivers:
water temperature
water level
secchi/transparency
light(PAR, photic zone)
turbidity
total suspended solids
total dissolved solids
other

chemical variables
pH
DO
DOC
DIC
DOM
POM
TP
SRP
ORP (orthophosphate)
TN
TKN
NO3
NO2
NH4
DIN
metals 
cond
salinity
other

## Phytoplankton drivers

```{r}
drivers <- data |> 
  # mutate(filt_chla = ifelse(grepl("filtered chlorophyll", bio_vars), 1, 0),                       #biotic variables
  #        phyto_micros = ifelse(grepl("phytoplankton microscopy", bio_vars), 1, 0), 
  #        single_cyano_taxon = ifelse(grepl("single cyanobacterial taxon", bio_vars), 1, 0), 
  #        sensor_chla = ifelse(grepl("sensor", bio_vars), 1, 0),
  #       # sensor_chla = ifelse(grepl("sensor", bio_vars), 1, 0),
  #        other_bio = ifelse(grepl("Other", bio_vars), 1, 0)) |> 
  mutate(tp = ifelse(grepl("TP", chem_drive), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_drive), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_drive), 1, 0), 
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         do = ifelse(grepl("DO", chem_drive), 1, 0),
         srp = ifelse(grepl("SRP", chem_drive), 1, 0),
         no3 = ifelse(grepl("NO3", chem_drive), 1, 0),
         cond = ifelse(grepl("conductivity", chem_drive), 1, 0),
         metals = ifelse(grepl("metals", chem_drive), 1, 0),
         orp = ifelse(grepl("ORP", chem_drive), 1, 0),
         no2 = ifelse(grepl("NO2", chem_drive), 1, 0),
         dic = ifelse(grepl("DIC", chem_drive), 1, 0),
       dom = ifelse(grepl("DOM", chem_drive), 1, 0),
       pom = ifelse(grepl("POM", chem_drive), 1, 0),
       tkn = ifelse(grepl("TKN", chem_drive), 1, 0),
         doc = ifelse(grepl("DOC", chem_drive), 1, 0),
         din = ifelse(grepl("DIN", chem_drive), 1, 0),
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         salinity = ifelse(grepl("salinity", chem_drive), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_drive), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_drive), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_drive), 1, 0),
         transparency = ifelse(grepl("transparency", phys_drive), 1, 0),
         turb = ifelse(grepl("turbidity", phys_drive), 1, 0),
         light = ifelse(grepl("PAR", phys_drive), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid", phys_drive), 1, 0), 
         strat = ifelse(grepl("mix|stab|strat", phys_drive), 1, 0),
         restime = ifelse(grepl("residence", phys_drive), 1, 0), 
         flush_discharge = ifelse(grepl("flush|discharge", phys_drive), 1, 0), 
         inflow = ifelse(grepl("inflow", phys_drive), 1, 0),
         precip = ifelse(grepl("precip|rain", phys_drive), 1, 0),
         air_temp = ifelse(grepl("air", phys_drive), 1, 0),
         other_phys = ifelse(grepl("Other", phys_drive), 1, 0))

other_phys <- drivers |> 
  filter(other_phys == 1) |> 
  select(name_yr, title, num_water, phys_drive)

  unique(data$phys_response)
  
  #WLchangevars is whether or not vars changed under WL change
WLchangevars <- data |> 
mutate(not_assessed = ifelse(grepl("not assessed", phys_response), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_response), 1, 0),
         transparency = ifelse(grepl("transparency", phys_response), 1, 0),
         turb = ifelse(grepl("turbidity", phys_response), 1, 0),
         light = ifelse(grepl("PAR", phys_response), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_response), 1, 0), 
         #strat = ifelse(grepl("mix|stab|strat", phys_response), 1, 0),
         #restime = ifelse(grepl("residence", phys_response), 1, 0), 
         #flush_discharge = ifelse(grepl("flush|discharge", phys_response), 1, 0), 
         #inflow = ifelse(grepl("inflow", phys_response), 1, 0),
         #precip = ifelse(grepl("precip|rain", phys_response), 1, 0),
         #air_temp = ifelse(grepl("air", phys_response), 1, 0),
         other_phys = ifelse(grepl("Other", phys_response), 1, 0), 
       watercolor = ifelse(grepl("water color", phys_response), 1, 0)) |> 
mutate(tp = ifelse(grepl("TP", chem_response), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_response), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_response), 1, 0), 
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         do = ifelse(grepl("DO", chem_response), 1, 0),
         srp = ifelse(grepl("SRP", chem_response), 1, 0),
         no3 = ifelse(grepl("NO3", chem_response), 1, 0),
         cond = ifelse(grepl("conductivity", chem_response), 1, 0),
         metals = ifelse(grepl("metals", chem_response), 1, 0),
         orp = ifelse(grepl("ORP", chem_response), 1, 0),
         no2 = ifelse(grepl("NO2", chem_response), 1, 0),
         doc = ifelse(grepl("DOC", chem_response), 1, 0),
       dic = ifelse(grepl("DIC", chem_response), 1, 0),
       dom = ifelse(grepl("DOM", chem_response), 1, 0),
       pom = ifelse(grepl("POM", chem_response), 1, 0),
       tkn = ifelse(grepl("TKN", chem_response), 1, 0),
         din = ifelse(grepl("DIN", chem_response), 1, 0),
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         salinity = ifelse(grepl("salinity", chem_response), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_response), 1, 0)) 

#spread vars is whether or not the var was measured
spreadvars <- data |> 
  mutate(filt_chla = ifelse(grepl("filtered chlorophyll", bio_vars), 1, 0),                       #biotic variables
         phyto_micros = ifelse(grepl("phytoplankton microscopy", bio_vars), 1, 0), 
         single_cyano_taxon = ifelse(grepl("single cyanobacterial taxon", bio_vars), 1, 0), 
         sensor_chla = ifelse(grepl("sensor", bio_vars), 1, 0),
        # sensor_chla = ifelse(grepl("sensor", bio_vars), 1, 0),
         other_bio = ifelse(grepl("Other", bio_vars), 1, 0)) |> 
  mutate(tp = ifelse(grepl("TP", chem_vars), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_vars), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_vars), 1, 0), 
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         do = ifelse(grepl("DO", chem_vars), 1, 0),
         srp = ifelse(grepl("SRP", chem_vars), 1, 0),
         no3 = ifelse(grepl("NO3", chem_vars), 1, 0),
         cond = ifelse(grepl("conductivity", chem_vars), 1, 0),
         metals = ifelse(grepl("metals", chem_vars), 1, 0),
         orp = ifelse(grepl("ORP", chem_vars), 1, 0),
         no2 = ifelse(grepl("NO2", chem_vars), 1, 0),
         doc = ifelse(grepl("DOC", chem_vars), 1, 0),
          dic = ifelse(grepl("DIC", chem_vars), 1, 0),
       dom = ifelse(grepl("DOM", chem_vars), 1, 0),
       pom = ifelse(grepl("POM", chem_vars), 1, 0),
       tkn = ifelse(grepl("TKN", chem_vars), 1, 0),
         din = ifelse(grepl("DIN", chem_vars), 1, 0),
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         salinity = ifelse(grepl("salinity", chem_vars), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_vars), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_vars), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_vars), 1, 0),
         transparency = ifelse(grepl("transparency", phys_vars), 1, 0),
         turb = ifelse(grepl("turbidity", phys_vars), 1, 0),
         light = ifelse(grepl("PAR", phys_vars), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
        tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_vars), 1, 0), 
         #strat = ifelse(grepl("mix|stab|strat", phys_vars), 1, 0),
        # restime = ifelse(grepl("residence", phys_vars), 1, 0), 
        # flush_discharge = ifelse(grepl("flush|discharge", phys_vars), 1, 0), 
        # inflow = ifelse(grepl("inflow", phys_vars), 1, 0),
        # precip = ifelse(grepl("precip|rain", phys_vars), 1, 0),
        # air_temp = ifelse(grepl("air", phys_vars), 1, 0),
         other_phys = ifelse(grepl("Other", phys_vars), 1, 0)) 
        # watercolor = ifelse(grepl("water color", phys_vars), 1, 0))


unique(data$chem_drive)
unique(data$phys_drive)

# vars <- drivers |> select(c(filt_chla, phyto_micros, single_cyano_taxon, sensor_chla, other_bio, other_chem, tp, tn, nh4, ph, do, srp, no3, cond, metals, orp, no2, doc, din, ph, salinity))

chemvars_drive <- drivers |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
chemvarsWL <- WLchangevars |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
physvars_drive <- drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))
#physvarsWL <- WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp))
physvarsWL <- WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys))
# summary(vars) 

freqchem <- as.data.frame(colSums(chemvars_drive))

chemvars_drive <- setNames(cbind(rownames(freqchem), freqchem, row.names = NULL), 
         c("names", "sum"))

freqchemWL <- as.data.frame(colSums(chemvarsWL))

chemvarsWL <- setNames(cbind(rownames(freqchemWL), freqchemWL, row.names = NULL), 
         c("names", "sum"))

freqphys <- as.data.frame(colSums(physvars_drive))
freqphysWL <- as.data.frame(colSums(physvarsWL))
rm(physvars_drive)
physvars2_drive <- setNames(cbind(rownames(freqphys), freqphys, row.names = NULL), 
         c("names", "sum"))

physvars2WL <- setNames(cbind(rownames(freqphysWL), freqphysWL, row.names = NULL), 
         c("names", "sum"))
```

```{r}
#separate out phytoplankton and cyanobacteria drivers
phyto_only_drivers <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:125)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_drivers <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:125)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

#WL change vars
phyto_only_WLchangevars <- WLchangevars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_WLchangevars <- WLchangevars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

#not WL change vars! This is anything measured
phyto_only_vars <- spreadvars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:124)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_vars <- spreadvars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:124)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

chemvars_drive_phytos <- phyto_only_drivers |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))

#only including variables that were measured at least 10 times in the extracted data
chemvars_phytos <- phyto_only_vars |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))

physvars_drive_phytos <- phyto_only_drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))

# physvarsWL_phytos <- phyto_only_WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp))

physvars_phytos <- phyto_only_vars |> select(c(water_temp, water_level, transparency, turb, light, tss_tds, other_phys))
# summary(vars) 

freqchem_drive_phytos <- as.data.frame(colSums(chemvars_drive_phytos))
chemvars_drive_phytos <- setNames(cbind(rownames(freqchem_drive_phytos), freqchem_drive_phytos, row.names = NULL), 
         c("names", "sum"))
# freqchemWL_phyto <- as.data.frame(colSums(chemvarsWL_phytos))
freqchem_phyto <- as.data.frame(colSums(chemvars_phytos))
# chemvarsWL_phytos <- setNames(cbind(rownames(freqchemWL), freqchemWL_phyto, row.names = NULL), 
#          c("names", "sum"))
chemvars_phytos <- setNames(cbind(rownames(freqchem_phyto), 
                                  freqchem_phyto, row.names = NULL), c("names", "sum"))

# chemvarsWL |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +  ggtitle("Chem variables that changed in response to WL")
# want a comparison with what was measured!
chemvars1_phytos <- chemvars_phytos |> 
  rename(measured = sum, 
         name = names)

# drivers identified out of all vars measured
chemicalvars_phytos <- cbind(chemvars1_phytos, chemvars_drive_phytos)

chem <- chemicalvars_phytos |> 
  filter(!name == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, id, notid, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = "fill", stat = "identity") + 
  ggtitle("Chemical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#99bbab", "#ddefcd")) +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

#physical variables

freqphys_drive <- as.data.frame(colSums(physvars_drive_phytos))
# freqphysWL <- as.data.frame(colSums(physvarsWL_phytos))
freqphys_phytos <- as.data.frame(colSums(physvars_phytos))

physvars2_drive_phytos <- setNames(cbind(rownames(freqphys_drive), freqphys_drive, row.names = NULL), 
         c("names", "sum"))

physvars1_phytos <- setNames(cbind(rownames(freqphys_phytos), freqphys_phytos, row.names = NULL), 
         c("name", "measured"))

physvars_phytos <- cbind(physvars1_phytos, physvars2_drive_phytos)

phys <- physvars_phytos |> 
    select(c(names, measured, sum)) |> 
  filter(!names == "other_phys") |> 
  rename(id = sum) |> 
   mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, id, notid, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>    
  select(c(names, id, notid)) |> 
  mutate(notid = ifelse(notid <0, 0, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = "fill", stat = "identity") +
  # theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  theme(axis.text.x = element_text(vjust = 0.5)) + 
  ggtitle("Physical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#99bbab", "#ddefcd")) + theme_bw() + 
   labs(y = "Proportion") +
    theme(axis.text.x  = element_text(hjust = .5, size = 6),
          axis.title.x = element_blank(),
          axis.title.y = element_text(angle = 90, vjust = .5, size = 12),
          legend.position = "none")

phytos <- phys/chem
phytos
ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/phytovars.jpg", plot = phytos, height = 5, width = 6.5, units = "in")

```

```{r cyanos vars only}
# chemvars_drive_cyanos <- cyano_only_drivers |> select(c(other_chem, tp, tn, nh4, ph, do, srp, no3, cond, metals, orp, no2, doc, din, ph, salinity, dom, pom, dic, tkn))

# only vars that have more than 10
chemvars_drive_cyanos <- cyano_only_drivers |> select(c(cond, do, nh4, no2, no3, ph, srp, tn, tp))

chemvars_cyanos <- cyano_only_vars |> select(c(cond, do, nh4, no2, no3, ph, srp, tn, tp))

physvars_drive_cyanos <- cyano_only_drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds))

# physvarsWL_phytos <- phyto_only_WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp))

physvars_cyanos <- cyano_only_vars |> select(c(water_temp, transparency, turb, light, tss_tds, water_level))

physvars_cyanos <- cyano_only_vars |> select(c(water_temp, transparency, turb, light, tss_tds, water_level))
# summary(vars) 

freqchem_drive_cyanos <- as.data.frame(colSums(chemvars_drive_cyanos))
chemvars_drive_cyanos <- setNames(cbind(rownames(freqchem_drive_cyanos), freqchem_drive_cyanos, row.names = NULL), 
         c("names", "sum"))
# freqchemWL_cyano <- as.data.frame(colSums(chemvarsWL_cyanos))
freqchem_cyano <- as.data.frame(colSums(chemvars_cyanos))
# chemvarsWL_cyanos <- setNames(cbind(rownames(freqchemWL), freqchemWL_cyano, row.names = NULL), 
#          c("names", "sum"))
chemvars_cyanos <- setNames(cbind(rownames(freqchem_cyano), 
                                  freqchem_cyano, row.names = NULL), c("names", "sum"))

# chemvarsWL |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +  ggtitle("Chem variables that changed in response to WL")
# want a comparison with what was measured!
chemvars1_cyanos <- chemvars_cyanos |> 
  rename(measured = sum, 
         name = names)
chemicalvars_cyanos <- cbind(chemvars1_cyanos, chemvars_drive_cyanos)

chemvars_cyanos_drivers <- chemicalvars_cyanos |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
   mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, id, notid, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>    
  select(c(names, id, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names))+
  geom_bar(position = "fill", stat = "identity") + 
 # ggtitle("Proportion Chemical variables that were identified as drivers, cyano only" ) + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#4dd0e1","#b2ebf2")) + theme_bw() + 
    labs(y = "") +
    theme(axis.text.x  = element_text(hjust = .5, size = 6),
          axis.title.x = element_blank(),
          axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

#physical variables

freqphys_drive <- as.data.frame(colSums(physvars_drive_cyanos))
# freqphysWL <- as.data.frame(colSums(physvarsWL_cyanos))
freqphys_cyanos <- as.data.frame(colSums(physvars_cyanos))

physvars2_drive_cyanos <- setNames(cbind(rownames(freqphys_drive), freqphys_drive, row.names = NULL), 
         c("names", "sum"))

physvars1_cyanos <- setNames(cbind(rownames(freqphys_cyanos), freqphys_cyanos, row.names = NULL), 
         c("name", "measured"))

physvars_cyanos <- cbind(physvars1_cyanos, physvars2_drive_cyanos)

physvars_cyanos_drivers <- physvars_cyanos |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
 mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, id, notid, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>    
  select(c(names, id, notid)) |> 
  mutate(notid = ifelse(notid <0, 0, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = "fill", stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
 # ggtitle("Phys driver variables, cyano only") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#4dd0e1","#b2ebf2")) + theme_bw() + 
    labs(y = "Proportion") +
    theme(axis.text.x  = element_text(hjust = .5, size = 6),
          axis.title.x = element_blank(),
          axis.title.y = element_text(angle = 90, vjust = .5, size = 12),
          legend.position = "none")

cyanos <- physvars_cyanos_drivers/chemvars_cyanos_drivers
cyanos
ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/cyanodrivervars.jpg", plot = cyanos, height = 5, width = 6.5, units = "in")

combined <- (physvars_cyanos_drivers/chemvars_cyanos_drivers) | (phys / chem)
combined <- (phys/physvars_cyanos_drivers) | (chem/chemvars_cyanos_drivers)
combined

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/drivers_cyanophyto.jpg", plot = combined, height = 4, width = 8, units = "in")
```

## Map

Make map figure.

```{r}

na_strings <- c("NA", "N A", "N / A", "N/A", "N/ A", "Not Available", "NOt available", "", "not reported")

 latlong <- data %>%
  mutate(
    study_years = ifelse(
      grepl("36 \\(satellite data\\); 12 \\(field data\\)", study_years),"12", 
      ifelse(grepl("36 \\(satellite data\\); 2 \\(field data\\)", study_years),"2",study_years)), 
    study_years = as.numeric(study_years)) |> 
  mutate(depth_m = ifelse(!is.na(fullpond_maxdepth), fullpond_maxdepth, obs_maxdepth), 
         depth_m = as.numeric(depth_m), 
         years_studied = ifelse(study_years == 1, "1", 
                                ifelse(study_years >1 & study_years <=5, "2-5", 
                                       ifelse(study_years >5 & study_years <= 10, "6-10", 
                                              ifelse(study_years > 10, "more than 10", NA)))), 
         years_studied = factor(years_studied, levels = c("1", "2-5", "6-10", "more than 10"))) |> 
  select("lat_dd", "long_dd", "res_name", "decrease_phyto_response", "increase_phyto_response", "decrease_cyano_response", "increase_cyano_response", "depth_m", "trophic_status_reduced", "years_studied") |> 
  mutate(lat_dd = ifelse(lat_dd == 103.466667 & res_name == "Zipingpu Reservoir", 30.983333, lat_dd)) |> 
  mutate(long_dd = ifelse(long_dd == 30.983333 & res_name == "Zipingpu Reservoir", 103.466667, long_dd)) |> 
  mutate(long_dd = ifelse(long_dd == 6.77888889 & res_name == "Serra Serrada", -6.77888889, long_dd)) |> 
    replace_with_na_all(condition = ~.x %in% na_strings) |> 
    subset(!is.na(long_dd)) |>  
  subset(!is.na(lat_dd)) |> 
  mutate(increase_phyto_response = ifelse(increase_phyto_response == "Not measured"|increase_phyto_response == ""|(grepl("Other", increase_phyto_response)), NA, increase_phyto_response)) |>
    mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured"|decrease_phyto_response == ""|(grepl("Other", decrease_phyto_response)), NA, decrease_phyto_response)) |>
   mutate(increase_cyano_response = ifelse(increase_cyano_response == "Not measured"|increase_cyano_response == ""|(grepl("Other", increase_cyano_response)), NA, increase_cyano_response)) |>
   mutate(decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured"|decrease_cyano_response == ""|(grepl("Other", decrease_cyano_response)), NA, decrease_cyano_response)) |>
  filter(!(is.na(decrease_phyto_response) & is.na(increase_phyto_response) & is.na(decrease_cyano_response) & is.na(increase_cyano_response))) %>%
  mutate(increase_decrease_mosaic = ifelse(((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)) & (!is.na(increase_phyto_response) | !is.na(increase_cyano_response))),"Both",
                                           ifelse((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)),"Decrease",
                                                  ifelse((!is.na(increase_phyto_response) | !is.na(increase_cyano_response)),"Increase", NA))))

world <- ne_countries(scale = "medium", returnclass = "sf")
#world <- map_data("world")

litdata_map <- latlong %>% 
  mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured", NA, decrease_phyto_response), 
         increase_phyto_response = ifelse(increase_phyto_response == "Not measured", NA, increase_phyto_response), 
         decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured", NA, decrease_cyano_response), 
         increase_cyano_response = ifelse(increase_cyano_response == "Not measured", NA, increase_cyano_response)) |> 
   mutate(mapvals = ifelse(((!is.na(decrease_phyto_response)|!is.na(increase_phyto_response)) & (!is.na(decrease_cyano_response)|!is.na(increase_cyano_response))), "Both",
                           ifelse(!is.na(decrease_phyto_response)|!is.na(increase_phyto_response), "Phytoplankton", 
                                 ifelse(!is.na(decrease_cyano_response)|!is.na(increase_cyano_response), "Cyanobacteria", NA)))) 

litdata_map_sf <- st_as_sf(litdata_map,
                           coords = c("long_dd", "lat_dd"),
                           crs = 4326,
                           remove = FALSE)

sitemap <- ggplot(world) +
  geom_sf(fill = "white") +
 # coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white", color = "black") +
  #geom_sf_label(stat = "sf_coordinates")+
  geom_sf(data = litdata_map_sf,
          aes(shape = as.factor(mapvals),
              color = as.factor(mapvals)),
          size = 2) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("#608687", "#4dd0e1", "darkgreen"),
                     name = "Variable") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  coord_sf() +
  theme_minimal()
sitemap

sitemap_trophic <- ggplot(world) +
  geom_sf(data = world, fill = "white") +
  #coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white") +
  #geom_sf_label(stat = "sf_coordinates")+
  # geom_sf(data = litdata_map_sf,
  #         aes(shape = as.factor(mapvals),
  #             color = as.factor(increase_decrease_mosaic))) +
    geom_sf(data = litdata_map_sf, size = 2, alpha = 0.7,
          aes(color = as.factor(increase_decrease_mosaic))) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("#7A5FA8", "#C65A6A", "#3F7FA6"),
                     name = "Does water level \nincrease or decrease? \n") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  theme_minimal() + 
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(legend.text  = element_text(size = 12),
        legend.title = element_text(size = 12))
sitemap_trophic

sitemap_years <- ggplot(world) +
  geom_sf(data = world, fill = "white") +
  #coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white") +
  #geom_sf_label(stat = "sf_coordinates")+
  # geom_sf(data = litdata_map_sf,
  #         aes(shape = as.factor(mapvals),
  #             color = as.factor(increase_decrease_mosaic))) +
    geom_sf(data = litdata_map_sf, size = 2, alpha = 0.6,
          aes(color = as.factor(years_studied))) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("lightgreen", "darkgreen", "blue", "purple"),
                     name = "How many \nyears studied?") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  theme_minimal() + 
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(legend.text  = element_text(size = 16),
        legend.title = element_text(size = 20))
sitemap_years

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/sitemap.jpg", plot = sitemap, height = 4, width = 6.5, units = "in")

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/sitemap_years.jpg", plot = sitemap_years, height = 4, width = 6.5, units = "in")

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/sitemapposter.jpg", plot = sitemap_trophic, height = 4, width = 10, units = "in")
```

## Prop Water Level

```{r max depth and proportion water level}
prop_dl <- data |>
  mutate(obs_maxdepth = ifelse(obs_maxdepth == "", NA, obs_maxdepth)) |> 
  mutate(usedepth = ifelse(is.na(obs_maxdepth), fullpond_maxdepth, obs_maxdepth)) |> 
  #select(c("obs_maxdepth", "fullpond_maxdepth", "usedepth")) |> 
  mutate(elev_decrease_max = as.numeric(elev_decrease_max)) |> 
  mutate(usedepth = as.numeric(usedepth)) |> 
  mutate(prop_1delta_wl_decrease = ifelse(is.na(elev_increase_max), elev_decrease_max/usedepth, NA)) |> 
  mutate(prop_alldelta_wl_decrease = elev_decrease_max/usedepth) |> 
  mutate(prop_alldelta_wl_decrease = ifelse(prop_alldelta_wl_decrease > 0, prop_alldelta_wl_decrease* -1, prop_alldelta_wl_decrease)) |> 
  mutate(prop_1delta_wl_decrease = ifelse(prop_1delta_wl_decrease > 0, prop_1delta_wl_decrease* -1, prop_1delta_wl_decrease)) |>
  mutate(prop_alldelta_wl_increase = elev_increase_max/usedepth) |> 
  mutate(prop_1delta_wl_increase = elev_increase_max/usedepth) |> 
  #select(c("obs_maxdepth", "fullpond_maxdepth", "usedepth", "prop_1delta_wl_decrease", "prop_alldelta_wl_decrease", "elev_decrease_max", "elev_increase_max", "prop_alldelta_wl_increase", "prop_1delta_wl_increase"))
  mutate(prop_delta_wl = ifelse(is.na(prop_alldelta_wl_decrease), prop_alldelta_wl_increase, prop_alldelta_wl_decrease)) |> 
  mutate(prop_delta_wl_priorityincrease = ifelse(is.na(prop_alldelta_wl_increase), prop_alldelta_wl_decrease, prop_alldelta_wl_increase))

boxplotdata2 <- prop_dl %>% 
  #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
    mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 0, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 0, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    mutate(combined_phyto_response_binary = case_when(
  is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(phyto_binary_decrease) ~ phyto_binary_increase,
  TRUE ~ phyto_binary_decrease
)) |> 
    mutate(combined_phyto_response_binary_increasepriority = case_when(
 prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
  is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
  TRUE ~ phyto_binary_increase
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response_binary = case_when(
  is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(cyano_binary_decrease) ~ cyano_binary_increase,
  TRUE ~ cyano_binary_decrease
)) |>
       mutate(combined_cyano_response_binary_increasepriority = case_when(
  prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
  prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
  TRUE ~ cyano_binary_increase
)) |> 
   mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
  mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
  mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))

```

## Set up logistic regression

```{r}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary_increasepriority, 
         response = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "increase")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary_increasepriority, 
         response = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")|> 
  mutate(wl_priority = "increase")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_increase <- rbind(cyanodata, phytodata)
#"#7A524B","#3C7F29","#8DBAE9"

cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_decrease <- rbind(cyanodata, phytodata)
logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"

```

## Logreg: combined
```{r}
logregdata_increase1 <- logregdata_increase |>
  rename(prop_delta_wl = prop_delta_wl_priorityincrease)

new_all <- bind_rows(logregdata_decrease, logregdata_increase1) |>
  select(-c(wl_priority)) |>
  unique()

all <- new_all %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#4dd0e1", "#99bbab"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) +
  #geom_jitter()+
 ylab("Response to changes in water level") +
  xlab("Proportion water level change") +
  #ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10),
        axis.title.y=element_text(angle=90, vjust = .5, size = 12),
        legend.title=element_blank())
all
#117733
#factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic"))
logreg_phyto <- new_all %>% 
  filter(type == "phyto") |> 
  # ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  ggplot(aes(x = prop_delta_wl, y = response_binary, shape = as.factor(type))) +
  scale_color_manual(values = c("#3355DB", "#FFC51E", "#B1C3A2", "#549C79", "gray"), name = "")+
  scale_shape_manual(values = c(20, 10, 11, 12, 13)) +
  # geom_point(size = 10, alpha = 0.5, aes(color = as.factor(trophic_status_reduced))) +
    geom_point(
    aes(
      color = as.factor(factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic", "not reported"))),
      shape = as.factor(type),
      text = paste(
        "Name / Year:", name_yr,
        "<br>Trophic status:", trophic_status_reduced,
        "<br>Δ WL:", prop_delta_wl,
        "<br>Response:", response_binary
      )
    ),
    size = 14,
    alpha = 0.75
  ) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 6, color = "black") +
  #geom_jitter()+
 # ylab("Phytoplankton response to changes in water level") +
 #  xlab("Proportion water level change") +
   ylab("") +
  xlab("") +
  #ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 0.5, size = 20),
        axis.text.y = element_text(hjust = 1, size = 20),
        axis.title.y=element_text(angle=90, hjust = 1, size = 28),
        axis.title.x=element_text(vjust = 1, size = 28),
        legend.text  = element_text(size = 28),
        legend.title=element_blank())

plotly_p <- ggplotly(logreg_phyto)

plotly_p # I have this as a plotly to be able to see exactly what's going into this for now. Useful, but can probably be removed for publishing
logreg_phyto

logregdata_phyto <- new_all %>% subset(type == "phyto")
logregdata_cyano <- new_all %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
results <- summary(logregform_phyto)
exp(coef(results))

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/logregposter.jpg", plot = logreg_phyto, height = 9.5, width = 20, units = "in")

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/logreg.jpg", plot = all, height = 4, width = 8, units = "in")

```

## Set up and run X2

```{r}
dat <- boxplotdata2 %>% 
    mutate(depth_m = ifelse(is.na(fullpond_maxdepth), obs_maxdepth, fullpond_maxdepth),
           depth_m = as.numeric(depth_m)) |> 
  select(cov_num, res_name, trophic_status_combined, tp_ugL, decrease_phyto_response, increase_phyto_response,
         decrease_cyano_response, increase_cyano_response, chla_ugL, combined_phyto_response, combined_cyano_response, name_yr, depth_m) %>%
  mutate(increase_phyto_response = ifelse(increase_phyto_response == "Not measured"|increase_phyto_response == ""|(grepl("Other", increase_phyto_response)), NA, increase_phyto_response)) |>
    mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured"|decrease_phyto_response == ""|(grepl("Other", decrease_phyto_response)), NA, decrease_phyto_response)) |>
   mutate(increase_cyano_response = ifelse(increase_cyano_response == "Not measured"|increase_cyano_response == ""|(grepl("Other", increase_cyano_response)), NA, increase_cyano_response)) |>
   mutate(decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured"|decrease_cyano_response == ""|(grepl("Other", decrease_cyano_response)), NA, decrease_cyano_response)) |>
  filter(!(is.na(decrease_phyto_response) & is.na(increase_phyto_response) & is.na(decrease_cyano_response) & is.na(increase_cyano_response))) %>%
  mutate(increase_decrease_mosaic = ifelse(((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)) & (!is.na(increase_phyto_response) | !is.na(increase_cyano_response))),"Both",
                                           ifelse((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)),"Decrease",
                                                  ifelse((!is.na(increase_phyto_response) | !is.na(increase_cyano_response)),"Increase", NA)))) 

fluctuation_studies_keep_increase <- dat %>%
  filter(increase_decrease_mosaic == "Both") %>%
  mutate(increase_decrease_mosaic = "Increase",
         decrease_phyto_response = NA,
         decrease_cyano_response = NA) 

fluctuation_studies_keep_decrease <- dat %>%
  filter(increase_decrease_mosaic == "Both") %>%
  mutate(increase_decrease_mosaic = "Decrease",
         increase_phyto_response = NA,
         increase_cyano_response = NA) 

dat_no_fluctuation_studies <- dat %>%
  filter(!increase_decrease_mosaic == "Both")

dat2 <- bind_rows(dat_no_fluctuation_studies, fluctuation_studies_keep_increase) %>%
  bind_rows(., fluctuation_studies_keep_decrease) %>%
  mutate(decrease_phyto_response = ifelse(is.na(decrease_phyto_response),"not reported",decrease_phyto_response),
         increase_phyto_response = ifelse(is.na(increase_phyto_response),"not reported",increase_phyto_response),
         decrease_cyano_response = ifelse(is.na(decrease_cyano_response),"not reported",decrease_cyano_response),
         increase_cyano_response = ifelse(is.na(increase_cyano_response),"not reported",increase_cyano_response)) %>%
  mutate(trophic_status_mosaic = ifelse((grepl("oligo",trophic_status_combined) | grepl("meso",trophic_status_combined)),"oligo/mesotrophic",
                                        ifelse(grepl("eu", trophic_status_combined),"eu/hypereutrophic","not reported")),
         increase_phyto = ifelse((increase_decrease_mosaic == "Increase" & increase_phyto_response == "Increase in phytoplankton"),"yes",
                                 ifelse(increase_decrease_mosaic == "Increase" & increase_phyto_response == "not reported","not reported",
                                        ifelse(increase_decrease_mosaic == "Decrease" & decrease_phyto_response == "Increase in phytoplankton","yes",
                                               ifelse(increase_decrease_mosaic == "Decrease" & decrease_phyto_response == "not reported","not reported","no")))),
         increase_cyano = ifelse((increase_decrease_mosaic == "Increase" & increase_cyano_response == "Increase in cyanobacteria"),"yes",
                                 ifelse(increase_decrease_mosaic == "Increase" & increase_cyano_response == "not reported","not reported",
                                        ifelse(increase_decrease_mosaic == "Decrease" & decrease_cyano_response == "Increase in cyanobacteria","yes",
                                               ifelse(increase_decrease_mosaic == "Decrease" & decrease_cyano_response == "not reported","not reported","no"))))) %>%
  select(cov_num, name_yr, depth_m, res_name, increase_decrease_mosaic, trophic_status_mosaic, increase_phyto, increase_cyano) %>%
  mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("not reported","eu/hypereutrophic","oligo/mesotrophic")),
         increase_phyto = factor(increase_phyto, levels = c("yes","no","not reported")),
         increase_cyano = factor(increase_cyano, levels = c("yes","no","not reported"))) 

dat_phyto1 <- dat2 %>%
  filter(!increase_phyto == "not reported") %>%
  select(trophic_status_mosaic, increase_decrease_mosaic, increase_phyto) %>%
  mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_phyto, sep = "_")) %>%
  group_by(x_axis_barplot, trophic_status_mosaic) %>%
  summarise(count = n()) |>
  mutate(response.count = sum(count),
         ts.prop = count/sum(count)) |>
  ungroup() %>%
  mutate(response.prop = response.count/sum(count)) %>%
  ungroup() %>%
  separate(x_axis_barplot,c("increase_decrease_mosaic","increase_phyto"), remove = FALSE) %>%
  mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("oligo/mesotrophic","eu/hypereutrophic","not reported")))

dat_phyto1 <- dat2 %>%
  filter(!increase_phyto == "not reported") %>%
  select(trophic_status_mosaic, increase_decrease_mosaic, increase_phyto) %>%
  mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_phyto, sep = "_")) %>%
  group_by(x_axis_barplot, trophic_status_mosaic) %>%
  summarise(count = n(), .groups = "drop") %>%

  # Ensure all combinations of x_axis_barplot and trophic_status_mosaic exist
  complete(
    x_axis_barplot,
    trophic_status_mosaic,
    fill = list(count = 0)
  ) %>%

  # Recompute decomposed vars after completing
  separate(x_axis_barplot, c("increase_decrease_mosaic", "increase_phyto"), remove = FALSE) %>%
  group_by(x_axis_barplot) %>%
  mutate(response.count = sum(count),
         ts.prop = count / sum(count)) %>%
  ungroup() %>%
  mutate(response.prop = response.count / sum(count)) %>%
  mutate(
    trophic_status_mosaic = factor(
      trophic_status_mosaic,
      levels = c("oligo/mesotrophic", "eu/hypereutrophic", "not reported")
    )
  )


dat_phyto_new <- dat2 %>%
  filter(!increase_phyto == "not reported") %>%
  select(increase_decrease_mosaic, increase_phyto) %>%
  mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_phyto, sep = "_")) %>%
  group_by(x_axis_barplot, increase_decrease_mosaic) %>%
  summarise(count = n()) |>
  ungroup() |> 
  group_by(increase_decrease_mosaic) |> 
  mutate(response.count = sum(count),
         ts.prop = count/sum(count)) |>
  ungroup() %>%
  mutate(response.prop = count/response.count) %>%
  ungroup() %>%
  separate(x_axis_barplot, c("increase_decrease_mosaic","increase_phyto"), remove = FALSE)

dat_cyano_new <- dat2 %>%
  filter(!increase_cyano == "not reported") %>%
  select(increase_decrease_mosaic, increase_cyano) %>%
  mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_cyano, sep = "_")) %>%
  group_by(x_axis_barplot, increase_decrease_mosaic) %>%
  summarise(count = n()) |>
  ungroup() |> 
  group_by(increase_decrease_mosaic) |> 
  mutate(response.count = sum(count),
         ts.prop = count/sum(count)) |>
  ungroup() %>%
  mutate(response.prop = count/response.count) %>%
  ungroup() %>%
  separate(x_axis_barplot, c("increase_decrease_mosaic","increase_cyano"), remove = FALSE)

#for datphytonew:
# New facet label names for water level/response variable
wl.labs <- c("WL decrease","WL increase")
names(wl.labs) <- unique(dat_phyto_new$increase_decrease_mosaic)
response.labs <- c("no phyto \n increase", "phyto \n increase", "no phyto \n increase", "phyto \n increase") 

# Calculate bar positions
w <- c((dat_phyto_new$response.prop))
pos <- data.frame(pos = 0.5 * (cumsum(w) + cumsum(c(0, w[-length(w)]))),
                  x_axis_barplot = unique(dat_phyto_new$x_axis_barplot))

## From Agresti(2007) p.39

# first for phytoplankton
phyto <- dat2 %>%
  select(increase_decrease_mosaic, increase_phyto)
phyto_increase_wl_increase <- phyto %>%
  filter(increase_decrease_mosaic == "Increase" & increase_phyto == "yes") %>%
  count(increase_phyto) %>%
  pull(n)
phyto_no_increase_wl_increase <- phyto %>%
  filter(increase_decrease_mosaic == "Increase" & increase_phyto == "no") %>%
  count(increase_phyto) %>%
  pull(n)
phyto_increase_wl_decrease <- phyto %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_phyto == "yes") %>%
  count(increase_phyto) %>%
  pull(n)
phyto_no_increase_wl_decrease <- phyto %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_phyto == "no") %>%
  count(increase_phyto) %>%
  pull(n)

# Question 1: are there differences in phytoplankton responses to increases vs.
# decreases in water level?
M <- as.table(rbind(c(phyto_increase_wl_increase, phyto_increase_wl_decrease), c(phyto_no_increase_wl_increase, phyto_no_increase_wl_decrease)))
dimnames(M) <- list(phyto_response = c("increase", "decrease or no change"),
                    water_level = c("increase","decrease"))
M
(Xsq <- chisq.test(M))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 2: are there differences in the likelihood of phytoplankton to increase
# or decrease given an increase in water level?
M_increase <- M[,1]
M_increase
(Xsq <- chisq.test(M_increase))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 3: are there differences in the likelihood of phytoplankton to increase
# or decrease given a decrease in water level?
M_decrease <- M[,2]
M_decrease
(Xsq <- chisq.test(M_decrease))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# now for cyanobacteria
cyano <- dat2 %>%
  select(increase_decrease_mosaic, increase_cyano)
cyano_increase_wl_increase <- cyano %>%
  filter(increase_decrease_mosaic == "Increase" & increase_cyano == "yes") %>%
  count(increase_cyano) %>%
  pull(n)
cyano_no_increase_wl_increase <- cyano %>%
  filter(increase_decrease_mosaic == "Increase" & increase_cyano == "no") %>%
  count(increase_cyano) %>%
  pull(n)
cyano_increase_wl_decrease <- cyano %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_cyano == "yes") %>%
  count(increase_cyano) %>%
  pull(n)
cyano_no_increase_wl_decrease <- cyano %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_cyano == "no") %>%
  count(increase_cyano) %>%
  pull(n)

# Question 1: are there differences in cyanobacterial responses to increases vs.
# decreases in water level?
M <- as.table(rbind(c(cyano_increase_wl_increase, cyano_increase_wl_decrease), c(cyano_no_increase_wl_increase, cyano_no_increase_wl_decrease)))
dimnames(M) <- list(cyano_response = c("increase", "decrease or no change"),
                    water_level = c("increase","decrease"))
M
f_chisq_test(M)
(Xsq <- chisq.test(M))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 2: are there differences in the likelihood of cyanobacteria to increase
# or decrease given an increase in water level?
M_increase <- M[,1]
M_increase
(Xsq <- chisq.test(M_increase))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 3: are there differences in the likelihood of cyanobacteria to increase
# or decrease given a decrease in water level?
M_decrease <- M[,2]
M_decrease
(Xsq <- chisq.test(M_decrease))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value


```

```{r x2 with three levels instead of 2}
library(dplyr)
library(tidyr)
library(rstatix)

# Create contingency table
tab_cyano_all <- boxplotdata2 %>%
  pivot_longer(
    cols = c(cyano_response_wl_up, cyano_response_wl_down),
    names_to = "water_direction",
    values_to = "response"
  ) %>%
  mutate(response = gsub(" ", "", response)) %>%
  filter(!is.na(response)) %>%
  with(table(water_direction, response))

# Run chi-square test
# Are there differences in cyanobacterial responses to increases vs. decreases in water level?
chisq_test(tab_cyano_all)
f_chisq_test(tab_cyano_all) #Bonferroni - adj. p

tab_cyano_down <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(cyano_response_wl_down) ~ "wl_down",
      is.na(cyano_response_wl_down) & !is.na(cyano_response_wl_up) ~ "wl_up",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(cyano_response_wl_down) ~ cyano_response_wl_down,
      is.na(cyano_response_wl_down) & !is.na(cyano_response_wl_up) ~ cyano_response_wl_up,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)

tab_cyano_down

tab_cyano_down <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_cyano_down)

chisq_test(tab_cyano_down)
f_chisq_test(tab_cyano_down)
fisher.test(tab_cyano_down)


tab_cyano_up <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(cyano_response_wl_up) ~ "wl_up",                     # priority
      is.na(cyano_response_wl_up) & !is.na(cyano_response_wl_down) ~ "wl_down",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(cyano_response_wl_up) ~ cyano_response_wl_up,        # priority
      is.na(cyano_response_wl_up) & !is.na(cyano_response_wl_down) ~ cyano_response_wl_down,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)


tab_cyano_up <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_cyano_up)
chi_cyano_up <- chisq.test(tab_cyano_up)
f_chisq_test(tab_cyano_up)


# Create contingency table
tab_phyto_all <- boxplotdata2 %>%
  pivot_longer(
    cols = c(phyto_response_wl_up, phyto_response_wl_down),
    names_to = "water_direction",
    values_to = "response"
  ) %>%
  mutate(response = gsub(" ", "", response)) %>%
  filter(!is.na(response)) %>%
  with(table(water_direction, response))

# Run chi-square test
# Are there differences in cyanobacterial responses to increases vs. decreases in water level?
chisq_test(tab_phyto_all)
f_chisq_test(tab_phyto_all) #Bonferroni - adj. p

#wl decrease priority
tab_phyto <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(phyto_response_wl_down) ~ "wl_down",
      is.na(phyto_response_wl_down) & !is.na(phyto_response_wl_up) ~ "wl_up",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(phyto_response_wl_down) ~ phyto_response_wl_down,
      is.na(phyto_response_wl_down) & !is.na(phyto_response_wl_up) ~ phyto_response_wl_up,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)
tab_phyto
tab_phyto_decrease <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_phyto)
tab_phyto_decrease

chisq.test(tab_phyto_decrease)
f_chisq_test(tab_phyto_decrease)

fisher.test(tab_cyano_down)
fisher.test(tab_cyano_up)
fisher.test(tab_phyto_decrease)
fisher.test(phyto_increase_chi)

tab_phyto_increase <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(phyto_response_wl_up) ~ "wl_up",                        # priority to wl_up
      is.na(phyto_response_wl_up) & !is.na(phyto_response_wl_down) ~ "wl_down",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(phyto_response_wl_up) ~ phyto_response_wl_up,           # priority to wl_up
      is.na(phyto_response_wl_up) & !is.na(phyto_response_wl_down) ~ phyto_response_wl_down,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)
tab_phyto_increase
phyto_increase_chi <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_phyto_increase)
f_chisq_test(phyto_increase_chi)

cyano_chi <- ggplot(df_plot_cyano, aes(x = Var1, y = Freq, fill = Var2)) +
  geom_bar(stat = "identity", position = "fill") +
     geom_text(aes(label = Freq),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 6) +
  labs(x = "",
       y = "Proportion",
       fill = "Response",
       title = "Cyanobacteria Response by Water Level Direction") +
 scale_fill_manual(
  values = c("#b2ebf2", "#4dd0e1", "#e0e0e0"),   # 3 colors
  name = "Did \ncyanobacteria\nincrease or \ndecrease?",
  labels = c("Decrease", "Increase", "Not Significant")) +
  theme_bw() + 
    theme(axis.text.x = element_text(hjust = 0.5, size = 12),
        #axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=90, vjust = 0.5, size = 12),
        axis.title.x=element_text(vjust = 1, size = 12),
        #legend.text  = element_blank(),
        legend.title=element_text(size = 12), 
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm')) 

df_plot_phyto <- as.data.frame(tab_phyto)

df_plot_phyto$Var1 <- factor(df_plot_phyto$Var1,
                       levels = c("phyto_response_wl_down",
                                  "phyto_response_wl_up"),
                       labels = c("Water Level Decrease",
                                  "Water Level Increase"))


phyto_chi <- ggplot(df_plot_phyto, aes(x = Var1, y = Freq, fill = Var2)) +
  geom_bar(stat = "identity", position = "fill") +
     geom_text(aes(label = Freq),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 6) +
  labs(x = "",
       y = "Proportion",
       fill = "Response",
       title = "Phytoplankton Response by Water Level Direction") +
  scale_fill_manual(values = c("#ddefcd", "#99bbab", "#e0e0e0"), 
                    name = "Did \nphytoplankton\nincrease or \ndecrease?",
  labels = c("Decrease", "Increase", "Not Significant")) +
  theme_bw() + 
    theme(axis.text.x = element_text(hjust = 0.5, size = 12),
        #axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=90, vjust = 0.5, size = 12),
        axis.title.x=element_text(vjust = 1, size = 12),
        #legend.text  = element_blank(),
        legend.title=element_text(size = 12), 
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm')) 

chi <- phyto_chi / cyano_chi #+ plot_annotation(title = "Chi-Square") 

chi

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/chi.jpg", plot = chi, height = 5, width = 6.5, units = "in")

```


## X2 plots

Possible Colors for chi-sq plots: "#8fc7c3", \# cyan-teal version of #99bbab "#d9f1ec" "#4dd0e1", "#b2ebf2" "#ddefcd" "#05A3CA", "#5dce97", "#608687", "#99bbab", "#ddefcd", "#80407c", "#45Cc80", "#a2eafc", "#99abbb""#9d99bb", "#cfcdef" "#99b3c8", \# blue version of #99bbab "#dde7f2") c("#8fb3d9", "#d6e4f5")
```{r}
# "#8fc7c3", 
# # cyan-teal version of 
# "#99bbab" "#d9f1ec" "#4dd0e1", "#b2ebf2" "#ddefcd" "#05A3CA", "#5dce97", "#608687", "#99bbab", "#ddefcd", "#80407c", "#45Cc80", "#a2eafc", "#99abbb""#9d99bb", "#cfcdef" "#99b3c8", \# blue version of #99bbab "#dde7f2") c("#8fb3d9", "#d6e4f5")
```



```{r barplots for chisq}
# #stacked barplots for chisq
# dat_phyto_new1 <- dat_phyto_new |> 
#   mutate(increase_decrease_mosaic = case_when(increase_decrease_mosaic == "Decrease" ~ "Decrease in Water Level", 
#                                               increase_decrease_mosaic == "Increase" ~ "Increase in Water Level"))
# 
# dat_cyano_new1 <- dat_cyano_new |> 
#   mutate(increase_decrease_mosaic = case_when(increase_decrease_mosaic == "Decrease" ~ "Decrease in Water Level", 
#                                               increase_decrease_mosaic == "Increase" ~ "Increase in Water Level"))
#  
# 
# 
# phyto_chi <-  ggplot(dat_phyto_new1, aes(x = increase_decrease_mosaic, y = count, fill = increase_phyto))  +
#     geom_bar(position="stack", stat="identity") + 
#     xlab("") + ylab("") + 
#    geom_text(aes(label = count),
#             position = position_stack(vjust = 0.5), 
#             color = "black", 
#             size = 6) + 
#   scale_fill_manual(values = c("#ddefcd", "#99bbab"), name = "Did phytoplankton \nincrease?") + 
#   theme_bw()+
#   theme(#axis.text.x = element_text(hjust = 0.5, size = 20),
#         axis.text.x = element_blank(),
#         axis.text.y = element_text(hjust = 1, size = 12),
#         axis.title.y=element_text(angle=90, hjust = 1, size = 12),
#         axis.title.x=element_text(vjust = 1, size = 12),
#         legend.text  = element_blank(),
#         legend.title=element_text(size = 12), 
#         legend.key.height = unit(1, 'cm'),
#         legend.key.width = unit(1, 'cm')) 
#   
# cyano_chi <- ggplot(dat_cyano_new1, aes(x = increase_decrease_mosaic, y = count, fill = increase_cyano)) + 
#     geom_bar(position="stack", stat="identity")+ 
#     xlab("") + ylab("") +
#    geom_text(aes(label = count),
#             position = position_stack(vjust = 0.5), 
#             color = "black", 
#             size = 6) + 
# 
#     scale_fill_manual(values = c("#b2ebf2", "#4dd0e1"), name = "Did cyanobacteria \nincrease?")+ 
#   theme_bw() + 
#     theme(axis.text.x = element_text(hjust = 0.5, size = 12),
#         #axis.text.x = element_blank(),
#         axis.text.y = element_text(hjust = 1, size = 12),
#         axis.title.y=element_text(angle=90, hjust = 1, size = 12),
#         axis.title.x=element_text(vjust = 1, size = 12),
#         legend.text  = element_blank(),
#         legend.title=element_text(size = 12), 
#         legend.key.height = unit(1, 'cm'),
#         legend.key.width = unit(1, 'cm')) 
# 
# chi <- phyto_chi / cyano_chi #+ plot_annotation(title = "Chi-Square") 
# 
# chi
# 
# ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/chi.jpg", plot = chi, height = 5, width = 6.5, units = "in")
```

# Supplemental Figures
logistic regression to prioritize increases or decreases. 
```{r}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary_increasepriority, 
         response = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "increase")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary_increasepriority, 
         response = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")|> 
  mutate(wl_priority = "increase")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_increase <- rbind(cyanodata, phytodata)
#"#7A524B","#3C7F29","#8DBAE9"

test <- logregdata_increase %>% ggplot(aes(x = prop_delta_wl_priorityincrease, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#4dd0e1", "#99bbab"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes in water level") + 
  xlab("Proportion water level change") + 
  ggtitle("Priority WL increases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
test

logregdata_phyto <- logregdata_increase %>% subset(type == "phyto")
logregdata_cyano <- logregdata_increase %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
```

## Logreg: decreases

```{r wl decreases}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_decrease <- rbind(cyanodata, phytodata)
logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"

test <- logregdata %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes in water level") + 
  xlab("Proportion water level change") + 
  ggtitle("Priority WL decrease")+
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
test

logregdata_phyto <- logregdata %>% subset(type == "phyto")
logregdata_cyano <- logregdata %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
```



```{r possible supplement figures}

#countries by continent
countries_sf <- data |> 
  select(lat_dd, long_dd, res_name, cov_num, name_yr) |> 
  filter(!is.na(lat_dd) & !is.na(long_dd)) |> 
  st_as_sf(coords = c("long_dd", "lat_dd"), crs = 4326) 
world <- ne_countries(scale = "medium", returnclass = "sf")
df_with_country <- st_join(countries_sf, world["name_long"])
df_final <- df_with_country %>% 
  st_drop_geometry() %>%
  rename(country = name_long)
df_final <- df_final %>% 
  mutate(continent = world$continent[match(country, world$name_long)])

countries <- left_join(data, df_final, by = c("res_name", "cov_num", "name_yr")) |> select(c("res_name", "name_yr", "country.x", "country.y", "continent")) |> 
  mutate(country.x = ifelse(country.x == "", NA, country.x)) |> 
  mutate(country = ifelse(!is.na(country.x), country.x, country.y)) |> 
  mutate(continent.a = case_when(
    country %in% c("Brazil", "Argentina", "Uruguay", "Argentina and Uruguay") ~ "South America",
      country %in% c("United States", "USA/Canada", "Canada", "Mexico") ~ "North America",
      country %in% c("Portugal", "England", "Italy", "France", "Serbia", 
                     "Estonia", "Estonia/Russia") ~ "Europe",
      country %in% c("Russia") ~ "Europe/Asia",
      country %in% c("China", "Taiwan", "Malaysia", "Mongolia", "Lebanon",
                     "Cambodia", "Vietnam") ~ "Asia",
       country %in% c("Australia") ~ "Oceania",
    country %in% c("South Africa") ~ "Africa",
      TRUE ~ NA_character_)) |> 
  mutate(continent = ifelse(!is.na(continent.a), continent.a, continent)) |> 
  select(res_name, country, continent) |> 
  unique()

countries |> 
  filter(!is.na(continent)) |> 
  ggplot(aes(x=continent))+
  geom_bar() + 
  xlab("Continent") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

#histogram of maxdepths

data |> 
  subset(!is.na(fullpond_maxdepth)) %>% 
  ggplot(aes(x=fullpond_maxdepth)) + 
  geom_histogram() +
  xlab("Max depth at Full Pond") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Old mapping figures

```{r old maps}
latlong %>% 
  subset(!is.na(long_dd)) %>% 
  subset(!is.na(lat_dd)) |> 
  mapview(xcol = "long_dd", ycol = "lat_dd", crs = 4269, grid = FALSE)

latlong |> 
  filter(lat_dd > 80)

litmap <- latlong %>% 
  subset(!is.na(long_dd)) %>% 
  mapview(xcol = "long_dd", ycol = "lat_dd", crs = 4269, grid = FALSE)

world <- ne_countries(scale = "medium", returnclass = "sf")
#world <- map_data("world")
my_countries <- c("Russia", "South Korea", "China", "Brazil", "Spain", "United States of America", "Burkina Faso", "Argentina", "Lebanon", "Vietnam", "Canada", "Israel", "Croatia", "Netherlands", "Estonia", "Italy", "Turkey", "Germany")

new_world <- mutate(world, fill = ifelse(sovereignt %in% c(my_countries), "yes", "no"))

litdata_map <- latlong %>% 
  mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured", NA, decrease_phyto_response), 
         increase_phyto_response = ifelse(increase_phyto_response == "Not measured", NA, increase_phyto_response), 
         decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured", NA, decrease_cyano_response), 
         increase_cyano_response = ifelse(increase_cyano_response == "Not measured", NA, increase_cyano_response)) |> 
   mutate(mapvals = ifelse(((!is.na(decrease_phyto_response)|!is.na(increase_phyto_response)) & (!is.na(decrease_cyano_response)|!is.na(increase_cyano_response))), "Both",
                           ifelse(!is.na(decrease_phyto_response)|!is.na(increase_phyto_response), "Phytoplankton", 
                                 ifelse(!is.na(decrease_cyano_response)|!is.na(increase_cyano_response), "Cyanobacteria", NA)))) 

litdata_map_sf <- st_as_sf(litdata_map,
                           coords = c("long_dd", "lat_dd"),
                           crs = 4326,
                           remove = FALSE)

new_world <- st_transform(new_world, 4326)

sitemap <- new_world %>% ggplot(fill = fill) +
    geom_sf(data = new_world, aes(fill = factor(fill))) +
 # coord_sf(crs = 4269) +
  scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries")+
    geom_point(data = litdata_map_sf, aes(x = long_dd, y = lat_dd, shape = as.factor(mapvals), color = as.factor(mapvals)), size = 2) + 
  scale_color_manual(values = c("black", "black", "#CC6677", "black", "black"), name = "Variable") +
  scale_shape_manual(values = c(16, 16, 18, 16, 16), name = "Variable")+
  ylab("Latitude") +
  xlab("Longitude") +
    theme(text = element_text(size = 12),
        axis.text.x = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=0, vjust = .5))+
  theme_bw() +
    theme(text = element_text(size = 12),
        axis.text.x = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=90, vjust = .5),
        legend.position="none")
library(scales)


# ggplot(data = world, aes(x = long, y = lat, group = group), fill = "white") + 
#   geom_polygon(fill = "white", color = "black")+
#   geom_sf(data = litdata_map_sf,
#           aes(shape = as.factor(mapvals),
#               color = as.factor(mapvals)),
#           size = 2)

sitemap <- ggplot(world) +
  geom_sf(fill = "white") +
 # coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white", color = "black") +
  #geom_sf_label(stat = "sf_coordinates")+
  geom_sf(data = litdata_map_sf,
          aes(shape = as.factor(mapvals),
              color = as.factor(mapvals)),
          size = 2) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("#608687", "#4dd0e1", "darkgreen"),
                     name = "Variable") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  coord_sf() +
  theme_minimal()
sitemap

sitemap_depth <- ggplot(world) +
  geom_sf(data = world, fill = "white") +
  #coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white") +
  #geom_sf_label(stat = "sf_coordinates")+
  geom_sf(data = litdata_map_sf,
          aes(shape = as.factor(mapvals),
              color = as.factor(mapvals), 
              size = (depth_m))) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("#608687", "#4dd0e1", "darkgreen"),
                     name = "Variable") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  theme_minimal()

sitemap_depth

sitemap_trophic <- ggplot(world) +
  geom_sf(data = world, fill = "white") +
  #coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white") +
  #geom_sf_label(stat = "sf_coordinates")+
  # geom_sf(data = litdata_map_sf,
  #         aes(shape = as.factor(mapvals),
  #             color = as.factor(increase_decrease_mosaic))) +
    geom_sf(data = litdata_map_sf, size = 3, alpha = 0.6,
          aes(color = as.factor(increase_decrease_mosaic))) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("lightgreen", "darkgreen", "blue", "blue"),
                     name = "Was it a \nwater level \nincrease or \ndecrease? \n") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  theme_minimal() + 
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(legend.text  = element_text(size = 16),
        legend.title = element_text(size = 20))
sitemap_trophic
# ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/sitemap.jpg", plot = sitemap, height = 4, width = 6.5, units = "in")

ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/sitemapposter.jpg", plot = sitemap_trophic, height = 4, width = 10, units = "in")
```

Old code that includes all vars for drivers
```{r}
#separate out phytoplankton and cyanobacteria drivers
phyto_only_drivers <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:125)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_drivers <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:125)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

#WL change vars
phyto_only_WLchangevars <- WLchangevars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_WLchangevars <- WLchangevars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")

#not WL change vars!
phyto_only_vars <- spreadvars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:124)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured" | decrease_phyto_response != "" & decrease_phyto_response != "Not measured")

cyano_only_vars <- spreadvars |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:124)) |> 
  filter(increase_cyano_response != "" & increase_cyano_response != "Not measured" | decrease_cyano_response != "" & decrease_cyano_response != "Not measured")


#many regerts
chemvars_drive_phytos <- phyto_only_drivers |> select(c(other_chem, tp, tn, nh4, ph, do, srp, no3, cond, metals, orp, no2, doc, din, ph, salinity, dic, tkn, dom, pom))

#only including variables that were measured at least 10 times in the extracted data
chemvars_phytos <- phyto_only_vars |> select(c(other_chem, tp, tn, nh4, ph, do, srp, no3, cond, metals, orp, no2, doc, din, ph, salinity, dic, tkn, dom, pom))

physvars_drive_phytos <- phyto_only_drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))

# physvarsWL_phytos <- phyto_only_WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp))

physvars_phytos <- phyto_only_vars |> select(c(water_temp, water_level, transparency, turb, light, tss_tds, other_phys))
# summary(vars) 

freqchem_drive_phytos <- as.data.frame(colSums(chemvars_drive_phytos))
chemvars_drive_phytos <- setNames(cbind(rownames(freqchem_drive_phytos), freqchem_drive_phytos, row.names = NULL), 
         c("names", "sum"))
# freqchemWL_phyto <- as.data.frame(colSums(chemvarsWL_phytos))
freqchem_phyto <- as.data.frame(colSums(chemvars_phytos))
# chemvarsWL_phytos <- setNames(cbind(rownames(freqchemWL), freqchemWL_phyto, row.names = NULL), 
#          c("names", "sum"))
chemvars_phytos <- setNames(cbind(rownames(freqchem_phyto), 
                                  freqchem_phyto, row.names = NULL), c("names", "sum"))

# chemvarsWL |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +  ggtitle("Chem variables that changed in response to WL")
# want a comparison with what was measured!
chemvars1_phytos <- chemvars_phytos |> 
  rename(measured = sum, 
         name = names)
chemicalvars_phytos <- cbind(chemvars1_phytos, chemvars_drive_phytos)

chem <- chemicalvars_phytos |> 
  filter(!name == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names))+
  geom_bar(position = "fill", stat = "identity") + 
  ggtitle("Chem variables, phytos only" ) + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))+ theme_bw()

#physical variables

freqphys_drive <- as.data.frame(colSums(physvars_drive_phytos))
# freqphysWL <- as.data.frame(colSums(physvarsWL_phytos))
freqphys_phytos <- as.data.frame(colSums(physvars_phytos))

physvars2_drive_phytos <- setNames(cbind(rownames(freqphys_drive), freqphys_drive, row.names = NULL), 
         c("names", "sum"))

physvars1_phytos <- setNames(cbind(rownames(freqphys_phytos), freqphys_phytos, row.names = NULL), 
         c("name", "measured"))

# physvars2WL <- setNames(cbind(rownames(freqphysWL), freqphysWL, row.names = NULL), 
#          c("names", "sum"))

# physvars2_drive_phytos |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + ggtitle("Phys variable Drivers")

# physvars2WL |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
#   ggtitle("Phys variable that change with water level")


# physvars1 <- freqphys_phytos |> 
#   rename(measured = sum, 
#          name = names)

physvars_phytos <- cbind(physvars1_phytos, physvars2_drive_phytos)

phys <- physvars_phytos |> 
    select(c(names, measured, sum)) |> 
  filter(!names == "other_phys") |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
  mutate(notid = ifelse(notid <0, 0, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = "fill", stat = "identity") +
  # theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  theme(axis.text.x = element_text(vjust = 0.5, hjust=1)) + 
  ggtitle("Phys variables, phyto only") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))

phytos <- phys/chem
phytos
ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/phytovars.jpg", plot = phytos, height = 5, width = 6.5, units = "in")
```

Code for variables that changed with water level change
```{r}
chemvars1 <- chemvars |> 
  rename(measured = sum, 
         name = names)
chemicalvars <- cbind(chemvars1, chemvars_drive)
chemvarsplot <- chemicalvars |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
   filter(!names == c("other_chem")) |>
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names))+
  geom_bar(position = "fill", stat = "identity") + 
  ggtitle("Proportion Chemical variables that were identified as drivers" ) + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))

chemicalvars |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names))+
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))

physvars1 <- physvars |> 
  rename(measured = sum, 
         name = names)

physvars <- cbind(physvars1, physvars2_drive)

physvars |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
  mutate(notid = ifelse(notid <0, 0, notid)) |> 
   filter(!names == c("other_phys")) |>
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = "stack", stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  ggtitle("Phys variables") + 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))

physvarsplot <- physvars |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
  mutate(notid = ifelse(notid <0, 0, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = "fill", stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  ggtitle("Phys variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3)  + 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))

chemvars1 <- chemvars |> 
  rename(measured = sum, 
         name = names)
chemicalvarsWL <- cbind(chemvars1, chemvarsWL)
WL_chem <- chemicalvarsWL |> 
  select(c(names, measured, sum)) |> 
  rename(change = sum) |> 
  mutate(nochange = measured - change) |> 
  select(c(names, change, nochange)) |> 
  filter(!names == c("other_chem")) |>
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names))+
  geom_bar(position = "fill", stat = "identity") + 
  ggtitle("Proportion Chemical variables that changed with WL change" ) + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3)+ 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))


# physvars1 <- physvars1 |> 
#   rename(measured = sum, 
#          name = names)

physvars_wl_spec <- physvars1 |> 
  filter(name != "water_level")
physicalvarsWL <- cbind(physvars_wl_spec, physvars2WL)

wl_phys <- physicalvarsWL |> 
  select(c(names, measured, sum)) |> 
  filter(!names == c("other_phys", "restime")) |> 
  rename(change = sum) |> 
  mutate(nochange = measured - change) |> 
  select(c(names, change, nochange)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names))+
  geom_bar(position = "fill", stat = "identity") + 
  ggtitle("Proportion physical variables that changed with WL change" ) + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#fba049","#d2d2fe"))
varsplot <- physvarsplot/chemvarsplot
deltawl <- wl_phys/WL_chem
(wl_phys + WL_chem) / (physvarsplot+chemvarsplot)
# chemicalvarsWL |> 
#   select(c(names, measured, sum)) |> 
#   rename(change = sum) |> 
#   mutate(nochange = measured - change) |> 
#   select(c(names, change, nochange)) |> 
#   pivot_longer(!names, names_to = "driver", values_to = "count") |> 
#   ggplot(aes(fill = driver, y = count, x = names))+
#   geom_bar(position = "stack", stat = "identity") + geom_text(aes(label = count),
#             position = position_stack(vjust = 0.5), 
#             color = "white", 
#             size = 3)
ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/varswlchange.jpg", plot = deltawl, height = 5, width = 6.5, units = "in")
ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/vars.jpg", plot = varsplot, height = 5, width = 6.5, units = "in")

```


# phyto drivers for WL decreases.
I don't think this actually adds much more info, so have abandoned this for now. Also, while we can distinguish the difference between increases and decreases in WL, so many studies had both an increase and a decrease that I don't think we can effectively differentiate. 
```{r WL change}
phyto_only_drivers_up <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
  filter(increase_phyto_response != "" & increase_phyto_response != "Not measured")

phyto_only_drivers_down <- drivers |> 
  select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
filter(!decrease_phyto_response %in% c("", "Not measured") &
   increase_phyto_response %in% c("", "Not measured"))

phyto_only_down <- phyto_only_vars |> 
  #select(c(1:3, increase_phyto_response, increase_cyano_response, decrease_phyto_response, decrease_cyano_response, tp, 90:120)) |> 
filter(!decrease_phyto_response %in% c("", "Not measured") &
   !increase_phyto_response %in% c("", "Not measured"))

chemvars_drive_phyto_down <- phyto_only_drivers_down |> select(c(cond, do, nh4, no2, no3, ph, srp, tn, tp))

chemvars_phyto_down <- phyto_only_down |> select(c(cond, do, nh4, no2, no3, ph, srp, tn, tp))

physvars_drive_phytos_down <- phyto_only_drivers_down |> select(c(water_level, water_temp, transparency, turb, light, tss_tds))

physvars_phytos_down <- phyto_only_down |> select(c(water_temp, transparency, turb, light, tss_tds, water_level))

#add plots for phytos down

freqchem_drive_phytos_down <- as.data.frame(colSums(chemvars_drive_phyto_down))
chemvars_drive_phytos_down <- setNames(cbind(rownames(freqchem_drive_phytos_down), freqchem_drive_phytos_down, row.names = NULL), 
         c("names", "sum"))
# freqchemWL_phyto <- as.data.frame(colSums(chemvarsWL_phytos))
freqchem_phyto_down <- as.data.frame(colSums(chemvars_phyto_down))
# chemvarsWL_phytos <- setNames(cbind(rownames(freqchemWL), freqchemWL_phyto, row.names = NULL), 
#          c("names", "sum"))
chemvars_phytos_down <- setNames(cbind(rownames(freqchem_phyto_down), 
                                  freqchem_phyto_down, row.names = NULL), c("names", "sum"))

# chemvarsWL |> ggplot(aes(x = names, y = sum)) + geom_bar(stat = "identity") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +  ggtitle("Chem variables that changed in response to WL")
# want a comparison with what was measured!
chemvars1_phytos_down <- chemvars_phytos_down |> 
  rename(measured = sum, 
         name = names)

# drivers identified out of all vars measured
chemicalvars_phytos_down <- cbind(chemvars1_phytos_down, chemvars_drive_phytos_down)

chem <- chemicalvars_phytos_down |> 
  filter(!name == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names))+
  geom_bar(position = "fill", stat = "identity") + 
  ggtitle("Chem variable drivers, phytoplankton only, WL decreases only" ) + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#99bbab", "#ddefcd"))+ theme_bw()

#physical variables

freqphys_drive_down <- as.data.frame(colSums(physvars_drive_phytos_down))
# freqphysWL <- as.data.frame(colSums(physvarsWL_phytos))
freqphys_phytos_down <- as.data.frame(colSums(physvars_phytos_down))

physvars2_drive_phytos <- setNames(cbind(rownames(freqphys_drive_down), freqphys_drive_down, row.names = NULL), 
         c("names", "sum"))

physvars1_phytos <- setNames(cbind(rownames(freqphys_phytos_down), freqphys_phytos_down, row.names = NULL), 
         c("name", "measured"))

physvars_phytos <- cbind(physvars1_phytos, physvars2_drive_phytos)

phys <- physvars_phytos |> 
    select(c(names, measured, sum)) |> 
  filter(!names == "other_phys") |> 
  rename(id = sum) |> 
  mutate(notid = measured - id) |> 
  select(c(names, id, notid)) |> 
  mutate(notid = ifelse(notid <0, 0, notid)) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = "fill", stat = "identity") +
  # theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  theme(axis.text.x = element_text(vjust = 0.5)) + 
  ggtitle("Phys variable drivers, phytoplankton only") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#99bbab", "#ddefcd"))

phytos <- phys/chem
phytos
ggsave("~/GitHubRepos/CareyLabVT/Powell/Systematic/Plots/phytovars.jpg", plot = phytos, height = 5, width = 6.5, units = "in")


```
