---
title: "SystematicReview"
format: html
---


libraries for the whole script!
```{r packages}
# recommend to add lines to install.packages if this script is going to be part
# of the published repo when the manuscript is submitted

library(janitor)
library(ggplot2)
library(tidyverse)
library(dplyr)

library(rnaturalearth)
library(rnaturalearthdata)
library(mapview)

library(naniar)

```


I preprocessed colnames for data. 

```{r read in data}
#data <- read.csv("~/GitHubRepos/CareyLabVT/Powell/Systematic/SRE_26Aug25_covonly.csv")
data1 <- read.csv("./Combined_08Oct.csv")
data1 <- clean_names(data1)
head(data1)
```

```{r exclude papers}
data <- data1 |> 
  filter(!name_yr %in% c("Beaver 2013","Jia 2022","Cao 2016","Wang 2022","Kisand 2004")) |> 
  filter(!title %in% c("Steady-state assemblages in a Mediterranean hypertrophic reservoir. The role of Microcystis ecomorphological variability in maintaining an apparent equilibrium"))
```

```{r fix some data}

data2 <- data |> 
  mutate(tp = mean_tp) |> 
   separate_wider_delim(mean_tp, " ", names = c("tp_ugL", "unit_tp"), too_few = "align_start") |> 
  mutate(mean_chla = ifelse(mean_chla == "0.07 to 34.62 ug/L", "8", mean_chla)) |> #change Rai 1978 to 8ug/L, actual range between 0 and 14 ug/L, origianl range from pheophyton
  mutate(mean_chla = ifelse(mean_chla == "15.7ug/L", "15.7 ug/L", mean_chla)) |> 
  mutate(chla = mean_chla) |> 
   separate_wider_delim(mean_chla, " ", names = c("chla_ugL", "unit_chla"), too_few = "align_start") #|> 
 # select(c(chla, chla_ugL, unit_chla, tp, tp_ugL, unit_tp))
  
 rm(data) 
  
data <- data2 |> 
  mutate(tp_ugL = as.numeric(tp_ugL)) |> 
  mutate(mean_tp = case_when(
    unit_tp == "mg/L" ~ tp_ugL * 1000, 
    unit_tp == "mg/m3" ~ tp_ugL / 1000, 
    TRUE ~ tp_ugL)) |> 
  # select(c(mean_tp, tp, tp_ugL, unit_tp)) |> 
   mutate(chla_ugL = as.numeric(chla_ugL)) |> 
  mutate(mean_chla = case_when(
    unit_chla == "mg/L" ~ chla_ugL * 1000, 
    unit_chla == "mg/m3" ~ chla_ugL / 1000, 
     TRUE ~ chla_ugL)) |> 
  mutate(trophic = ifelse(trophic == "Other: according to authors, varies according to water level; never more than mesotrophic unless water level is low", "mesotrophic", trophic)) %>% 
  mutate(trophic = ifelse(trophic == "", "not reported", trophic)) %>% 
  # mutate(trophic_status = ifelse(trophic_status == "eutophic-hypertrophic", "hypertrophic", trophic_status)) %>% #change eutrophic-hypereutrophic to hypertrophic
  # mutate(trophic_status = ifelse(trophic_status == "hypereutrophic; eutrophic", "hypertrophic", trophic_status)) %>%
  # mutate(mean_chlorophyll_concentration_units = ifelse(mean_chlorophyll_concentration_units == "2mg/L at surface", "not reported", mean_chlorophyll_concentration_units)) %>% #take out 2 g/L, could not find in the paper when I double-checked
  # separate(mean_chlorophyll_concentration_units, sep = " |-|m", into = c("chla_ugL", "unit")) %>%
  # mutate(chla_ugL = ifelse(chla_ugL == "not",NA,chla_ugL)) %>%
  # mutate(chla_ugL = ifelse(chla_ugL == "",NA,chla_ugL)) %>%
  mutate(chla_ugL = as.numeric(chla_ugL)) %>% 
  mutate(tp_ugL = as.numeric(tp_ugL)) |> 
  mutate(trophic_status_chla = ifelse(trophic == "not reported" & chla_ugL <= 2.6, "oligotrophic", 
                                   ifelse(trophic == "not reported" & chla_ugL >2.6 & chla_ugL <=7.3, "mesotrophic", 
                                          ifelse(trophic == "not reported" & chla_ugL >7.3 & chla_ugL <= 56, "eutrophic", 
                                                 ifelse(trophic == "not reported" & chla_ugL >56, "hypertrophic",
                                                        ifelse(trophic == "not reported" & is.na(chla_ugL), trophic, trophic)))))) %>% 
  mutate(trophic_status_tp = ifelse(tp_ugL <= 12, "oligotrophic", 
                                     ifelse(tp_ugL >=12 & tp_ugL <=24, "mesotrophic", 
                                            ifelse(tp_ugL >24 & tp_ugL <= 70, "eutrophic", 
                                                   ifelse(tp_ugL >70, "hypertrophic",
                                                          ifelse(is.na(tp_ugL), trophic, NA)))))) %>%  
  mutate(trophic_status_combined = ifelse(trophic == "not reported" & !is.na(trophic_status_tp), trophic_status_tp,
                                          ifelse(trophic == "not reported" & !is.na(trophic_status_chla), trophic_status_chla, trophic)))

#|> 
  # select(c(mean_tp, tp, tp_ugL, unit_tp, mean_chla, chla, chla_ugL, unit_chla))
```

Data wrangling for histograms set. 
```{r data wrangling - histograms}
phyto_responses <- c("Increase in phytoplankton", "No significant response", "Not measured", "Decrease in phytoplankton")

data |> 
  filter(decrease_phyto_response %in% phyto_responses) |> 
  subset(!is.na(decrease_phyto_response)) %>% 
  ggplot(aes(x=decrease_phyto_response)) + 
  geom_bar(fill = "#117733") +
  xlab("Phytoplankton response to water level decrease") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "white")

data |> 
  filter(increase_phyto_response %in% phyto_responses) |> 
  subset(!is.na(increase_phyto_response)) %>% 
  ggplot(aes(x=increase_phyto_response)) + 
  geom_bar(fill = "#117733") +
  xlab("Phytoplankton response to water level increase") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "white")
  
cyano_responses <- c("Increase in cyanobacteria", "No significant response", "Not measured", "Decrease in cyanobacteria")

data |> 
  filter(increase_cyano_response %in% cyano_responses) |> 
  subset(!is.na(increase_cyano_response)) %>% 
  ggplot(aes(x=increase_cyano_response)) + 
  geom_bar(fill = "#88CCEE") +
  xlab("Cyanobacteria response to water level increase") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

data |> 
  filter(decrease_cyano_response %in% cyano_responses) |> 
  subset(!is.na(decrease_cyano_response)) %>% 
  ggplot(aes(x=decrease_cyano_response)) + 
  geom_bar(fill = "#88CCEE") +
  xlab("Cyanobacteria response to water level decrease") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

```

This set of chunks is for histograms of the categorical information that we pulled. 
Number of waterbodies
Country
fullpond_maxdepth
fullpond_maxvol
trophic_status
decrease_phyto
decrease_cyano
increase_phyto
increase_cyano

```{r quick info - histograms}
data |> 
  subset(!is.na(country)) %>% 
  mutate(country = ifelse(country == "", "multires", country)) |> 
  ggplot(aes(x=country)) + 
  geom_bar(fill = "#DDCC77") +
  xlab("Country") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

#consider subsetting these to phyto and cyano specific studies?
data |> 
  subset(!is.na(fullpond_maxdepth)) %>% 
  ggplot(aes(x=fullpond_maxdepth)) + 
  geom_histogram() +
  xlab("Max depth at Full Pond") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

trophicstat <- c("eutrophic", "meso-eutrophic", "hypereutrophic", "mesotrophic", "meso-oligotrophic", "oligo-mesotrophic", "oligotrophic", "not reported")

data |> 
  filter(trophic %in% trophicstat) |> 
  ggplot(aes(x= factor(trophic, c("hypereutrophic", "eutrophic","meso-eutrophic", "mesotrophic", "meso-oligotrophic", "oligo-mesotrophic", "oligotrophic", "not reported")))) + 
  geom_bar(fill = "#DDCC77") +
  xlab("Trophic Status") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

```

```{r map from latlongs}
na_strings <- c("NA", "N A", "N / A", "N/A", "N/ A", "Not Available", "NOt available", "", "not reported")

latlong <- data %>%
  select("lat_dd", "long_dd", "res_name") |> 
    replace_with_na_all(condition = ~.x %in% na_strings) |> 
    subset(!is.na(long_dd)) |>  
  subset(!is.na(lat_dd))

str(latlong)

latlong %>% 
  subset(!is.na(long_dd)) %>% 
  subset(!is.na(lat_dd)) |> 
  mapview(xcol = "long_dd", ycol = "lat_dd", crs = 4269, grid = FALSE)
```



This set of chunks is for past analyses that we've run. 

```{r get water level depth}
#create a column for max depth of the waterbody
#this uses the observed max depth. If the observed max depth is not available, it takes the full pool max depth.
#we are losing some information by converting to numeric. Will need to sort that out later!!!!
prop_dl <- data |>
  mutate(obs_maxdepth = ifelse(obs_maxdepth == "", NA, obs_maxdepth)) |> 
  mutate(usedepth = ifelse(is.na(obs_maxdepth), fullpond_maxdepth, obs_maxdepth)) |> 
  #select(c("obs_maxdepth", "fullpond_maxdepth", "usedepth")) |> 
  mutate(elev_decrease_max = as.numeric(elev_decrease_max)) |> 
  mutate(usedepth = as.numeric(usedepth)) |> 
  mutate(prop_1delta_wl_decrease = ifelse(is.na(elev_increase_max), elev_decrease_max/usedepth, NA)) |> 
  mutate(prop_alldelta_wl_decrease = elev_decrease_max/usedepth) |> 
  mutate(prop_alldelta_wl_decrease = ifelse(prop_alldelta_wl_decrease > 0, prop_alldelta_wl_decrease* -1, prop_alldelta_wl_decrease)) |> 
  mutate(prop_1delta_wl_decrease = ifelse(prop_1delta_wl_decrease > 0, prop_1delta_wl_decrease* -1, prop_1delta_wl_decrease)) |>
  mutate(prop_alldelta_wl_increase = elev_increase_max/usedepth) |> 
  mutate(prop_1delta_wl_increase = elev_increase_max/usedepth) |> 
  #select(c("obs_maxdepth", "fullpond_maxdepth", "usedepth", "prop_1delta_wl_decrease", "prop_alldelta_wl_decrease", "elev_decrease_max", "elev_increase_max", "prop_alldelta_wl_increase", "prop_1delta_wl_increase"))
  mutate(prop_delta_wl = ifelse(is.na(prop_alldelta_wl_decrease), prop_alldelta_wl_increase, prop_alldelta_wl_decrease)) |> 
  mutate(prop_delta_wl_priorityincrease = ifelse(is.na(prop_alldelta_wl_increase), prop_alldelta_wl_decrease, prop_alldelta_wl_increase))

prop_dl |> 
  ggplot(aes(x=prop_alldelta_wl_decrease)) + 
  geom_histogram()

prop_dl |> 
  ggplot(aes(x=prop_alldelta_wl_increase)) + 
  geom_histogram()
 
# mutate(prop_delta_wl = /depth_m) %>% 
```

```{r}

#27Aug25 this currently loses a number of reservoirs because they have multiple responses recorded. Anything that does not fit "no significant response" an increase or a decrease is excluded. 
#Still true as of Oct. 8
#I've actually discovered that because of the way that I've written the ifelse statement maybe something weird is happening?
#have fixed this so that if the prop_delta_wl is <0, then it can't use the phyto_binary_increase data
boxplotdata2 <- prop_dl %>%  #changed 0 and 1 for phyto_wl_increase :( after poster print :(
  #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
    mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 0, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 0, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>%
  # MEL NOTE: make sure we are proceeding from most specific to most general case here
    mutate(combined_phyto_response_binary = case_when(
  is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(phyto_binary_decrease) ~ phyto_binary_increase,
  TRUE ~ phyto_binary_decrease
)) |> 
    mutate(combined_phyto_response_binary_increasepriority = case_when(
 prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
  is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
  TRUE ~ phyto_binary_increase
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response_binary = case_when(
  is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(cyano_binary_decrease) ~ cyano_binary_increase,
  TRUE ~ cyano_binary_decrease
)) |>
       mutate(combined_cyano_response_binary_increasepriority = case_when(
  prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
  prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
  TRUE ~ cyano_binary_increase
)) |> 
   mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
  mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
  mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))

cyano_box_2 <- boxplotdata2 %>% 
  subset(!is.na(combined_cyano_response)) %>% 
  ggplot(aes(x= combined_cyano_response, y = prop_delta_wl, fill = combined_cyano_response)) +
  scale_fill_manual(values = c("lightblue", "lightblue", "lightgreen", "lightgreen", "darkblue"), name = "Cyanobacteria Response")+
  #scale_color_manual(values = c("lightblue", "blue", "darkblue", "darkgreen", "lightgreen"))+
  geom_boxplot() + 
  ylab("Proportion Change in Water Level") + xlab("Cyanobacteria Response to Water Level Fluctuations") +
  geom_jitter(fill="white", size=3, alpha=0.9, width = 0, shape = 21) +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))
  # scale_x_discrete(limits=c("0", "1")) +
  # scale_fill_discrete(name = "test", labels = c("Decrease in cyanobacteria in response to water level decreases", "Increase in cyanobacteria in response to water level decreases"))

# MEL NOTE: I got an error for this figure
# Error in `palette()`:
# ! Insufficient values in manual scale. 10 needed but only 5 provided.
cyano_box_2

phyto_box_2 <- boxplotdata2 %>% 
  subset(!is.na(combined_phyto_response)) %>% 
  ggplot(aes(x= combined_phyto_response, y = prop_delta_wl, fill = combined_phyto_response)) +
  scale_fill_manual(values = c("lightblue", "lightblue", "lightgreen", "lightgreen", "darkgreen", "green"), name = "Phytoplankton Response")+
  #scale_color_manual(values = c("lightblue", "blue", "darkblue", "darkgreen", "lightgreen"))+
  geom_boxplot() + 
  ylab("Proportion Change in Water Level") + xlab("Phytoplankton Response to Water Level Fluctuations") +
  geom_jitter(fill="white", size=3, alpha=0.9, width = 0, shape = 21) +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))
  # scale_x_discrete(limits=c("0", "1")) +
  # scale_fill_discrete(name = "test", labels = c("Decrease in cyanobacteria in response to water level decreases", "Increase in cyanobacteria in response to water level decreases"))
phyto_box_2

phytobox_wldecrease <- boxplotdata2 %>% 
  subset(!is.na(phyto_response_wl_down)) %>% 
  subset(!is.na(prop_alldelta_wl_decrease)) %>%
  ggplot(aes(x= phyto_response_wl_down, y = prop_alldelta_wl_decrease, fill = phyto_response_wl_down)) +
  scale_fill_manual(values = c("lightblue", "lightblue", "lightgreen", "lightgreen", "darkgreen", "green"), name = "Phytoplankton Response")+
  #scale_color_manual(values = c("lightblue", "blue", "darkblue", "darkgreen", "lightgreen"))+
  geom_boxplot() + 
  ylab("Proportion Change in Water Level") + xlab("Phytoplankton Response to Water Level Fluctuations") +
  geom_jitter(fill="white", size=3, alpha=0.9, width = 0, shape = 21) +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))

phytobox_wlincrease <- boxplotdata2 %>% 
  subset(!is.na(phyto_response_wl_up)) %>% 
  subset(!is.na(prop_alldelta_wl_increase)) %>%
  ggplot(aes(x= phyto_response_wl_up, y = prop_alldelta_wl_increase, fill = phyto_response_wl_up)) +
  scale_fill_manual(values = c("lightblue", "lightblue", "lightgreen", "lightgreen", "darkgreen", "green"), name = "Phytoplankton Response")+
  #scale_color_manual(values = c("lightblue", "blue", "darkblue", "darkgreen", "lightgreen"))+
  geom_boxplot() + 
  ylab("Proportion Change in Water Level") + xlab("Phytoplankton Response to Water Level Fluctuations") +
  geom_jitter(fill="white", size=3, alpha=0.9, width = 0, shape = 21) +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))
phytobox_wldecrease
phytobox_wlincrease

cyanobox_wlincrease <- boxplotdata2 %>% 
  subset(!is.na(cyano_response_wl_up)) %>% 
  subset(!is.na(prop_alldelta_wl_increase)) %>%
  ggplot(aes(x= cyano_response_wl_up, y = prop_alldelta_wl_increase, fill = cyano_response_wl_up)) +
  scale_fill_manual(values = c("lightblue", "lightblue", "lightgreen", "lightgreen", "darkgreen", "green"), name = "Cyanobacteria Response")+
  #scale_color_manual(values = c("lightblue", "blue", "darkblue", "darkgreen", "lightgreen"))+
  geom_boxplot() + 
  ylab("Proportion Change in Water Level") + xlab("Cyanobacteria Response to Water Level Fluctuations") +
  geom_jitter(fill="white", size=3, alpha=0.9, width = 0, shape = 21) +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))

# MEL NOTE: not sure I understand why there are multiple boxes for increases and decreases
cyanobox_wldecrease <- boxplotdata2 %>% 
  subset(!is.na(cyano_response_wl_down)) %>% 
  subset(!is.na(prop_alldelta_wl_decrease)) %>%
  ggplot(aes(x= cyano_response_wl_down, y = prop_alldelta_wl_decrease, fill = cyano_response_wl_down)) +
  scale_fill_manual(values = c("lightblue", "lightblue", "lightgreen", "lightgreen", "darkgreen", "green"), name = "Cyanobacteria Response")+
  #scale_color_manual(values = c("lightblue", "blue", "darkblue", "darkgreen", "lightgreen"))+
  geom_boxplot() + 
  ylab("Proportion Change in Water Level") + xlab("Cyanobacteria Response to Water Level Fluctuations") +
  geom_jitter(fill="white", size=3, alpha=0.9, width = 0, shape = 21) +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))

# MEL NOTE: didn't find these objects in the environment
cyanobox_wlincrease
cyanobox_wldecrease
```

```{r adding 0.5 as non significant response}
logregdata_p5 <- prop_dl %>%  
    mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 0.5, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 0.5, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
  mutate(combined_phyto_response_binary = case_when(
  is.na(phyto_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(phyto_binary_decrease) ~ phyto_binary_increase,
  TRUE ~ phyto_binary_decrease
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0.5, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0.5, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
    mutate(combined_cyano_response_binary = case_when(
  is.na(cyano_binary_decrease) & prop_delta_wl < 0 ~ NA_real_,
  is.na(cyano_binary_decrease) ~ cyano_binary_increase,
  TRUE ~ cyano_binary_decrease
)) |>
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response = ifelse(is.na(cyano_response_wl_down), cyano_response_wl_up, cyano_response_wl_down)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) 
```


```{r to figure out issues with odd Naselli-Flores data caught mistake}
# lookatme <- logregdata_p5 |> 
#   select(c(name_yr, title, notes, res_name, country, fullpond_maxdepth, trophic, prop_delta_wl, combined_phyto_response_binary, phyto_binary_increase, phyto_binary_decrease))
# head(logregdata_p5)
```


```{r}
phytoresponsebinary <- logregdata_p5 %>% 
  ggplot(aes(x = prop_delta_wl, y = combined_phyto_response_binary)) + 
  geom_point() + geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE) + 
  ylab("Phytoplankton response to changes in water level") + 
  xlab("Proportion water level change") +
    theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))

cyanoresponsebinary <- logregdata_p5 %>% 
  ggplot(aes(x = prop_delta_wl, y = combined_cyano_response_binary)) + 
  geom_point() + geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE) + 
  ylab("Cyanobacterial response to changes in water level") + 
  xlab("Proportion water level change") +   
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16))

phytoresponsebinary
cyanoresponsebinary

help <- boxplotdata2 |> 
  filter(!is.na(prop_delta_wl))
help1 <- boxplotdata2 |> 
  filter(!is.na(country))

```


Could use this code to get both halves of water level increases and decreases.
```{r wl increases and decreases together}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_decrease <- rbind(cyanodata, phytodata)
logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"

test <- logregdata %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes in water level") + 
  xlab("Proportion water level change") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
test

# logregform <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata, family=binomial) #this is weighed down because both phyto and cyano are in here. Oops.
# summary(logregform)

logregform <-  glm(formula=combined_phyto_response_binary ~ prop_delta_wl, data=logregdata_p5, family=binomial) #this is where it becomes a question of how we assign not significant responses. Technically, they aren't a decrease or an increase
summary(logregform)

logregdata_phyto <- logregdata %>% subset(type == "phyto")
logregdata_cyano <- logregdata %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
```


```{r}
cyanodecreasedata <- logregdata_p5 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- logregdata_p5 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- logregdata_p5 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- logregdata_p5 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- logregdata_p5 %>% 
  select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytodata <- logregdata_p5 %>% 
  select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata <- rbind(cyanodata, phytodata)

logregdata_phyto <- logregdata %>% subset(type == "phyto")
logregdata_cyano <- logregdata %>% subset(type == "cyano")
#"#7A524B","#3C7F29","#8DBAE9"

test <- logregdata %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes in water level") + 
  xlab("Proportion water level change") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
test

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
```

```{r remove non significant responses}
#remove non significant responses
logregdata_p5_2 <- logregdata_p5 |> 
  mutate(combined_cyano_response_binary = ifelse(combined_cyano_response_binary == 0.5, NA, combined_cyano_response_binary)) |> 
  mutate(combined_phyto_response_binary = ifelse(combined_phyto_response_binary == 0.5, NA, combined_phyto_response_binary))

cyanodecreasedata <- logregdata_p5_2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- logregdata_p5_2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- logregdata_p5_2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- logregdata_p5_2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- logregdata_p5_2 %>% 
  select(prop_delta_wl, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytodata <- logregdata_p5_2 %>% 
  select(prop_delta_wl, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata <- rbind(cyanodata, phytodata)

logregdata_phyto <- logregdata %>% subset(type == "phyto")
logregdata_cyano <- logregdata %>% subset(type == "cyano")
#"#7A524B","#3C7F29","#8DBAE9"

test <- logregdata %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes in water level") + 
  xlab("Proportion water level change") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
test

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
```


```{r logreg prioritize wl increases}
#I think this is still very wrong, will need some fiddling if pursued. 
#this makes new column for responses with a WL increase priority.

#I think I've reprioritized the responses to wl increases
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary_increasepriority, 
         response = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "increase")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary_increasepriority, 
         response = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")|> 
  mutate(wl_priority = "increase")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_increase <- rbind(cyanodata, phytodata)
#"#7A524B","#3C7F29","#8DBAE9"

test <- logregdata_increase %>% ggplot(aes(x = prop_delta_wl_priorityincrease, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes in water level") + 
  xlab("Proportion water level change") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
test

# logregform <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata, family=binomial) #this is weighed down because both phyto and cyano are in here. Oops.
# summary(logregform)

# logregform <-  glm(formula=combined_phyto_response_binary ~ prop_delta_wl_priorityincrease, data=boxplotdata2, family=binomial) #this is where it becomes a question of how we assign not significant responses. Technically, they aren't a decrease or an increase
# summary(logregform)

logregdata_phyto <- logregdata_increase %>% subset(type == "phyto")
logregdata_cyano <- logregdata_increase %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)

```

Logistic regression
```{r combined for wl increases and decreases}

logregdata_increase1 <- logregdata_increase |>
  rename(prop_delta_wl = prop_delta_wl_priorityincrease)

new_all <- bind_rows(logregdata_decrease, logregdata_increase1) |>
  select(-c(wl_priority)) |>
  unique()

all <- new_all %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) +
  #geom_jitter()+
 ylab("Response to changes in water level") +
  xlab("Proportion water level change") +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10),
        axis.title.y=element_text(angle=90, vjust = .5, size = 12),
        legend.title=element_blank())
all

logregdata_phyto <- new_all %>% subset(type == "phyto")
logregdata_cyano <- new_all %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
```


Entirely commented out this failed experiment of a chunk. working chunk is below.
```{r}
# #code from Mary, mosaic plots
# 
# # get data
# dat <- boxplotdata2 %>% 
#   select(cov_num, res_name, trophic_status_combined, tp_ugL, phyto_response_wl_down, phyto_response_wl_up,
#          cyano_response_wl_down, cyano_response_wl_up, chla_ugL, combined_phyto_response, combined_cyano_response) %>%
#   filter(!(is.na(phyto_response_wl_down) & is.na(phyto_response_wl_up) & is.na(cyano_response_wl_down) & is.na(cyano_response_wl_up))) %>%
#   mutate(increase_decrease_mosaic = ifelse(((!is.na(phyto_response_wl_down) | !is.na(cyano_response_wl_down)) & (!is.na(phyto_response_wl_up) | !is.na(cyano_response_wl_up))),"both",
#                                            ifelse((!is.na(phyto_response_wl_down) | !is.na(cyano_response_wl_down)),"Decrease",
#                                                   ifelse((!is.na(phyto_response_wl_up) | !is.na(cyano_response_wl_up)),"Increase",NA)))) 
# 
# dat <- boxplotdata2 %>% 
#   select(cov_num, res_name, trophic_status_combined, tp_ugL, phyto_response_wl_down, phyto_response_wl_up,
#          cyano_response_wl_down, cyano_response_wl_up, chla_ugL, combined_phyto_response, combined_cyano_response) %>%
#   filter(!(is.na(phyto_response_wl_down) & is.na(phyto_response_wl_up) & is.na(cyano_response_wl_down) & is.na(cyano_response_wl_up))) %>%
#   mutate(increase_decrease_mosaic = ifelse(((!is.na(phyto_response_wl_down) | !is.na(cyano_response_wl_down)) & (!is.na(phyto_response_wl_up) | !is.na(cyano_response_wl_up))),"both",
#                                            ifelse((!is.na(phyto_response_wl_down) | !is.na(cyano_response_wl_down)),"Decrease",
#                                                   ifelse((!is.na(phyto_response_wl_up) | !is.na(cyano_response_wl_up)),"Increase",NA)))) 
# 
# fluctuation_studies_keep_increase <- dat %>%
#   filter(increase_decrease_mosaic == "both") %>%
#   mutate(increase_decrease_mosaic = "Increase",
#          phyto_response_wl_down = NA,
#          cyano_response_wl_down = NA) 
# 
# fluctuation_studies_keep_decrease <- dat %>%
#   filter(increase_decrease_mosaic == "both") %>%
#   mutate(increase_decrease_mosaic = "Decrease",
#          phyto_response_wl_up = NA,
#          cyano_response_wl_up = NA) 
# 
# dat_no_fluctuation_studies <- dat %>%
#   filter(!increase_decrease_mosaic == "both")
# 
# dat2 <- bind_rows(dat_no_fluctuation_studies, fluctuation_studies_keep_increase) %>%
#   bind_rows(., fluctuation_studies_keep_decrease) %>%
#   mutate(phyto_response_wl_down = ifelse(is.na(phyto_response_wl_down),"not reported",phyto_response_wl_down),
#          phyto_response_wl_up = ifelse(is.na(phyto_response_wl_up),"not reported",phyto_response_wl_up),
#          cyano_response_wl_down = ifelse(is.na(cyano_response_wl_down),"not reported",cyano_response_wl_down),
#          cyano_response_wl_up = ifelse(is.na(cyano_response_wl_up),"not reported",cyano_response_wl_up)) %>%
#   mutate(trophic_status_mosaic = ifelse((grepl("oligo",trophic_status_combined) | grepl("meso",trophic_status_combined)),"oligo/mesotrophic",
#                                         ifelse(grepl("eu", trophic_status_combined),"eu/hypereutrophic","not reported")),
#          increase_phyto = ifelse((increase_decrease_mosaic == "Increase" & phyto_response_wl_up == "Increase"),"yes",
#                                  ifelse(increase_decrease_mosaic == "Increase" & phyto_response_wl_up == "not reported","not reported",
#                                         ifelse(increase_decrease_mosaic == "Decrease" & phyto_response_wl_down == "Increase","yes",
#                                                ifelse(increase_decrease_mosaic == "Decrease" & phyto_response_wl_down == "not reported","not reported","no")))),
#          increase_cyano = ifelse((increase_decrease_mosaic == "Increase" & cyano_response_wl_up == "Increase"),"yes",
#                                  ifelse(increase_decrease_mosaic == "Increase" & cyano_response_wl_up == "not reported","not reported",
#                                         ifelse(increase_decrease_mosaic == "Decrease" & cyano_response_wl_down == "Increase","yes",
#                                                ifelse(increase_decrease_mosaic == "Decrease" & cyano_response_wl_down == "not reported","not reported","no"))))) %>%
#   select(cov_num, res_name, increase_decrease_mosaic, trophic_status_mosaic, increase_phyto, increase_cyano) %>%
#   mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("not reported","eu/hypereutrophic","oligo/mesotrophic")),
#          increase_phyto = factor(increase_phyto, levels = c("yes","no","not reported")),
#          increase_cyano = factor(increase_cyano, levels = c("yes","no","not reported"))) 
# 
# # dat2 <- bind_rows(dat_no_fluctuation_studies, fluctuation_studies_keep_increase) %>%
# #   bind_rows(., fluctuation_studies_keep_decrease) %>%
# #   mutate(phyto_response_wl_down = ifelse(is.na(phyto_response_wl_down),"not reported",phyto_response_wl_down),
# #          phyto_response_wl_up = ifelse(is.na(phyto_response_wl_up),"not reported",phyto_response_wl_up),
# #          cyano_response_wl_down = ifelse(is.na(cyano_response_wl_down),"not reported",cyano_response_wl_down),
# #          cyano_response_wl_up = ifelse(is.na(cyano_response_wl_up),"not reported",cyano_response_wl_up)) %>%
# #  mutate(increase_phyto = case_when(increase_decrease_mosaic == "Increase" & phyto_response_wl_up == "Increase" ~ "yes",
# #             increase_decrease_mosaic == "Increase" & phyto_response_wl_up == "not reported" ~ "not reported",
# #             increase_decrease_mosaic == "Increase" & phyto_response_wl_up == "Decrease" ~ "no",
# #             increase_decrease_mosaic == "Decrease" & phyto_response_wl_down == "Increase" ~ "yes",
# #             increase_decrease_mosaic == "Decrease" & phyto_response_wl_down == "not reported" ~ "not reported",
# #             increase_decrease_mosaic == "Decrease" & phyto_response_wl_down == "Decrease" ~ "no"))
# # 
# # write.csv(dat2,"./Data/mosaic_plot_data_9OCT25.csv", row.names = FALSE)
# 
# dat_phyto1 <- dat2 %>%
#   filter(!increase_phyto == "not reported") %>%
#   select(trophic_status_mosaic, increase_decrease_mosaic, increase_phyto) %>%
#   mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_phyto, sep = "_")) %>%
#   group_by(x_axis_barplot, trophic_status_mosaic) %>%
#   summarise(count = n()) |>
#   mutate(response.count = sum(count),
#          ts.prop = count/sum(count)) |>
#   ungroup() %>%
#   mutate(response.prop = response.count/sum(count)) %>%
#   ungroup() %>%
#   separate(x_axis_barplot,c("increase_decrease_mosaic","increase_phyto"), remove = FALSE) %>%
#   mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("oligo/mesotrophic","eu/hypereutrophic","not reported")))
# 
# # New facet label names for water level/response variable
# wl.labs <- c("WL decrease","WL increase")
# names(wl.labs) <- unique(dat_phyto$increase_decrease_mosaic)
# response.labs <- c("no phyto \n increase", "phyto \n increase", "no phyto \n increase") 
# 
# # Calculate bar positions
# w <- c(unique(dat_phyto$response.prop)[1],unique(dat_phyto$response.prop))
# pos <- data.frame(pos = 0.5 * (cumsum(w) + cumsum(c(0, w[-length(w)]))),
#                   x_axis_barplot = unique(dat_phyto$x_axis_barplot))
# 
# dat_phyto <- left_join(dat_phyto1, pos, by = "x_axis_barplot")
# #rm(dat_phyto)
# 
# barplot_phyto <- ggplot(data = dat_phyto, aes(x = pos, y = ts.prop, width = response.prop, fill = trophic_status_mosaic)) +
#   geom_bar(stat = "identity", colour = "black") +
#   scale_x_continuous(breaks = unique(dat_phyto$pos), labels = response.labs, expand = c(0,0))+
#   scale_y_continuous(expand = c(0,0))+
#   geom_text(aes(label = paste0("n=",count)), position = position_stack(vjust = 0.5))+ # if labels are desired
#   facet_wrap(~increase_decrease_mosaic, scales = "free_x",labeller = labeller(increase_decrease_mosaic = wl.labs)) +
#   scale_fill_manual(values = c("#88CCEE","#117733","gray90")) +
#   theme_classic()+
#   theme(axis.ticks.y = element_blank(),
#         axis.text.y = element_blank(),
#         axis.title.y = element_blank(),
#         axis.title.x = element_blank())+  # if no spacing preferred between bars
#   labs(fill = "Trophic status",title = "Phytoplankton response to water level fluctuation")
#   
# barplot_phyto
# 
# ggsave(barplot_phyto, filename = "./Plots/barplot_phyto.png",dev = "png",width = 7, height = 4,
#        units = "in")
# 
# # cyanobacteria plot
# 
# dat_cyano <- dat2 %>%
#   filter(!increase_cyano == "not reported") %>%
#   select(trophic_status_mosaic, increase_decrease_mosaic, increase_cyano) %>%
#   mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_cyano, sep = "_")) %>%
#   group_by(x_axis_barplot, trophic_status_mosaic) %>%
#   summarise(count = n()) |>
#   mutate(response.count = sum(count),
#          ts.prop = count/sum(count)) |>
#   ungroup() %>%
#   mutate(response.prop = response.count/sum(count)) %>%
#   ungroup() %>%
#   separate(x_axis_barplot, c("increase_decrease_mosaic","increase_phyto"), remove = FALSE) %>%
#   mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("oligo/mesotrophic","eu/hypereutrophic","not reported")))
# 
# # New facet label names for water level/response variable
# wl.labs <- c("WL decrease","WL increase")
# names(wl.labs) <- unique(dat_cyano$increase_decrease_mosaic)
# response.labs <- c("no cyano \n increase", "cyano \n increase", "no cyano \n increase", "cyano \n increase") 
# 
# # Calculate bar positions
# w <- c(unique(dat_cyano$response.prop),unique(dat_cyano$response.prop)[3])
# pos <- data.frame(pos = 0.5 * (cumsum(w) + cumsum(c(0, w[-length(w)]))),
#                   x_axis_barplot = unique(dat_cyano$x_axis_barplot))
# 
# dat_cyano <- left_join(dat_cyano, pos, by = "x_axis_barplot")
# 
# 
# barplot_cyano <- ggplot(data = dat_cyano, aes(x = pos, y = ts.prop, width = response.prop, fill = trophic_status_mosaic)) +
#   geom_bar(stat = "identity", colour = "black") +
#   scale_x_continuous(breaks = unique(dat_cyano$pos), labels = response.labs, expand = c(0,0))+
#   scale_y_continuous(expand = c(0,0))+
#   geom_text(aes(label = paste0("n=",count)), position = position_stack(vjust = 0.5))+ # if labels are desired
#   facet_wrap(~increase_decrease_mosaic, scales = "free_x",labeller = labeller(increase_decrease_mosaic = wl.labs)) +
#   scale_fill_manual(values = c("#88CCEE","#117733","gray90")) +
#   theme_classic()+
#   theme(axis.ticks.y = element_blank(),
#         axis.text.y = element_blank(),
#         axis.title.y = element_blank(),
#         axis.title.x = element_blank())+  # if no spacing preferred between bars
#   labs(fill = "Trophic status",title = "Cyanobacteria response to water level fluctuation")
# 
# barplot_cyano
# 
# ggsave(barplot_cyano, filename = "./Plots/barplot_cyano.png",dev = "png",width = 7, height = 4,
#        units = "in")
# 
# ## From Agresti(2007) p.39
# 
# # first for phytoplankton
# phyto <- dat2 %>%
#   select(increase_decrease_mosaic, increase_phyto)
# phyto_increase_wl_increase <- phyto %>%
#   filter(increase_decrease_mosaic == "increase" & increase_phyto == "yes") %>%
#   count(increase_phyto) %>%
#   pull(n)
# phyto_no_increase_wl_increase <- phyto %>%
#   filter(increase_decrease_mosaic == "increase" & increase_phyto == "no") %>%
#   count(increase_phyto) %>%
#   pull(n)
# phyto_increase_wl_decrease <- phyto %>%
#   filter(increase_decrease_mosaic == "decrease" & increase_phyto == "yes") %>%
#   count(increase_phyto) %>%
#   pull(n)
# phyto_no_increase_wl_decrease <- phyto %>%
#   filter(increase_decrease_mosaic == "decrease" & increase_phyto == "no") %>%
#   count(increase_phyto) %>%
#   pull(n)
# 
# # Question 1: are there differences in phytoplankton responses to increases vs.
# # decreases in water level?
# M <- as.table(rbind(c(phyto_increase_wl_increase, phyto_increase_wl_decrease), c(phyto_no_increase_wl_increase, phyto_no_increase_wl_decrease)))
# dimnames(M) <- list(phyto_response = c("increase", "decrease or no change"),
#                     water_level = c("increase","decrease"))
# M
# (Xsq <- chisq.test(M))  # Prints test summary
# Xsq$observed   # observed counts (same as M)
# Xsq$expected   # expected counts under the null
# Xsq$residuals  # Pearson residuals
# Xsq$stdres     # standardized residuals
# Xsq$p.value
# 
# # Question 2: are there differences in the likelihood of phytoplankton to increase
# # or decrease given an increase in water level?
# M_increase <- M[,1]
# M_increase
# (Xsq <- chisq.test(M_increase))  # Prints test summary
# Xsq$observed   # observed counts (same as M)
# Xsq$expected   # expected counts under the null
# Xsq$residuals  # Pearson residuals
# Xsq$stdres     # standardized residuals
# Xsq$p.value
# 
# # Question 3: are there differences in the likelihood of phytoplankton to increase
# # or decrease given a decrease in water level?
# M_decrease <- M[,2]
# M_decrease
# (Xsq <- chisq.test(M_decrease))  # Prints test summary
# Xsq$observed   # observed counts (same as M)
# Xsq$expected   # expected counts under the null
# Xsq$residuals  # Pearson residuals
# Xsq$stdres     # standardized residuals
# Xsq$p.value
# 
# # now for cyanobacteria
# cyano <- dat2 %>%
#   select(increase_decrease_mosaic, increase_cyano)
# cyano_increase_wl_increase <- cyano %>%
#   filter(increase_decrease_mosaic == "increase" & increase_cyano == "yes") %>%
#   count(increase_cyano) %>%
#   pull(n)
# cyano_no_increase_wl_increase <- cyano %>%
#   filter(increase_decrease_mosaic == "increase" & increase_cyano == "no") %>%
#   count(increase_cyano) %>%
#   pull(n)
# cyano_increase_wl_decrease <- cyano %>%
#   filter(increase_decrease_mosaic == "decrease" & increase_cyano == "yes") %>%
#   count(increase_cyano) %>%
#   pull(n)
# cyano_no_increase_wl_decrease <- cyano %>%
#   filter(increase_decrease_mosaic == "decrease" & increase_cyano == "no") %>%
#   count(increase_cyano) %>%
#   pull(n)
# 
# # Question 1: are there differences in cyanobacterial responses to increases vs.
# # decreases in water level?
# M <- as.table(rbind(c(cyano_increase_wl_increase, cyano_increase_wl_decrease), c(cyano_no_increase_wl_increase, cyano_no_increase_wl_decrease)))
# dimnames(M) <- list(cyano_response = c("increase", "decrease or no change"),
#                     water_level = c("increase","decrease"))
# M
# (Xsq <- chisq.test(M))  # Prints test summary
# Xsq$observed   # observed counts (same as M)
# Xsq$expected   # expected counts under the null
# Xsq$residuals  # Pearson residuals
# Xsq$stdres     # standardized residuals
# Xsq$p.value
# 
# # Question 2: are there differences in the likelihood of cyanobacteria to increase
# # or decrease given an increase in water level?
# M_increase <- M[,1]
# M_increase
# (Xsq <- chisq.test(M_increase))  # Prints test summary
# Xsq$observed   # observed counts (same as M)
# Xsq$expected   # expected counts under the null
# Xsq$residuals  # Pearson residuals
# Xsq$stdres     # standardized residuals
# Xsq$p.value
# 
# # Question 3: are there differences in the likelihood of cyanobacteria to increase
# # or decrease given a decrease in water level?
# M_decrease <- M[,2]
# M_decrease
# (Xsq <- chisq.test(M_decrease))  # Prints test summary
# Xsq$observed   # observed counts (same as M)
# Xsq$expected   # expected counts under the null
# Xsq$residuals  # Pearson residuals
# Xsq$stdres     # standardized residuals
# Xsq$p.value
# 
# 
# ggplot(data = dat, aes(x = trophic_status_combined))+
#   geom_bar(stat = "count")
```

Chi squares based on Mary's code. I think this is correct but not 100%, have someone take a look before pubs. 
```{r chisq use me}
dat <- boxplotdata2 %>% 
  select(cov_num, res_name, trophic_status_combined, tp_ugL, decrease_phyto_response, increase_phyto_response,
         decrease_cyano_response, increase_cyano_response, chla_ugL, combined_phyto_response, combined_cyano_response) %>%
  mutate(increase_phyto_response = ifelse(increase_phyto_response == "Not measured"|increase_phyto_response == ""|(grepl("Other", increase_phyto_response)), NA, increase_phyto_response)) |>
    mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured"|decrease_phyto_response == ""|(grepl("Other", decrease_phyto_response)), NA, decrease_phyto_response)) |>
   mutate(increase_cyano_response = ifelse(increase_cyano_response == "Not measured"|increase_cyano_response == ""|(grepl("Other", increase_cyano_response)), NA, increase_cyano_response)) |>
   mutate(decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured"|decrease_cyano_response == ""|(grepl("Other", decrease_cyano_response)), NA, decrease_cyano_response)) |>
  filter(!(is.na(decrease_phyto_response) & is.na(increase_phyto_response) & is.na(decrease_cyano_response) & is.na(increase_cyano_response))) %>%
  mutate(increase_decrease_mosaic = ifelse(((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)) & (!is.na(increase_phyto_response) | !is.na(increase_cyano_response))),"both",
                                           ifelse((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)),"Decrease",
                                                  ifelse((!is.na(increase_phyto_response) | !is.na(increase_cyano_response)),"Increase", NA)))) 

# dat <- boxplotdata2 %>% 
#   select(cov_num, res_name, trophic_status_combined, tp_ugL, decrease_phyto_response, increase_phyto_response,
#          decrease_cyano_response, increase_cyano_response, chla_ugL, combined_phyto_response, combined_cyano_response) %>%
#   filter(!(is.na(decrease_phyto_response) & is.na(increase_phyto_response) & is.na(decrease_cyano_response) & is.na(increase_cyano_response))) %>%
#   mutate(increase_decrease_mosaic = ifelse(((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)) & (!is.na(increase_phyto_response) | !is.na(increase_cyano_response))),"both",
#                                            ifelse((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)),"Decrease",
#                                                   ifelse((!is.na(increase_phyto_response) | !is.na(increase_cyano_response)),"Increase",NA)))) 

fluctuation_studies_keep_increase <- dat %>%
  filter(increase_decrease_mosaic == "both") %>%
  mutate(increase_decrease_mosaic = "Increase",
         decrease_phyto_response = NA,
         decrease_cyano_response = NA) 

fluctuation_studies_keep_decrease <- dat %>%
  filter(increase_decrease_mosaic == "both") %>%
  mutate(increase_decrease_mosaic = "Decrease",
         increase_phyto_response = NA,
         increase_cyano_response = NA) 

dat_no_fluctuation_studies <- dat %>%
  filter(!increase_decrease_mosaic == "both")

dat2 <- bind_rows(dat_no_fluctuation_studies, fluctuation_studies_keep_increase) %>%
  bind_rows(., fluctuation_studies_keep_decrease) %>%
  mutate(decrease_phyto_response = ifelse(is.na(decrease_phyto_response),"not reported",decrease_phyto_response),
         increase_phyto_response = ifelse(is.na(increase_phyto_response),"not reported",increase_phyto_response),
         decrease_cyano_response = ifelse(is.na(decrease_cyano_response),"not reported",decrease_cyano_response),
         increase_cyano_response = ifelse(is.na(increase_cyano_response),"not reported",increase_cyano_response)) %>%
  mutate(trophic_status_mosaic = ifelse((grepl("oligo",trophic_status_combined) | grepl("meso",trophic_status_combined)),"oligo/mesotrophic",
                                        ifelse(grepl("eu", trophic_status_combined),"eu/hypereutrophic","not reported")),
         increase_phyto = ifelse((increase_decrease_mosaic == "Increase" & increase_phyto_response == "Increase in phytoplankton"),"yes",
                                 ifelse(increase_decrease_mosaic == "Increase" & increase_phyto_response == "not reported","not reported",
                                        ifelse(increase_decrease_mosaic == "Decrease" & decrease_phyto_response == "Increase in phytoplankton","yes",
                                               ifelse(increase_decrease_mosaic == "Decrease" & decrease_phyto_response == "not reported","not reported","no")))),
         increase_cyano = ifelse((increase_decrease_mosaic == "Increase" & increase_cyano_response == "Increase in cyanobacteria"),"yes",
                                 ifelse(increase_decrease_mosaic == "Increase" & increase_cyano_response == "not reported","not reported",
                                        ifelse(increase_decrease_mosaic == "Decrease" & decrease_cyano_response == "Increase in cyanobacteria","yes",
                                               ifelse(increase_decrease_mosaic == "Decrease" & decrease_cyano_response == "not reported","not reported","no"))))) %>%
  select(cov_num, res_name, increase_decrease_mosaic, trophic_status_mosaic, increase_phyto, increase_cyano) %>%
  mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("not reported","eu/hypereutrophic","oligo/mesotrophic")),
         increase_phyto = factor(increase_phyto, levels = c("yes","no","not reported")),
         increase_cyano = factor(increase_cyano, levels = c("yes","no","not reported"))) 

# dat2 <- bind_rows(dat_no_fluctuation_studies, fluctuation_studies_keep_increase) %>%
#   bind_rows(., fluctuation_studies_keep_decrease) %>%
#   mutate(decrease_phyto_response = ifelse(is.na(decrease_phyto_response),"not reported",decrease_phyto_response),
#          increase_phyto_response = ifelse(is.na(increase_phyto_response),"not reported",increase_phyto_response),
#          decrease_cyano_response = ifelse(is.na(decrease_cyano_response),"not reported",decrease_cyano_response),
#          increase_cyano_response = ifelse(is.na(increase_cyano_response),"not reported",increase_cyano_response)) %>%
#  mutate(increase_phyto = case_when(increase_decrease_mosaic == "Increase" & increase_phyto_response == "Increase" ~ "yes",
#             increase_decrease_mosaic == "Increase" & increase_phyto_response == "not reported" ~ "not reported",
#             increase_decrease_mosaic == "Increase" & increase_phyto_response == "Decrease" ~ "no",
#             increase_decrease_mosaic == "Decrease" & decrease_phyto_response == "Increase" ~ "yes",
#             increase_decrease_mosaic == "Decrease" & decrease_phyto_response == "not reported" ~ "not reported",
#             increase_decrease_mosaic == "Decrease" & decrease_phyto_response == "Decrease" ~ "no"))
# 
# write.csv(dat2,"./Data/mosaic_plot_data_9OCT25.csv", row.names = FALSE)

# MEL NOTE: Katie can you remind me why this additional code is needed? (lines 1260-1303)
# especially because it is only done for phytoplankton and not cyanobacteria -
# is all of this to be able to show 0 studies falling into the phyto increase when
# water level increases in eu/hypereutrophic lakes?

dat_phyto1 <- dat2 %>%
  filter(!increase_phyto == "not reported") %>%
  select(trophic_status_mosaic, increase_decrease_mosaic, increase_phyto) %>%
  mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_phyto, sep = "_")) %>%
  group_by(x_axis_barplot, trophic_status_mosaic) %>%
  summarise(count = n()) |>
  mutate(response.count = sum(count),
         ts.prop = count/sum(count)) |>
  ungroup() %>%
  mutate(response.prop = response.count/sum(count)) %>%
  ungroup() %>%
  separate(x_axis_barplot,c("increase_decrease_mosaic","increase_phyto"), remove = FALSE) %>%
  mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("oligo/mesotrophic","eu/hypereutrophic","not reported")))

library(dplyr)
library(tidyr)

dat_phyto1 <- dat2 %>%
  filter(!increase_phyto == "not reported") %>%
  select(trophic_status_mosaic, increase_decrease_mosaic, increase_phyto) %>%
  mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_phyto, sep = "_")) %>%
  group_by(x_axis_barplot, trophic_status_mosaic) %>%
  summarise(count = n(), .groups = "drop") %>%

  # Ensure all combinations of x_axis_barplot and trophic_status_mosaic exist
  complete(
    x_axis_barplot,
    trophic_status_mosaic,
    fill = list(count = 0)
  ) %>%

  # Recompute decomposed vars after completing
  separate(x_axis_barplot, c("increase_decrease_mosaic", "increase_phyto"), remove = FALSE) %>%
  group_by(x_axis_barplot) %>%
  mutate(response.count = sum(count),
         ts.prop = count / sum(count)) %>%
  ungroup() %>%
  mutate(response.prop = response.count / sum(count)) %>%
  mutate(
    trophic_status_mosaic = factor(
      trophic_status_mosaic,
      levels = c("oligo/mesotrophic", "eu/hypereutrophic", "not reported")
    )
  )

# New facet label names for water level/response variable
wl.labs <- c("WL decrease","WL increase")
names(wl.labs) <- unique(dat_phyto1$increase_decrease_mosaic)
response.labs <- c("no phyto \n increase", "phyto \n increase", "no phyto \n increase", "phyto \n increase") 

# Calculate bar positions
w <- c(unique(dat_phyto1$response.prop))
pos <- data.frame(pos = 0.5 * (cumsum(w) + cumsum(c(0, w[-length(w)]))),
                  x_axis_barplot = unique(dat_phyto1$x_axis_barplot))

dat_phyto <- left_join(dat_phyto1, pos, by = "x_axis_barplot")
#rm(dat_phyto)

barplot_phyto <- ggplot(data = dat_phyto, aes(x = pos, y = ts.prop, width = response.prop, fill = trophic_status_mosaic)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_x_continuous(breaks = unique(dat_phyto$pos), labels = response.labs, expand = c(0,0))+
  scale_y_continuous(expand = c(0,0))+
  geom_text(aes(label = paste0("n=",count)), position = position_stack(vjust = 0.5))+ # if labels are desired
  facet_wrap(~increase_decrease_mosaic, scales = "free_x",labeller = labeller(increase_decrease_mosaic = wl.labs)) +
  scale_fill_manual(values = c("#88CCEE","#117733","gray90")) +
  theme_classic()+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank())+  # if no spacing preferred between bars
  labs(fill = "Trophic status",title = "Phytoplankton response to water level fluctuation")
  
barplot_phyto

#ggsave(barplot_phyto, filename = "./Plots/barplot_phyto.png",dev = "png",width = 7, height = 4,
#       units = "in")

# cyanobacteria plot

dat_cyano <- dat2 %>%
  filter(!increase_cyano == "not reported") %>%
  select(trophic_status_mosaic, increase_decrease_mosaic, increase_cyano) %>%
  mutate(x_axis_barplot = paste(increase_decrease_mosaic, increase_cyano, sep = "_")) %>%
  group_by(x_axis_barplot, trophic_status_mosaic) %>%
  summarise(count = n()) |>
  mutate(response.count = sum(count),
         ts.prop = count/sum(count)) |>
  ungroup() %>%
  mutate(response.prop = response.count/sum(count)) %>%
  ungroup() %>%
  separate(x_axis_barplot, c("increase_decrease_mosaic","increase_phyto"), remove = FALSE) %>%
  mutate(trophic_status_mosaic = factor(trophic_status_mosaic, levels = c("oligo/mesotrophic","eu/hypereutrophic","not reported")))

# New facet label names for water level/response variable
wl.labs <- c("WL decrease","WL increase")
names(wl.labs) <- unique(dat_cyano$increase_decrease_mosaic)
response.labs <- c("no cyano \n increase", "cyano \n increase", "no cyano \n increase", "cyano \n increase") 

# Calculate bar positions
w <- c(unique(dat_cyano$response.prop))
pos <- data.frame(pos = 0.5 * (cumsum(w) + cumsum(c(0, w[-length(w)]))),
                  x_axis_barplot = unique(dat_cyano$x_axis_barplot))

dat_cyano <- left_join(dat_cyano, pos, by = "x_axis_barplot")


barplot_cyano <- ggplot(data = dat_cyano, aes(x = pos, y = ts.prop, width = response.prop, fill = trophic_status_mosaic)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_x_continuous(breaks = unique(dat_cyano$pos), labels = response.labs, expand = c(0,0))+
  scale_y_continuous(expand = c(0,0))+
  geom_text(aes(label = paste0("n=",count)), position = position_stack(vjust = 0.5))+ # if labels are desired
  facet_wrap(~increase_decrease_mosaic, scales = "free_x",labeller = labeller(increase_decrease_mosaic = wl.labs)) +
  scale_fill_manual(values = c("#88CCEE","#117733","gray90")) +
  theme_classic()+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank())+  # if no spacing preferred between bars
  labs(fill = "Trophic status",title = "Cyanobacteria response to water level fluctuation")

barplot_cyano

# ggsave(barplot_cyano, filename = "./Plots/barplot_cyano.png",dev = "png",width = 7, height = 4,
#        units = "in")

## From Agresti(2007) p.39

# first for phytoplankton
phyto <- dat2 %>%
  select(increase_decrease_mosaic, increase_phyto)
phyto_increase_wl_increase <- phyto %>%
  filter(increase_decrease_mosaic == "Increase" & increase_phyto == "yes") %>%
  count(increase_phyto) %>%
  pull(n)
phyto_no_increase_wl_increase <- phyto %>%
  filter(increase_decrease_mosaic == "Increase" & increase_phyto == "no") %>%
  count(increase_phyto) %>%
  pull(n)
phyto_increase_wl_decrease <- phyto %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_phyto == "yes") %>%
  count(increase_phyto) %>%
  pull(n)
phyto_no_increase_wl_decrease <- phyto %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_phyto == "no") %>%
  count(increase_phyto) %>%
  pull(n)

# Question 1: are there differences in phytoplankton responses to increases vs.
# decreases in water level?
M <- as.table(rbind(c(phyto_increase_wl_increase, phyto_increase_wl_decrease), c(phyto_no_increase_wl_increase, phyto_no_increase_wl_decrease)))
dimnames(M) <- list(phyto_response = c("increase", "decrease or no change"),
                    water_level = c("increase","decrease"))
M
(Xsq <- chisq.test(M))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 2: are there differences in the likelihood of phytoplankton to increase
# or decrease given an increase in water level?
M_increase <- M[,1]
M_increase
(Xsq <- chisq.test(M_increase))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 3: are there differences in the likelihood of phytoplankton to increase
# or decrease given a decrease in water level?
M_decrease <- M[,2]
M_decrease
(Xsq <- chisq.test(M_decrease))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# now for cyanobacteria
cyano <- dat2 %>%
  select(increase_decrease_mosaic, increase_cyano)
cyano_increase_wl_increase <- cyano %>%
  filter(increase_decrease_mosaic == "Increase" & increase_cyano == "yes") %>%
  count(increase_cyano) %>%
  pull(n)
cyano_no_increase_wl_increase <- cyano %>%
  filter(increase_decrease_mosaic == "Increase" & increase_cyano == "no") %>%
  count(increase_cyano) %>%
  pull(n)
cyano_increase_wl_decrease <- cyano %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_cyano == "yes") %>%
  count(increase_cyano) %>%
  pull(n)
cyano_no_increase_wl_decrease <- cyano %>%
  filter(increase_decrease_mosaic == "Decrease" & increase_cyano == "no") %>%
  count(increase_cyano) %>%
  pull(n)

# Question 1: are there differences in cyanobacterial responses to increases vs.
# decreases in water level?
M <- as.table(rbind(c(cyano_increase_wl_increase, cyano_increase_wl_decrease), c(cyano_no_increase_wl_increase, cyano_no_increase_wl_decrease)))
dimnames(M) <- list(cyano_response = c("increase", "decrease or no change"),
                    water_level = c("increase","decrease"))
M
(Xsq <- chisq.test(M))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 2: are there differences in the likelihood of cyanobacteria to increase
# or decrease given an increase in water level?
M_increase <- M[,1]
M_increase
(Xsq <- chisq.test(M_increase))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value

# Question 3: are there differences in the likelihood of cyanobacteria to increase
# or decrease given a decrease in water level?
M_decrease <- M[,2]
M_decrease
(Xsq <- chisq.test(M_decrease))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals
Xsq$p.value


ggplot(data = dat, aes(x = trophic_status_combined))+
  geom_bar(stat = "count")

```

